---
title: "MASEM on Nohe et al. (2015) data"
author: "Suzanne Jak and Mike Cheung"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
  word_document: default
editor_options: 
  chunk_output_type: console
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 100)
knitr::opts_chunk$set(message=FALSE)
```

# TSSEM (with complete data)

## A model without any moderator
```{r}
library(metaSEM)

## Proposed model in lavaan syntax
model1 <- 'W2 ~ w2w*W1 + s2w*S1
           S2 ~ w2s*W1 + s2s*S1
           W1 ~~ w1WITHs1*S1
           W2 ~~ w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ Errw2*W2
           S2 ~~ Errs2*S2'
     
RAM1 <- lavaan2RAM(model1, obs.variables=c("W1", "S1", "W2", "S2"))
RAM1

## Display the number of data points
pattern.na(Nohe15A1$data, show.na=FALSE)

## Stage 1 analysis
random1 <- tssem1(Nohe15A1$data, Nohe15A1$n, method="REM", RE.type="Diag")
summary(random1)

## Stage 2 analysis
random2 <- tssem2(random1, Amatrix=RAM1$A, Smatrix=RAM1$S)
summary(random2)

## Plot the model
plot(random2, col="green")
```

## Models with three subgroup analysis
```{r}
## Get the necessary functions
## source("http://www.suzannejak.nl/subgroup.functions.R")
source("./subgroup.functions.R")

data <- Nohe15A1$data
n <- Nohe15A1$n
Lag <- Nohe15A1$Lag

# Data for studies with short Lag 
data_g1 <- data[Lag<7]
n_g1 <- n[Lag<7]

# Data for studies with medium Lag 
data_g2 <- data[Lag>=7&Lag<13]
n_g2 <- n[Lag>=7&Lag<13]

# Data for studies with long Lag 
data_g3 <- data[Lag>=13]
n_g3 <- n[Lag>=13]
```

### Fitting a random-effects Stage 1 model in three subgroups
```{r, message=FALSE}
## Stage 1 analysis per subgroup (random-effects analysis)
stage1_g1.fit <- tssem1(Cov = data_g1, n = n_g1, method = "REM", RE.type = "Diag")
stage1_g2.fit <- tssem1(Cov = data_g2, n = n_g2, method = "REM", RE.type = "Diag")
stage1_g3.fit <- tssem1(Cov = data_g3, n = n_g3, method = "REM", RE.type = "Diag")

## Rerun it to remove the error code
stage1_g3.fit <- rerun(stage1_g3.fit)

## Results
summary(stage1_g1.fit)
summary(stage1_g2.fit)
summary(stage1_g3.fit)
```

### Fitting the Stage 2 model in three subgroups
```{r}
## Stage 2 analysis per subgroup (random-effect analysis)
stage2_g1.fit <- tssem2(stage1_g1.fit, Amatrix=RAM1$A, Smatrix=RAM1$S)
stage2_g2.fit <- tssem2(stage1_g2.fit, Amatrix=RAM1$A, Smatrix=RAM1$S)
stage2_g3.fit <- tssem2(stage1_g3.fit, Amatrix=RAM1$A, Smatrix=RAM1$S)

## Results
summary(stage2_g1.fit)
summary(stage2_g2.fit)
summary(stage2_g3.fit)
```

### Testing the equality of regression coefficients
* We create and fit a model with equal direct effects (we use the same matrix A for both groups), but different variances and covariances, so we create an S matrix with different labels for group 2 and group 3.
```{r}
## Proposed model g2
model2 <- 'W2 ~ w2w*W1 + s2w*S1
           S2 ~ w2s*W1 + s2s*S1
           W1 ~~ g2w1WITHs1*S1
           W2 ~~ g2w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ g2Errw2*W2
           S2 ~~ g2Errs2*S2'
     
RAM2 <- lavaan2RAM(model2, obs.variables=c("W1", "S1", "W2", "S2"))

## Proposed model g3
model3 <- 'W2 ~ w2w*W1 + s2w*S1
           S2 ~ w2s*W1 + s2s*S1
           W1 ~~ g3w1WITHs1*S1
           W2 ~~ g3w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ g3Errw2*W2
           S2 ~~ g3Errs2*S2'
     
RAM3 <- lavaan2RAM(model3, obs.variables=c("W1", "S1", "W2", "S2"))

## Create the models for the two groups, make sure to set the argument run=FALSE
stage2_g1 <- tssem2(stage1_g1.fit, Amatrix=RAM1$A, Smatrix=RAM1$S, run=FALSE, model.name="g1")

stage2_g2 <- tssem2(stage1_g2.fit, Amatrix=RAM1$A, Smatrix=RAM2$S, run=FALSE, model.name="g2")

stage2_g3 <- tssem2(stage1_g3.fit, Amatrix=RAM1$A, Smatrix=RAM3$S, run=FALSE, model.name="g3")

## Create the multigroup model
stage2_constrained <- mxModel(model="same_regression_coef", stage2_g1, stage2_g2,stage2_g3,
                              mxFitFunctionMultigroup(c("g1", "g2", "g3")))

## Fit multigroup model with equality constraints
Stage2_constrained.fit <- mxRun(stage2_constrained, intervals=TRUE)

## first make a list of the fitted models in the separate groups
submodels.fit <- list(stage2_g1.fit, stage2_g2.fit, stage2_g3.fit)

subgroup.summary(submodels.fit, Stage2_constrained.fit)
```

### Testing the equality of one regression coefficient (w2w)
* We create and fit a model with equal direct effects (we use the same matrix A for both groups), but different variances and covariances, so we create an S matrix with different labels for group 2 and group 3.

```{r}
## Proposed model g2
model2 <- 'W2 ~ w2w*W1 + g2s2w*S1
           S2 ~ g2w2s*W1 + g2s2s*S1
           W1 ~~ g2w1WITHs1*S1
           W2 ~~ g2w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ g2Errw2*W2
           S2 ~~ g2Errs2*S2'
     
RAM2 <- lavaan2RAM(model2, obs.variables=c("W1", "S1", "W2", "S2"))

## Proposed model g3
model3 <- 'W2 ~ w2w*W1 + g3s2w*S1
           S2 ~ g3w2s*W1 + g3s2s*S1
           W1 ~~ g3w1WITHs1*S1
           W2 ~~ g3w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ g3Errw2*W2
           S2 ~~ g3Errs2*S2'
     
RAM3 <- lavaan2RAM(model3, obs.variables=c("W1", "S1", "W2", "S2"))

## Create the models for the two groups, make sure to set the argument run=FALSE
stage2_g1 <- tssem2(stage1_g1.fit, Amatrix=RAM1$A, Smatrix=RAM1$S, run=FALSE, model.name="g1")

stage2_g2 <- tssem2(stage1_g2.fit, Amatrix=RAM2$A, Smatrix=RAM2$S, run=FALSE, model.name="g2")

stage2_g3 <- tssem2(stage1_g3.fit, Amatrix=RAM3$A, Smatrix=RAM3$S, run=FALSE, model.name="g3")

## Create the multigroup model
stage2_constrained <- mxModel(model="same_regression_coef", stage2_g1, stage2_g2,stage2_g3,
                              mxFitFunctionMultigroup(c("g1", "g2", "g3")))

## Fit multigroup model with equality constraints
Stage2_constrained.fit <- mxRun(stage2_constrained, intervals=TRUE)

## First make a list of the fitted models in the separate groups
submodels.fit <- list(stage2_g1.fit,stage2_g2.fit,stage2_g3.fit)

subgroup.summary(submodels.fit,Stage2_constrained.fit)
```

## Models with two subgroup analysis
```{r}
# Data for studies with short Lag 
data_g1 <- data[Lag<12]
n_g1 <- n[Lag<12]

# Data for studies with long Lag 
data_g2 <- data[Lag>=12]
n_g2 <- n[Lag>=12]
```

### Fitting a random-effects Stage 1 model in two subgroups
```{r, message=FALSE}
## Stage 1 analysis per subgroup (random-effects analysis)
stage1_g1.fit <- tssem1(Cov = data_g1, n = n_g1, method = "REM", RE.type = "Diag")
stage1_g2.fit <- tssem1(Cov = data_g2, n = n_g2, method = "REM", RE.type = "Diag")
## Rerun the analysis
stage1_g2.fit <- rerun(stage1_g2.fit)

summary(stage1_g1.fit)
summary(stage1_g2.fit)
```

### Fitting the Stage 2 model in both subgroups
```{r}
## Stage 2 analysis per subgroup (random-effect analysis)
stage2_g1.fit <- tssem2(stage1_g1.fit, Amatrix=RAM1$A, Smatrix=RAM1$S)
stage2_g2.fit <- tssem2(stage1_g2.fit, Amatrix=RAM1$A, Smatrix=RAM1$S)

summary(stage2_g1.fit)
summary(stage2_g2.fit)
```

### Testing the equality of regression coefficients
* We create and fit a model with equal direct effects (we use the same matrix A for both groups), but different variances and covariances, so we create an S matrix with different labels for group 2.

```{r}
## Proposed model g2
model2 <- 'W2 ~ w2w*W1 + s2w*S1
           S2 ~ w2s*W1 + s2s*S1
           W1 ~~ g2w1WITHs1*S1
           W2 ~~ g2w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ g2Errw2*W2
           S2 ~~ g2Errs2*S2'
     
RAM2 <- lavaan2RAM(model2, obs.variables=c("W1", "S1", "W2", "S2"))

# Create the models for the two groups, make sure to set the argument run=FALSE
stage2_g1 <- tssem2(stage1_g1.fit, Amatrix=RAM1$A, Smatrix=RAM1$S, run=FALSE, model.name="g1")

stage2_g2 <- tssem2(stage1_g2.fit, Amatrix=RAM1$A, Smatrix=RAM2$S, run=FALSE, model.name="g2")

# Create the multigroup model
stage2_constrained <- mxModel(model="same_regression_coef", stage2_g1, stage2_g2,
                              mxFitFunctionMultigroup(c("g1", "g2")))

# Fit multigroup model with equality constraints
Stage2_constrained.fit <- mxRun(stage2_constrained, intervals=TRUE)

# first make a list of the fitted models in the separate groups
submodels.fit <- list(stage2_g1.fit,stage2_g2.fit)

subgroup.summary(submodels.fit,Stage2_constrained.fit)
```

# OSMASEM (with complete data)

## Data preparation
```{r}
## Get the data
data <- Nohe15A1$data
n <- Nohe15A1$n
Lag <- Nohe15A1$Lag
  
## Calculate the sampling covariance matrix of the correlations
my.df <- Cor2DataFrame(data, n, acov = "weighted")

## Add standardized Lag as a moderator.
## Standardization of the moderator improves the convergence.
my.df$data <- data.frame(my.df$data, Lag=scale(Nohe15A1$Lag),
                         check.names=FALSE)
head(my.df$data)

## Check the number of studies
pattern.na(my.df, show.na=FALSE, type="osmasem")

## Proposed model
model1 <- 'W2 ~ w2w*W1 + s2w*S1
           S2 ~ w2s*W1 + s2s*S1
           W1 ~~ w1WITHs1*S1
           W2 ~~ w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ Errw2*W2
           S2 ~~ Errs2*S2'

plot(model1, col="yellow")     

## Convert the lavaan syntax into the RAM specification
RAM1 <- lavaan2RAM(model1, obs.variables=c("W1", "S1", "W2", "S2"))
RAM1
```

## Model without any moderator
```{r}
## Create the model implied correlation structure with implicit diagonal constraints
## M0 <- create.vechsR(A0=RAM1$A, S0=RAM1$S)

## Create the heterogeneity variance-covariance matrix
## RE.type= either "Diag" or "Symm"
## Transform= either "expLog" or "sqSD" for better estimation on variances
## T0 <- create.Tau2(RAM=RAM1, RE.type="Diag", Transform="expLog", RE.startvalues=0.05)

## mx.fit0 <- osmasem(model.name="No moderator", Mmatrix=M0, Tmatrix=T0, data=my.df)

mx.fit0 <- osmasem(model.name="No moderator", RAM=RAM1, data=my.df)
summary(mx.fit0)

## The variance-covariance matrix in mx.fit0 is based on the untransformed matrix
## Extract the heterogeneity variance-covariance matrix
VarCorr(mx.fit0)
```

## Model with `Lag` as a moderator on the A matrix
```{r, message=FALSE}
Ax <- matrix(c(0,0,0,0,
               0,0,0,0,
               "0*data.Lag","0*data.Lag",0,0,
               "0*data.Lag","0*data.Lag",0,0),
             nrow=4, ncol=4, byrow=TRUE)
Ax              

## When there are more than one moderators
## Ax <- list(A1, A2, A3)

## Create the model implied correlation structure with the standardized Lag as the moderator
## M1 <- create.vechsR(A0=RAM1$A, S0=RAM1$S, Ax=Ax)

## mx.fit1 <- osmasem(model.name="Ax as moderator", Mmatrix=M1, Tmatrix=T0, data=my.df)

mx.fit1 <- osmasem(model.name="Ax as moderator", RAM=RAM1, Ax=Ax, data=my.df)
summary(mx.fit1)

## Extract the residual heterogeneity variance-covariance matrix
VarCorr(mx.fit1)

## Calculate the R2
## Tau2.0: Heterogeneity variances without the predictors
## Tau2.1: Heterogeneity variances with the predictors
## R2: (Tau2.0-Tau2.1)/Tau2.0
osmasemR2(mx.fit1, mx.fit0)

## Compare the models with and without the moderator
anova(mx.fit1, mx.fit0)

## Get the estimated A0 and A1
A0 <- mxEval(A0, mx.fit1$mx.fit)
A0
A1 <- mxEval(A1, mx.fit1$mx.fit)
A1

## Compute the estimated A matrix at -1SD (-1) of the standardized Lag
A0 - A1

## Compute the estimated A matrix at 0 (mean) of the standardized Lag
A0

## Compute the estimated A matrix at +1SD (+1) of the standardized Lag
A0 + A1
```

# TSSEM (with 1/4 variables (3/6 correlations) per study randomly deleted)

## A model without any moderator
```{r}
## Set seed for reproducibility
set.seed(345678)

## A function to create missing data: 1/4 of the variables were randomly deleted.
del_rand <- function(x, pattern=c(TRUE, TRUE, TRUE, FALSE)) {
  filter <- sample(pattern)
  x[!filter, ] <- NA
  x[, !filter] <- NA
  diag(x) <- 1
  x
}

data.missing <- lapply(Nohe15A1$data, del_rand)

## Display the number of data points
pattern.na(data.missing, show.na=FALSE)

## Proposed model in lavaan syntax
model1 <- 'W2 ~ w2w*W1 + s2w*S1
           S2 ~ w2s*W1 + s2s*S1
           W1 ~~ w1WITHs1*S1
           W2 ~~ w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ Errw2*W2
           S2 ~~ Errs2*S2'
     
RAM1 <- lavaan2RAM(model1, obs.variables=c("W1", "S1", "W2", "S2"))
RAM1

## Stage 1 analysis
random1 <- tssem1(data.missing, Nohe15A1$n, method="REM", RE.type="Diag")
summary(random1)

## Stage 2 analysis
random2 <- tssem2(random1, Amatrix=RAM1$A, Smatrix=RAM1$S)
summary(random2)

## Plot the model
plot(random2, col="green")
```

## Models with three subgroup analysis
```{r}
## Get the necessary functions
## source("http://www.suzannejak.nl/subgroup.functions.R")

data <- data.missing
n <- Nohe15A1$n
Lag <- Nohe15A1$Lag

# Data for studies with short Lag 
data_g1 <- data[Lag<7]
n_g1 <- n[Lag<7]

# Data for studies with medium Lag 
data_g2 <- data[Lag>=7&Lag<13]
n_g2 <- n[Lag>=7&Lag<13]

# Data for studies with long Lag 
data_g3 <- data[Lag>=13]
n_g3 <- n[Lag>=13]
```

### Fitting a fix-effects Stage 1 model in three subgroups as there is not enough data
```{r, message=FALSE}
## Stage 1 analysis per subgroup (random-effects analysis)
stage1_g1.fit <- tssem1(Cov = data_g1, n = n_g1, method = "REM", RE.type = "Zero")
stage1_g2.fit <- tssem1(Cov = data_g2, n = n_g2, method = "REM", RE.type = "Zero")
stage1_g3.fit <- tssem1(Cov = data_g3, n = n_g3, method = "REM", RE.type = "Zero")

## Results
summary(stage1_g1.fit)
summary(stage1_g2.fit)
summary(stage1_g3.fit)
```

### Fitting the Stage 2 model in three subgroups
```{r}
## Stage 2 analysis per subgroup (random-effect analysis)
stage2_g1.fit <- tssem2(stage1_g1.fit, Amatrix=RAM1$A, Smatrix=RAM1$S)
stage2_g2.fit <- tssem2(stage1_g2.fit, Amatrix=RAM1$A, Smatrix=RAM1$S)
stage2_g3.fit <- tssem2(stage1_g3.fit, Amatrix=RAM1$A, Smatrix=RAM1$S)

## Results
summary(stage2_g1.fit)
summary(stage2_g2.fit)
summary(stage2_g3.fit)
```

### Testing the equality of regression coefficients
* We create and fit a model with equal direct effects (we use the same matrix A for both groups), but different variances and covariances, so we create an S matrix with different labels for group 2 and group 3.
```{r}
## Proposed model g2
model2 <- 'W2 ~ w2w*W1 + s2w*S1
           S2 ~ w2s*W1 + s2s*S1
           W1 ~~ g2w1WITHs1*S1
           W2 ~~ g2w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ g2Errw2*W2
           S2 ~~ g2Errs2*S2'
     
RAM2 <- lavaan2RAM(model2, obs.variables=c("W1", "S1", "W2", "S2"))

## Proposed model g3
model3 <- 'W2 ~ w2w*W1 + s2w*S1
           S2 ~ w2s*W1 + s2s*S1
           W1 ~~ g3w1WITHs1*S1
           W2 ~~ g3w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ g3Errw2*W2
           S2 ~~ g3Errs2*S2'
     
RAM3 <- lavaan2RAM(model3, obs.variables=c("W1", "S1", "W2", "S2"))

## Create the models for the two groups, make sure to set the argument run=FALSE
stage2_g1 <- tssem2(stage1_g1.fit, Amatrix=RAM1$A, Smatrix=RAM1$S, run=FALSE, model.name="g1")

stage2_g2 <- tssem2(stage1_g2.fit, Amatrix=RAM1$A, Smatrix=RAM2$S, run=FALSE, model.name="g2")

stage2_g3 <- tssem2(stage1_g3.fit, Amatrix=RAM1$A, Smatrix=RAM3$S, run=FALSE, model.name="g3")

## Create the multigroup model
stage2_constrained <- mxModel(model="same_regression_coef", stage2_g1, stage2_g2,stage2_g3,
                              mxFitFunctionMultigroup(c("g1", "g2", "g3")))

## Fit multigroup model with equality constraints
Stage2_constrained.fit <- mxRun(stage2_constrained, intervals=TRUE)

## first make a list of the fitted models in the separate groups
submodels.fit <- list(stage2_g1.fit, stage2_g2.fit, stage2_g3.fit)

subgroup.summary(submodels.fit,Stage2_constrained.fit)
```

### Testing the equality of one regression coefficient (w2w)
* We create and fit a model with equal direct effects (we use the same matrix A for both groups), but different variances and covariances, so we create an S matrix with different labels for group 2 and group 3.

```{r}
## Proposed model g2
model2 <- 'W2 ~ w2w*W1 + g2s2w*S1
           S2 ~ g2w2s*W1 + g2s2s*S1
           W1 ~~ g2w1WITHs1*S1
           W2 ~~ g2w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ g2Errw2*W2
           S2 ~~ g2Errs2*S2'
     
RAM2 <- lavaan2RAM(model2, obs.variables=c("W1", "S1", "W2", "S2"))

## Proposed model g3
model3 <- 'W2 ~ w2w*W1 + g3s2w*S1
           S2 ~ g3w2s*W1 + g3s2s*S1
           W1 ~~ g3w1WITHs1*S1
           W2 ~~ g3w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ g3Errw2*W2
           S2 ~~ g3Errs2*S2'
     
RAM3 <- lavaan2RAM(model3, obs.variables=c("W1", "S1", "W2", "S2"))

## Create the models for the two groups, make sure to set the argument run=FALSE
stage2_g1 <- tssem2(stage1_g1.fit, Amatrix=RAM1$A, Smatrix=RAM1$S, run=FALSE, model.name="g1")

stage2_g2 <- tssem2(stage1_g2.fit, Amatrix=RAM2$A, Smatrix=RAM2$S, run=FALSE, model.name="g2")

stage2_g3 <- tssem2(stage1_g3.fit, Amatrix=RAM3$A, Smatrix=RAM3$S, run=FALSE, model.name="g3")

## Create the multigroup model
stage2_constrained <- mxModel(model="same_regression_coef", stage2_g1, stage2_g2,stage2_g3,
                              mxFitFunctionMultigroup(c("g1", "g2", "g3")))

## Fit multigroup model with equality constraints
Stage2_constrained.fit <- mxRun(stage2_constrained, intervals=TRUE)

## First make a list of the fitted models in the separate groups
submodels.fit <- list(stage2_g1.fit,stage2_g2.fit,stage2_g3.fit)

subgroup.summary(submodels.fit,Stage2_constrained.fit)
```

## Models with two subgroup analysis
```{r}
# Data for studies with short Lag 
data_g1 <- data[Lag<12]
n_g1 <- n[Lag<12]

# Data for studies with long Lag 
data_g2 <- data[Lag>=12]
n_g2 <- n[Lag>=12]
```

### Fitting a fixed-effects Stage 1 model in two subgroups as there is not enough data
```{r, message=FALSE}
## Stage 1 analysis per subgroup (random-effects analysis)
stage1_g1.fit <- tssem1(Cov = data_g1, n = n_g1, method = "REM", RE.type = "Zero")
stage1_g2.fit <- tssem1(Cov = data_g2, n = n_g2, method = "REM", RE.type = "Zero")

summary(stage1_g1.fit)
summary(stage1_g2.fit)
```

### Fitting the Stage 2 model in both subgroups
```{r}
## Stage 2 analysis per subgroup (random-effect analysis)
stage2_g1.fit <- tssem2(stage1_g1.fit, Amatrix=RAM1$A, Smatrix=RAM1$S)
stage2_g2.fit <- tssem2(stage1_g2.fit, Amatrix=RAM1$A, Smatrix=RAM1$S)

summary(stage2_g1.fit)
summary(stage2_g2.fit)
```

### Testing the equality of regression coefficients
* We create and fit a model with equal direct effects (we use the same matrix A for both groups), but different variances and covariances, so we create an S matrix with different labels for group 2.

```{r}
## Proposed model g2
model2 <- 'W2 ~ w2w*W1 + s2w*S1
           S2 ~ w2s*W1 + s2s*S1
           W1 ~~ g2w1WITHs1*S1
           W2 ~~ g2w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ g2Errw2*W2
           S2 ~~ g2Errs2*S2'
     
RAM2 <- lavaan2RAM(model2, obs.variables=c("W1", "S1", "W2", "S2"))

# Create the models for the two groups, make sure to set the argument run=FALSE
stage2_g1 <- tssem2(stage1_g1.fit, Amatrix=RAM1$A, Smatrix=RAM1$S, run=FALSE, model.name="g1")

stage2_g2 <- tssem2(stage1_g2.fit, Amatrix=RAM1$A, Smatrix=RAM2$S, run=FALSE, model.name="g2")

# Create the multigroup model
stage2_constrained <- mxModel(model="same_regression_coef", stage2_g1, stage2_g2,
                              mxFitFunctionMultigroup(c("g1", "g2")))

# Fit multigroup model with equality constraints
Stage2_constrained.fit <- mxRun(stage2_constrained, intervals=TRUE)

# first make a list of the fitted models in the separate groups
submodels.fit <- list(stage2_g1.fit,stage2_g2.fit)

subgroup.summary(submodels.fit,Stage2_constrained.fit)
```

# OSMASEM (with 1/4 variables (3/6 correlations) per study randomly deleted)

## Data preparation
```{r}
## Get the data
n <- Nohe15A1$n
Lag <- Nohe15A1$Lag
  
## Calculate the sampling covariance matrix of the correlations
my.df <- Cor2DataFrame(data.missing, n, acov = "weighted")

## Add standardized Lag as a moderator.
## Standardization of the moderator improves the convergence.
my.df$data <- data.frame(my.df$data, Lag=scale(Nohe15A1$Lag),
                         check.names=FALSE)
head(my.df$data)

## Check the number of studies
pattern.na(my.df, show.na=FALSE, type="osmasem")

## Proposed model
model1 <- 'W2 ~ w2w*W1 + s2w*S1
           S2 ~ w2s*W1 + s2s*S1
           W1 ~~ w1WITHs1*S1
           W2 ~~ w2WITHs2*S2
           W1 ~~ 1*W1
           S1 ~~ 1*S1
           W2 ~~ Errw2*W2
           S2 ~~ Errs2*S2'

plot(model1, col="yellow")     

## Convert the lavaan syntax into the RAM specification
RAM1 <- lavaan2RAM(model1, obs.variables=c("W1", "S1", "W2", "S2"))
RAM1
```

## Model without any moderator
```{r}
## Create the model implied correlation structure with implicit diagonal constraints
## M0 <- create.vechsR(A0=RAM1$A, S0=RAM1$S)

## Create the heterogeneity variance-covariance matrix
## RE.type= either "Diag" or "Symm"
## Transform= either "expLog" or "sqSD" for better estimation on variances
## T0 <- create.Tau2(RAM=RAM1, RE.type="Diag", Transform="expLog", RE.startvalues=0.05)

## mx.fit0 <- osmasem(model.name="No moderator", Mmatrix=M0, Tmatrix=T0, data=my.df)

mx.fit0 <- osmasem(model.name="No moderator", RAM=RAM1, data=my.df)
summary(mx.fit0)

## The variance-covariance matrix in mx.fit0 is based on the untransformed matrix
## Extract the heterogeneity variance-covariance matrix
VarCorr(mx.fit0)
```

## Model with `Lag` as a moderator on the A matrix
```{r, message=FALSE}
Ax <- matrix(c(0,0,0,0,
               0,0,0,0,
               "0*data.Lag","0*data.Lag",0,0,
               "0*data.Lag","0*data.Lag",0,0),
             nrow=4, ncol=4, byrow=TRUE)
Ax              

## When there are more than one moderators
## Ax <- list(A1, A2, A3)

## Create the model implied correlation structure with the standardized Lag as the moderator
## M1 <- create.vechsR(A0=RAM1$A, S0=RAM1$S, Ax=Ax)

## mx.fit1 <- osmasem(model.name="Ax as moderator", Mmatrix=M1, Tmatrix=T0, data=my.df)

mx.fit1 <- osmasem(model.name="Ax as moderator", RAM=RAM1, Ax=Ax, data=my.df)
summary(mx.fit1)

## Extract the residual heterogeneity variance-covariance matrix
VarCorr(mx.fit1)

## Calculate the R2
## Tau2.0: Heterogeneity variances without the predictors
## Tau2.1: Heterogeneity variances with the predictors
## R2: (Tau2.0-Tau2.1)/Tau2.0
osmasemR2(mx.fit1, mx.fit0)

## Compare the models with and without the moderator
anova(mx.fit1, mx.fit0)

## Get the estimated A0 and A1
A0 <- mxEval(A0, mx.fit1$mx.fit)
A0
A1 <- mxEval(A1, mx.fit1$mx.fit)
A1

## Compute the estimated A matrix at -1SD (-1) of the standardized Lag
A0 - A1

## Compute the estimated A matrix at 0 (mean) of the standardized Lag
A0

## Compute the estimated A matrix at +1SD (+1) of the standardized Lag
A0 + A1
```

```{r}
sessionInfo()
```