---
title: 'Synthesizing Indirect Effects in Mediation Models with Meta-Analytic Methods: Supplementary Materials 1'
author: "Mike Cheung"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
---

* This file demonstrates how to compute effect sizes and their sampling covariance matrix with two approaches using the delta method. The first one uses a numeric approach with the structural equation modeling (SEM) framework. The second approach computes the sampling covariance matrix with the symbolic calculations.

# Numeric calculations with the SEM approach

## One mediator
```{r, message=FALSE}
library(metaSEM)

## Model with one mediator: x -> m -> y
model1 <- "y ~ c*x + b*m
           m ~ a*x
           # Define indirect and direct effects
           Indirect := a*b
           Direct := c"

## Diplay the model
plot(model1)

## Sample correlation matrix
my.cor <- matrix(c(1, .5, .3,
                   .5, 1, .4,
                   .3, .4, 1),
                 nrow = 3,
                 ncol = 3,
                 dimnames = list(c("y", "m", "x"),
                                 c("y", "m", "x")))
my.cor

## Calculate the indirect and direct effects and their sampling covariance matrix
calEffSizes(model=model1, n=300, Cov=my.cor)
```

## Two parallel mediators
```{r}
## Model with two specific mediators: x -> m1 -> y and x -> m2 -> y
model2 <- "y ~ e*x + b*m1 + d*m2
           m1 ~ a*x
           m2 ~ c*x
           # m1 and m2 are correlated
           m1 ~~ m2
           # Define indirect and direct effects
           Ind_m1 := a*b
           Ind_m2 := c*d
           Direct := e"

plot(model2, layout="circle")

## Sample correlation matrix
my.cor <- matrix(c(1, .5, .6, .3, 
                   .5, 1, .4, .2,
                   .6, .4, 1, .3,
                   .3, .2, .3, 1),
                 nrow = 4,
                 ncol = 4,
                 dimnames = list(c("y", "m1", "m2", "x"),
                                 c("y", "m1", "m2", "x")))
my.cor

## Calculate the indirect and direct effects and their sampling covariance matrix
calEffSizes(model=model2, n=300, Cov=my.cor)
```

## Two serial mediators
```{r}
## Model with two intermediate mediators: x -> m1 -> m2 -> y
model3 <- "y ~ e*x + f*m1 + c*m2
           m1 ~ a*x
           m2 ~ b*m1
           m2 ~ d*x
           # Define indirect and direct effects
           Ind_m1m2 := a*b*c
           Ind_m1 := a*f
           Ind_m2 := d*c
           Direct := e"

plot(model3, layout="circle")

## Calculate the indirect and direct effects and their sampling covariance matrix
calEffSizes(model=model3, n=300, Cov=my.cor)
```

# Symbolic calculations

## One mediator
```{r}
library(symSEM)

## fn: The effect sizes
## Covfn: Sampling covariance matrix of fn: "b^2*Va+2*b*a*Cba+a^2*Vb"
## Va: Sampling variance of a
## Vb: Sampling variance of b
## Cba: Sampling covariance of a and b
deltamethod(fn="a*b")
```

## Two parallel mediators
```{r}
deltamethod(fn=c("a*b", "c*d"))
```

## Two serial mediators
```{r}
deltamethod(fn="a*b*c")

sessionInfo()
```




