---
title: 'Synthesizing Indirect Effects in Mediation Models with Meta-Analytic Methods: Supplementary Materials 2'
author: "Mike Cheung"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
---

# Data preparation
```{r, message=FALSE}
library(metaSEM)

## Get the data from the local file
source("Hagger18.R")

## Check whether the correlation matrices are positive definite
which(is.pd(Hagger18$data)==FALSE)

## Drop the 21th study as the correlation matrix is not positive definite.
Hagger18 <- lapply(Hagger18, function(x) x[-21])
```

# Illustration 1 with one mediator
```{r}
## Use df1 as the data file in illustration 1
df1 <- Hagger18

## Select Beh, Int, and PB for the illustration
obs.vars1 <- c("Beh", "Int", "PB")
df1$data <- lapply(df1$data, function(x) x[obs.vars1, obs.vars1])

## NA is not allowed in computing the indirect and direct effects.
## Studies with NA are excluded first.
index1 <- sapply(df1$data, function(x) any(is.na(vechs(x))) )
df1 <- lapply(df1, function(x) x[!index1])

## Show the first few studies
head(df1)

## Show the no. of studies per correlation
pattern.na(df1$data, show.na = FALSE)

## Show the total sample sizes per correlation
pattern.n(df1$data, df1$n)
```

## Meta-analyzing the indirect and direct effects
### Caculation of indirect and direct effects
```{r}
## Calculate the indirect and direct effects and their sampling covariance matrices
## The variables are arranged as outcome, mediator, and predictor.
IE.df1 <- indirectEffect(df1$data, df1$n)

## Add behavior frequency to the data
IE.df1 <- data.frame(IE.df1, beh_freq_high=df1$beh_freq_high)

## Show the first few studies
head(IE.df1)
```

### Meta-analysis of indirect and direct effects
```{r, message=FALSE}
## Random-effects model
IE0 <- meta(y=cbind(ind_eff, dir_eff),
            v=cbind(ind_var, ind_dir_cov, dir_var),
            data=IE.df1,
            model.name = "Random")
summary(IE0)

## Variance-covariance matrix of the random effects
VarCorr(IE0)

## Correlation matrix of the random effects
cov2cor(VarCorr(IE0))

## Plot the effect sizes
plot(IE0, axis.labels = c("Indirect effect", "Direct effect"))

## Mixed-effects model with behavior frequency as a moderator
IE1 <- meta(y=cbind(ind_eff, dir_eff),
            v=cbind(ind_var, ind_dir_cov, dir_var),
            x=beh_freq_high,
            data=IE.df1,
            model.name = "Mixed")
summary(IE1)

## Test the statistical significance between the models
anova(IE1, IE0)
```

```{r, eval=FALSE, echo=FALSE}
svg("Figure4.svg", width=7, height=5)
plot(IE0, axis.labels = c("Indirect effect", "Direct effect"), add.margin = 0.05)
dev.off()
```

## TSSEM
### Stage 1 analysis
```{r, message=FALSE}
## Random-effects model
random1 <- tssem1(df1$data, df1$n, method="REM")
summary(random1)

## Average correlation matrix under a random-effects model
averageR <- vec2symMat(coef(random1, select="fixed"), diag = FALSE)
dimnames(averageR) <- list(obs.vars1, obs.vars1)
averageR

## Heterogeneity variances of the random-effects
coef(random1, select="random")
```

### Stage 2 analysis
```{r, message=FALSE}
## Proposed model in lavaan syntax
model1 <- "Beh ~ c*PB + b*Int
           Int ~ a*PB
           PB ~~ 1*PB"
plot(model1)

## Convert the lavaan syntax to RAM specification used in metaSEM
RAM1 <- lavaan2RAM(model1, obs.variables=obs.vars1)
RAM1

## Request the likelihood-based confidence interval
## Indirect effect: ind = a*b
## Direct effect: dir = c
tssem.fit <- tssem2(random1, RAM=RAM1, intervals.type = "LB",
                    mx.algebras = list(ind=mxAlgebra(a*b, name="ind"),
                                       dir=mxAlgebra(c, name="dir")))
summary(tssem.fit)
plot(tssem.fit, color="green")
```

```{r, eval=FALSE, echo=FALSE}
svg("Figure4.svg", width=5, height=5)
plot(tssem.fit, color="gray")
dev.off()
```

## OSMASEM
### Model without any moderator
```{r, message=FALSE}
## Convert the data format from TSSEM to OSMASEM
osmasem.df1 <- Cor2DataFrame(df1)

## Show the first few studies
head(osmasem.df1)

## Fit a model without any moderator
osmasem.fit0 <- osmasem(model.name="No moderator", RAM=RAM1, data=osmasem.df1)
summary(osmasem.fit0)

## Get the heterogeneity of variances
VarCorr(osmasem.fit0)

## Plot the fitted model
plot(osmasem.fit0, color="yellow")
```

### Model with a moderator
```{r}
## Create A1 to represent the moderator on the A matrix
A1 <- create.modMatrix(RAM1, output="A", "beh_freq_high")
A1

## Fit a model with female as a moderator
osmasem.fit1 <- osmasem(model.name="Female as a moderator", RAM=RAM1, Ax=A1, data=osmasem.df1)
summary(osmasem.fit1)

## Test the statistical significance between the models
anova(osmasem.fit1, osmasem.fit0)

## Get the R2 on the correlation coefficients
osmasemR2(osmasem.fit0, osmasem.fit1)
```


# Illustration 2 with two parallel mediators (Aut and Cap)
```{r}
## Use new.df2 as the data file in illustration 2
df2 <- Hagger18

## Select Aut, Cap, Beh, and PB for the illustration
obs.vars2 <- c("Aut", "Cap", "Beh", "PB")
df2$data <- lapply(df2$data, function(x) x[obs.vars2, obs.vars2])

## Drop studies do not include all correlation in c("Aut", "Cap", "Beh", "PB")
index2 <- sapply(df2$data, function(x) any(is.na(vechs(x))))
df2 <- lapply(df2, function(x) x[!index2])

## Show the first few studies
head(df2)

## Show the no. of studies per correlation
pattern.na(df2$data, show.na = FALSE)

## Show the total sample sizes per correlation
pattern.n(df2$data, df2$n)
```

## Meta-analyzing the indirect and direct effects
### Caculation of indirect and direct effects
```{r}
## Calculate indirect and direct effects as effect sizes in each study
## PB -> Cap -> Beh
## PB -> Aut -> Beh
model2 <- "Cap ~ a*PB
           Aut ~ c*PB
           Beh ~ b*Cap + d*Aut + e*PB
           Cap ~~ f*Aut
           PB ~~ 1*PB
           ## Define indirect and direct effects
           Ind_Cap := a*b
           Ind_Aut := c*d
           Dir_PB := f"

## Display the proposed model
plot(model2, layout="circle")

## Estimate the indirect and direct effects
IE.df2 <- mapply(function(x, y) {calEffSizes(model=model2, n=y, Cov=x)},
                                 df2$data, df2$n, SIMPLIFY = FALSE)

head(IE.df2)
  
## Rename the variances and covariances of the effect sizes from Cov1 to Cov6 for ease of reference
IE.df2 <- t(sapply(IE.df2, 
                   function(x) { acov <- vech(x$VCOV)
                                 names(acov) <- paste0("Cov", 1:6)
                                 c(x$ES, acov)} ))

## Show the first few studies
head(IE.df2)
```

### Meta-analysis of indirect and direct effects
```{r, message=FALSE}
## Random-effects model with independent random effects
## Tau2_2_2 is close to negative. It is fixed at 0.
IE2 <- meta(y=IE.df2[, c("Ind_Cap", "Ind_Aut", "Dir_PB")],
            v=IE.df2[, paste0("Cov", 1:6)],
            RE.constraints = Diag(c("0.01*Tau2_1_1", "0", "0.01*Tau2_3_3")))
summary(IE2)
```

## TSSEM
### Stage 1 analysis
```{r}
## Random-effects model
random1 <- tssem1(df2$data, df2$n)
summary(random1)

## Average correlation matrix under a random-effects model
averageR <- vec2symMat(coef(random1, select="fixed"), diag = FALSE)
dimnames(averageR) <- list(obs.vars2, obs.vars2)
averageR

## Heterogeneity variances of the random-effects
coef(random1, select="random")
```

### Stage 2 analysis
```{r}
RAM2 <- lavaan2RAM(model2, obs.variables=obs.vars2)
RAM2

tssem.fit <- tssem2(random1, RAM=RAM2, intervals.type = "LB",
                    mx.algebras = list(Ind_Cap=mxAlgebra(a*b, name="Ind_Cap"),
                                       Ind_Aut=mxAlgebra(c*d, name="Ind_Aut"),
                                       Dir_PB=mxAlgebra(f, name="Dir_PB")))
summary(tssem.fit)
plot(tssem.fit, layout="circle", color="green")
```

```{r, eval=FALSE, echo=FALSE}
svg("Figure5.svg", width=5, height=5)
plot(tssem.fit, layout="circle", color="gray")
dev.off()
```

# Illustration 3 with two serial mediators (Aut and Cap)
```{r}
## Use df3 as the data file in illustration 3
df3 <- Hagger18

## Select Cap, Int, Beh, and PB for the illustration
obs.vars3 <- c("Cap", "Int", "Beh", "PB")
df3$data <- lapply(df3$data, function(x) x[obs.vars3, obs.vars3])

## Drop studies do not include all correlations in c("Cap", "Int", "Beh", "PB")
index3 <- sapply(df3$data, function(x) any(is.na(vechs(x))))
df3 <- lapply(df3, function(x) x[!index3])

## Show the first few studies
head(df3)

## Show the no. of studies per correlation
pattern.na(df3$data, show.na = FALSE)

## Show the total sample sizes per correlation
pattern.n(df3$data, df3$n)
```

## Meta-analyzing the indirect and direct effects
### Caculation of indirect and direct effects
```{r}
## Calculate indirect and direct effects as effect sizes in each study
## PB -> Cap -> Int -> Beh
model3 <- "Cap ~ a*PB
           Int ~ b*Cap + d*PB
           Beh ~ e*Cap + c*Int + f*PB
           PB ~~ 1*PB
           ## Define indirect and direct effects
           Ind_CapInt := a*b*c
           Ind_Cap := a*e
           Ind_Int := d*c
           Dir_PB := f"

## Display the proposed model
plot(model3, layout="circle")

## Estimate the indirect and direct effects
IE.df3 <- mapply(function(x, y) {calEffSizes(model=model3, n=y, Cov=x)},
                                 df3$data, df3$n, SIMPLIFY = FALSE)
  
## Rename the variances and covariances of the effect sizes from Cov1 to Cov10 for ease of reference
IE.df3 <- t(sapply(IE.df3, 
                   function(x) { acov <- vech(x$VCOV)
                                 names(acov) <- paste0("Cov", 1:10)
                                 c(x$ES, acov)} ))

## Show the first few studies
head(IE.df3)
```

### Meta-analysis of indirect and direct effects
```{r}
## Random-effects model with independent random effects
IE3 <- meta(y=IE.df3[, c("Ind_CapInt", "Ind_Cap", "Ind_Int", "Dir_PB")],
            v=IE.df3[, paste0("Cov", 1:10)],
            RE.constraints = Diag(paste0("0.01*Tau2_", 1:4, "_", 1:4)))
summary(IE3)
```

## TSSEM
### Stage 1 analysis
```{r}
## Random-effects model
random1 <- tssem1(df3$data, df3$n)
summary(random1)

## Average correlation matrix under a random-effects model
averageR <- vec2symMat(coef(random1, select="fixed"), diag = FALSE)
dimnames(averageR) <- list(obs.vars3, obs.vars3)
averageR

## Heterogeneity variances of the random-effects
coef(random1, select="random")
```

### Stage 2 analysis
```{r}
RAM3 <- lavaan2RAM(model3, obs.variables=obs.vars3)
RAM3

tssem.fit <- tssem2(random1, RAM=RAM3, intervals.type = "LB",
                    mx.algebras = list(Ind_CapInt=mxAlgebra(a*b*c, name="Ind_CapInt"),
                                       Ind_Cap=mxAlgebra(a*e, name="Ind_Cap"),
                                       Ind_Int=mxAlgebra(d*c, name="Ind_Int"),
                                       Dir_PB=mxAlgebra(f, name="Dir_PB")))
summary(tssem.fit)
plot(tssem.fit, layout="circle", color="green")
```

```{r, eval=FALSE, echo=FALSE}
svg("Figure6.svg", width=5, height=5)
plot(tssem.fit, layout="circle", color="gray")
dev.off()
```

```{r}
sessionInfo()
```

