---
title: 'Supplementary Materials for "A Guide to Conducting a Meta-Analysis with Non-Independent Effect Sizes"'
author: Mike W.-L. Cheung
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
  word_document: default
editor_options: 
  chunk_output_type: console
---

* Latest version of the R and Mplus code is available at https://github.com/mikewlcheung/code-in-articles.

# Multivariate meta-analysis with a dataset from Nam, Mengersen, and Garthwaite (2003)

## Preparation
```{r, message=FALSE}
## Required packages in this paper
lib2install <- c("metaSEM", "metafor")

## Install them automatically if they are not available in your computer
for (i in lib2install) {
  if (!(i %in% rownames(installed.packages()))) install.packages(i)
}

## Load the libraries
library(metaSEM)
library(metafor)

## Display the first few lines of the data
head(Nam03)
```

## Univariate meta-analyses
```{r}
## Univariate meta-analysis on Asthma
summary(meta(y=Asthma_logOR, v=Asthma_v, data=Nam03))

## Univariate meta-analysis on LRD
summary(meta(y=LRD_logOR, v=LRD_v, data=Nam03))
```

## Multivariate meta-analysis (assuming cor=0.5 between asthma and LRD)
### Without any moderator
```{r, warning=FALSE}
fit0a <- meta(y=cbind(Asthma_logOR, LRD_logOR), 
              v=cbind(Asthma_v, AsthmaLRD_cov_05, LRD_v),
              data=Nam03,
              model.name="Multi MA")
summary(fit0a)

## Correlation between the population effect sizes
cov2cor(VarCorr(fit0a))

## Plot the (complete) effect sizes and their 95% confidence ellipses
plot(fit0a, xlim=c(-0.5, 1), ylim=c(-0.5, 1), axis.labels=c("Asthma", "LRD"),
     study.ellipse.plot = FALSE, diag.panel=TRUE,
     randeff.ellipse.lty=2, 
     main="Log-odds ratio")

forest(rma(yi=Asthma_logOR, vi=Asthma_v, slab=ID, data=Nam03),
       xlab="Log-odds ratio")
title("Forest plot of Asthma")

forest(rma(yi=LRD_logOR, vi=LRD_v, slab=ID, data=Nam03),
       xlab="Log-odds ratio")
title("Forest plot of LRD")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
pdf("Fig1.pdf", width = 10, height = 10)
## Plot the effect sizes (complete studies only)
plot(fit0a, xlim=c(-0.5, 1), ylim=c(-0.5, 1), axis.labels=c("Asthma", "LRD"),
     study.ellipse.plot = FALSE, diag.panel=TRUE,
     randeff.ellipse.lty=2, 
     main="Log-odds ratio (assuming Cor=0.5)")

forest(rma(yi=Asthma_logOR, vi=Asthma_v, slab=ID, data=Nam03),
       xlab="Log-odds ratio")
title("Forest plot of Asthma")

forest(rma(yi=LRD_logOR, vi=LRD_v, slab=ID, data=Nam03),
       xlab="Log-odds ratio")
title("Forest plot of LRD")
dev.off()
```

```{r}
## Test whether (1) the population effect sizes and 
## (2) heterogeneity variances are the same for Asthma and LRD
fit1a <- meta(y=cbind(Asthma_logOR, LRD_logOR), 
              v=cbind(Asthma_v, AsthmaLRD_cov_05, LRD_v),
              data=Nam03,
              intercept.constraints=c("0*Intercept", "0*Intercept"),
              RE.constraints=matrix(c("0.1*Tau2_1_1", "0*Tau2_2_1",
                                      "0*Tau2_2_1", "0.1*Tau2_1_1"),
                                    ncol=2, nrow=2),
              model.name="Equality constraints")
## Compare the models with and without the equality constraints
anova(fit0a, fit1a)

summary(fit1a)
```

### With the mean `Age` as a moderator
```{r}
fit2a <- meta(y=cbind(Asthma_logOR, LRD_logOR), 
              v=cbind(Asthma_v, AsthmaLRD_cov_05, LRD_v),
              ## Center Age for ease of interpretations
              x=scale(Age, scale=FALSE),
              ## Fix the covariance of between the random effects at 0
              ## as there is not enough data
              RE.constraints=matrix(c("0.1*Tau2_1_1", "0",
                                      "0", "0.1*Tau2_2_2"),
                                    ncol=2, nrow=2), 
              data=Nam03,
              model.name="Mixed MA")
summary(fit2a)

## Test whether the effect of mean Age the same for Asthma and LRD
fit3a <- meta(y=cbind(Asthma_logOR, LRD_logOR), 
              v=cbind(Asthma_v, AsthmaLRD_cov_05, LRD_v),
              ## Center Age for ease of interpretations
              x=scale(Age, scale=FALSE),
              ## Fix the covariance of between the random effects at 0
              ## as there is not enough data
              RE.constraints=matrix(c("0.1*Tau2_1_1", "0",
                                      "0", "0.1*Tau2_2_2"),
                                    ncol=2, nrow=2), 
              ## Fix the regression coefficients to be identical by
              ## using the same label `Slope1`.
              coef.constraints=matrix(matrix(c("0*Slope1", "0*Slope1"),
                                             nrow=1)),
              data=Nam03,
              model.name="Equality constraints")

## Compare the models with and without the equality constraints
anova(fit2a, fit3a)

summary(fit3a)
```

## Multivariate meta-analysis (assuming cor=0 between asthma and LRD)
### Without any moderator
```{r}
## Create a variable of sampling correlation=0
Nam03$AsthmaLRD_cov_0 <- 0

fit0b <- meta(y=cbind(Asthma_logOR, LRD_logOR), 
              v=cbind(Asthma_v, AsthmaLRD_cov_0, LRD_v),
              data=Nam03,
              model.name="Multi MA")
summary(fit0b)

## Correlation between the population effect sizes
cov2cor(VarCorr(fit0b))

## Plot the (complete) effect sizes and their 95% confidence ellipses
plot(fit0b, xlim=c(-0.5, 1), ylim=c(-0.5, 1), axis.labels=c("Asthma", "LRD"),
     study.ellipse.plot = FALSE, diag.panel=TRUE,
     randeff.ellipse.lty=2, 
     main="Log-odds ratio (assuming Cor=0)")

forest(rma(yi=Asthma_logOR, vi=Asthma_v, slab=ID, data=Nam03),
       xlab="Log-odds ratio")
title("Forest plot of Asthma")

forest(rma(yi=LRD_logOR, vi=LRD_v, slab=ID, data=Nam03),
       xlab="Log-odds ratio")
title("Forest plot of LRD")
```

```{r}
## Test whether (1) the population effect sizes and 
## (2) heterogeneity variances are the same for Asthma and LRD
fit1b <- meta(y=cbind(Asthma_logOR, LRD_logOR), 
              v=cbind(Asthma_v, AsthmaLRD_cov_0, LRD_v),
              data=Nam03,
              intercept.constraints=c("0*Intercept", "0*Intercept"),
              RE.constraints=matrix(c("0.1*Tau2_1_1", "0*Tau2_2_1",
                                      "0*Tau2_2_1", "0.1*Tau2_1_1"),
                                    ncol=2, nrow=2),
              model.name="Equality constraints")
## Compare the models with and without the equality constraints
anova(fit0b, fit1b)

summary(fit1b)
```

### With the mean `Age` as a moderator
```{r}
fit2b <- meta(y=cbind(Asthma_logOR, LRD_logOR), 
              v=cbind(Asthma_v, AsthmaLRD_cov_0, LRD_v),
              ## Center Age for ease of interpretations
              x=scale(Age, scale=FALSE),
              ## Fix the covariance of between the random effects at 0
              ## as there is not enough data
              RE.constraints=matrix(c("0.1*Tau2_1_1", "0",
                                      "0", "0.1*Tau2_2_2"),
                                    ncol=2, nrow=2), 
              data=Nam03,
              model.name="Mixed MA")
summary(fit2b)

## Test whether the effect of mean Age the same for Asthma and LRD
fit3b <- meta(y=cbind(Asthma_logOR, LRD_logOR), 
              v=cbind(Asthma_v, AsthmaLRD_cov_0, LRD_v),
              ## Center Age for ease of interpretations
              x=scale(Age, scale=FALSE),
              ## Fix the covariance of between the random effects at 0
              ## as there is not enough data
              RE.constraints=matrix(c("0.1*Tau2_1_1", "0",
                                      "0", "0.1*Tau2_2_2"),
                                    ncol=2, nrow=2), 
              ## Fix the regression coefficients to be identical by
              ## using the same label `Slope1`.
              coef.constraints=matrix(matrix(c("0*Slope1", "0*Slope1"),
                                             nrow=1)),
              data=Nam03,
              model.name="Equality constraints")

## Compare the models with and without the equality constraints
anova(fit2b, fit3b)

summary(fit3b)
```

## Multivariate meta-analysis (assuming cor=0.8 between asthma and LRD)
### Without any moderator
```{r}
## Create a variable of sampling correlation=0
Nam03$AsthmaLRD_cov_08 <- with(Nam03, 0.8*sqrt(Asthma_v)*sqrt(LRD_v))

fit0c <- meta(y=cbind(Asthma_logOR, LRD_logOR), 
              v=cbind(Asthma_v, AsthmaLRD_cov_08, LRD_v),
              data=Nam03,
              model.name="Multi MA")
summary(fit0c)

## Correlation between the population effect sizes
cov2cor(VarCorr(fit0c))

## Plot the (complete) effect sizes and their 95% confidence ellipses
plot(fit0c, xlim=c(-0.5, 1), ylim=c(-0.5, 1), axis.labels=c("Asthma", "LRD"),
     study.ellipse.plot = FALSE, diag.panel=TRUE,
     randeff.ellipse.lty=2, 
     main="Log-odds ratio (assuming Cor=0.8)")

forest(rma(yi=Asthma_logOR, vi=Asthma_v, slab=ID, data=Nam03),
       xlab="Log-odds ratio")
title("Forest plot of Asthma")

forest(rma(yi=LRD_logOR, vi=LRD_v, slab=ID, data=Nam03),
       xlab="Log-odds ratio")
title("Forest plot of LRD")
```

```{r}
## Test whether (1) the population effect sizes and 
## (2) heterogeneity variances are the same for Asthma and LRD
fit1c <- meta(y=cbind(Asthma_logOR, LRD_logOR), 
              v=cbind(Asthma_v, AsthmaLRD_cov_08, LRD_v),
              data=Nam03,
              intercept.constraints=c("0*Intercept", "0*Intercept"),
              RE.constraints=matrix(c("0.1*Tau2_1_1", "0*Tau2_2_1",
                                      "0*Tau2_2_1", "0.1*Tau2_1_1"),
                                    ncol=2, nrow=2),
              model.name="Equality constraints")
## Compare the models with and without the equality constraints
anova(fit0c, fit1c)

summary(fit1c)
```

### With the mean `Age` as a moderator
```{r}
fit2c <- meta(y=cbind(Asthma_logOR, LRD_logOR), 
              v=cbind(Asthma_v, AsthmaLRD_cov_08, LRD_v),
              ## Center Age for ease of interpretations
              x=scale(Age, scale=FALSE),
              ## Fix the covariance of between the random effects at 0
              ## as there is not enough data
              RE.constraints=matrix(c("0.1*Tau2_1_1", "0",
                                      "0", "0.1*Tau2_2_2"),
                                    ncol=2, nrow=2), 
              data=Nam03,
              model.name="Mixed MA")
summary(fit2c)

## Test whether the effect of mean Age the same for Asthma and LRD
fit3c <- meta(y=cbind(Asthma_logOR, LRD_logOR), 
              v=cbind(Asthma_v, AsthmaLRD_cov_08, LRD_v),
              ## Center Age for ease of interpretations
              x=scale(Age, scale=FALSE),
              ## Fix the covariance of between the random effects at 0
              ## as there is not enough data
              RE.constraints=matrix(c("0.1*Tau2_1_1", "0",
                                      "0", "0.1*Tau2_2_2"),
                                    ncol=2, nrow=2), 
              ## Fix the regression coefficients to be identical by
              ## using the same label `Slope1`.
              coef.constraints=matrix(matrix(c("0*Slope1", "0*Slope1"),
                                             nrow=1)),
              data=Nam03,
              model.name="Equality constraints")

## Compare the models with and without the equality constraints
anova(fit2c, fit3c)

summary(fit3c)
```

# Three-level meta-analysis with a dataset from Stadler et al. (2015)

## Univariate meta-analysis
```{r}
## Show the first few cases
head(Stadler15)

summary(meta(y=r, v=v, data=Stadler15))
```

## Three-level meta-analysis
### Without any moderator
```{r}
## Display the number of effect sizes per study
table(Stadler15$Authors)

fit4 <- meta3(y=r, v=v, cluster=Authors, data=Stadler15)
summary(fit4)
```

### With the `IntelligenceMeasure` as a moderator
```{r}
## Display the number of effect sizes per outcome measure
table(Stadler15$IntelligenceMeasure)

fit5 <- meta3(y=r, v=v, cluster=Authors, data=Stadler15,
              intercept.constraints=0,
              x=model.matrix(~ -1+IntelligenceMeasure))

## Compare the models with and without the equality constraint
anova(fit5, fit4)

## Slope_1: General intelligence
## Slope_2: Reasoning
summary(fit5)
```

# Running the analyses in Mplus
```{r, echo=FALSE}
mplus <- function(x, msg="my message") {  
 ## Mplus inp file
 my.inp <- paste("./", x,".inp", sep="")
 txt.inp <- readLines(my.inp, warn = FALSE)
 ## Mplus output file
 my.out <- paste("./", x,".out", sep="")
 txt.out <- readLines(my.out, warn = FALSE)
 
 cat("[",msg,"](", my.inp, ")\n", sep="")
 
 cat("```\n\n")
 cat(txt.inp, sep="\n", append=TRUE)
 cat("```\n\n\n")
 
 cat("[Mplus output file](", my.out, ")\n", sep="")
 
 cat("```\n\n")
 cat(txt.out, sep="\n", append=TRUE)
 cat("```\n\n\n")  
}
```

## Export the data to Mplus
* Multivariate meta-analysis
```{r, eval=FALSE}
## Get a copy of the known sampling variances and covariance
my.v <- Nam03[, c("Asthma_v", "AsthmaLRD_cov_05", "LRD_v")]

## Replace the missing value with 0.
## Since data with missing effect sizes are automatically filtered,
## the imputed value of 0 does not affect the analysis.
my.v[is.na(my.v)] <- 0  

## Combine the effect sizes with the known sampling variances and covariance.
my.df <- cbind(Nam03[, c("ID", "Age", "Asthma_logOR", "LRD_logOR")], my.v)

## Display the few cases
head(my.df)

## Write it as a data file
write.table(my.df, "Nam03.dat", na = ".", row.names = FALSE, col.names = FALSE)
```

* Three-level meta-analysis
```{r, eval=FALSE}
## Get unique numeric labels for the level 3 (study)
## as Mplus only allows numeric data
Study_ID <- with(Stadler15, as.numeric(factor(Authors, level=unique(Authors))))

## Create dummy codes for the Intelligence Measure
GenInt <- ifelse(Stadler15$IntelligenceMeasure=="General intelligence", yes=1, no=0)
Reason <- ifelse(Stadler15$IntelligenceMeasure=="Reasoning", yes=1, no=0)

my.df <- cbind(ID=Stadler15$ID, Study_ID,
               Stadler15[, c("r", "v")], GenInt, Reason)

## Display the few cases
head(my.df, n=8)

## Write it as a data file
write.table(my.df, "Stadler15.dat", na = ".", row.names = FALSE, col.names = FALSE)
```

## Multivariate meta-analysis
```{r, echo=FALSE, comment="#"}
mplus("Nam03a", "Multivariate random-effects meta-analysis")
mplus("Nam03b", "Multivariate random-effects meta-analysis with equality constraints")
mplus("Nam03c", "Multivariate mixed-effects meta-analysis")
```

## Three-level meta-analysis
```{r, echo=FALSE, comment="#"}
mplus("Stadler15a", "Three-level random-effects meta-analysis")
mplus("Stadler15b", "Three-level mixed-effects meta-analysis")
```

```{r}
cat("\n\n")
## Settings of the computing environment
sessionInfo()
```
