---
title: 'Supplemental materials of "A Tutorial on Fitting Flexible Meta-Analytic Models with Structural Equation Modeling in R"'
author: "Mike Cheung"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
  word_document: default
editor_options: 
  chunk_output_type: console
---

# Univariate Meta-Analyses

## Data preparation
```{r, message=FALSE}
## Load the required libraries
library(metaSEM)
library(symSEM)

## Prepare a sample data
dat1 <- data.frame(yi=Jaramillo05$r, 
                   vi=Jaramillo05$r_v,
                   xi=scale(Jaramillo05$IDV, scale=FALSE),
                   zi=scale(Jaramillo05$Citations, scale=TRUE))
```

## Models without any external variables
### Fixed-effect model
```{r}
## Fixed-effect model with a latent variable ei
m1a <- "ei =~ 1*yi           ## e as a latent variable and yi is its indicator
        yi ~ mu*1            ## Mean(yi) = mu
        yi ~~ 0*yi           ## Set the error variance of yi at 0
        ei ~~ data.vi*ei     ## Var(yi) = vi, which is assigned from data
"

## Fixed-effects model without specifying ei
m1b <- "yi ~ mu*1            ## Mean(yi) = mu
        yi ~~ data.vi*yi     ## Var(yi) = vi, which is assigned from data
"

## Display the conceptual model using the semPlot package
## sizeInt=7: Set 7 as the size of the intercept, see help(semPaths) in R
plot(m1a, color = "yellow", sizeInt = 7)

## Convert the lavaan model to RAM specification
## obs.variables = "yi": "yi" is the observed variable
ram1 <- lavaan2RAM(m1a, obs.variables = "yi", std.lv = FALSE)

## Display the RAM model
ram1

## Get the model-implied mean and variance structures
impliedS(ram1)

## Fit the model to the data
fit1 <- sem("Fixed", RAM = ram1, data = dat1)
summary(fit1)

## Display the model with results using the semPlot package
## nDigits=4: Display 4 digits
plot(fit1, color = "green", sizeInt = 7, nDigits = 4)
```

### Random-effects model
```{r}
## Random-effects model
m2 <- "ui =~ 1*yi          ## ui is the random effect
       ui ~~ tau2*ui       ## Var(ui) = tau2
       yi ~ mu*1           ## Mean(yi) = mu
       yi ~~ data.vi*yi    ## Var(yi) = vi, which is assigned from data
"

## Display the conceptual model
plot(m2, color="yellow", sizeInt = 7)

## Convert the lavaan model to RAM specification
## std.lv = FALSE: Ensure the latent variable is not standardized
ram2 <- lavaan2RAM(m2, obs.variables = "yi", std.lv = FALSE)

## Get the model-implied mean and variance structures
impliedS(ram2)

## intervals.type = "LB": Request likelihood-based CI
fit2 <- sem("Random", RAM = ram2, data = dat1, intervals.type = "LB")
summary(fit2)

## Display the model with results
plot(fit2, color="green", sizeInt = 7, nDigits = 4)
```

### Multiplicative-error model
```{r}
## Multiplicative error, same as UWLS model
m3 <- "ei =~ phi_sqrt*yi   ## phi (phi_sqrt^2) is the heterogeneity scaling factor
       ei ~~ data.vi*ei    ## Var(ei) = vi
       yi ~ mu*1           ## Mean(yi) = mu
       yi ~~ 0*yi          ## Set the default error variance of yi at 0
       phi := phi_sqrt^2   ## Define phi as a function of phi_sqrt
"

plot(m3, color = "yellow", sizeInt = 7)

ram3 <- lavaan2RAM(m3, obs.variables = "yi", std.lv = FALSE)

impliedS(ram3)

fit3 <- sem("Multiplicative", RAM = ram3, data = dat1, intervals.type = "LB")
summary(fit3)
plot(fit3, color="green", sizeInt=6, nDigits=4)
```

### Hybrid models of additive and multiplicative errors
```{r}
## Hybrid model 1: multiplicative and addictive error
m4 <- "yi ~ mu*1          ## Mean(yi) = mu 
       yi ~~ 0*yi
       ## additive error
       ui =~ 1*yi
       ui ~~ tau2*ui      ## Var(ui) = tau2
       ## mulitplicative error
       ei =~ phi_sqrt*yi  ## phi is a scaling factor
       ei ~~ data.vi*ei   ## Var(ei) = vi
       phi := phi_sqrt^2  ## Define phi as a function of phi_sqrt
"
plot(m4, color="yellow", sizeInt=6)

ram4 <- lavaan2RAM(m4, obs.variables = "yi", std.lv = FALSE)

impliedS(ram4)

fit4 <- sem("Hybrid1", RAM = ram4, data = dat1, intervals.type = "LB")
summary(fit4)

plot(fit4, color="green", sizeInt=6, nDigits=4)


## Hybrid model 2: multiplicative and addictive error
m5 <- "yi ~ mu*1         ## Mean(yi) = mu
       yi ~~ Vary*yi     ## Var(yi) = vary
       ## Define Vary as a function of vi, tau2, and omega
       Vary == data.vi + tau2*data.vi^omega
"

ram5 <- lavaan2RAM(m5, obs.variables = "yi", std.lv = FALSE)

impliedS(ram5, replace.constraints = TRUE)

fit5 <- sem("Hybrid2", RAM = ram5, data = dat1, intervals.type = "LB", 
            replace.constraints = TRUE)
summary(fit5)

## Hybrid model 3: Hartung-Knapp-Sidik-Jonkman method
m6 <- "yi ~ mu*1            ## Mean(yi) = mu
       yi ~~ 0*yi
       ## additive error
       ui =~ phi_sqrt*yi    ## phi is a scaling factor
       ui ~~ 0.01703407*ui  ## tau2 obtained from the random-effects model
       ## mulitplicative error
       ei =~ phi_sqrt*yi    ## phi is a scaling factor
       ei ~~ data.vi*ei     ## Var(ei) = vi
       phi := phi_sqrt^2    ## Define phi as a function of phi_sqrt
"
plot(m6, color="yellow", sizeInt=6)

ram6 <- lavaan2RAM(m6, obs.variables = "yi", std.lv = FALSE)

impliedS(ram6)

fit6 <- sem("Hybrid3", RAM = ram6, data = dat1, intervals.type = "LB")
summary(fit6)

plot(fit6, color="green", sizeInt=6, nDigits=4)
```

### Comparison of the six models
```{r}
## Refit some models to get SEs rather than LBCIs for comparisons
fit2 <- sem("Random", RAM = ram2, data = dat1)
fit3 <- sem("Multiplicative", RAM = ram3, data = dat1)
fit4 <- sem("Hybrid1", RAM = ram4, data = dat1)
fit5 <- sem("Hybrid2", RAM = ram5, data = dat1, replace.constraints = TRUE)
fit6 <- sem("Hybrid3", RAM = ram6, data = dat1)

## A helper function to extract the results
extractResults <- function(x) {
  out <- summary(x)
  para <- out$coefficients[c("mu", "tau2", "phi_sqrt", "omega"), 
                           c("Estimate", "Std.Error")]
  out <- c(unlist(para), out$estPara, out$Minus2LL, 
           out$informationCriteria["AIC:", "Parameters Penalty"],
           out$informationCriteria["BIC:", "Parameters Penalty"])
  names(out) <- c("mu", "tau2", "phi_sqrt", "omega",
                  "mu_SE", "tau2_SE", "phi_sqrt_SE", "omega_SE",
                  "No_parameters",  "-2LL", "AIC", "BIC")
  out
}

table1 <- sapply(list(fit1, fit2, fit3, fit4, fit5, fit6), extractResults)
table1 <- data.frame(Name=rownames(table1), table1)
rownames(table1) <- NULL
colnames(table1) <- c("Name", "Fixed", "Random", "Multiplicative", "Hybrid1", 
                      "Hybrid2", "Hybrid3")
## Add the estimated tau2 and tau2_SE from the random-effects model to
## the table in an ad-hoc manner
table1[table1$Name=="tau2", "Hybrid3"] <- table1[table1$Name=="tau2", "Random"]
table1[table1$Name=="tau2_SE", "Hybrid3"] <- table1[table1$Name=="tau2_SE", "Random"]
knitr::kable(table1, digits=4, caption="Results of the Six Meta-Analytic Models")
```

## Models with external variables
### A fixed-x meta-regression
```{r, message=FALSE}
## Mixed-effects model with the covariate as a design matrix
m7 <- "yi ~ b0*1         ## Intercept of yi: b0
       yi ~~ data.vi*yi  ## Known sampling variance vi
       ui =~ 1*yi
       ui ~~ tau2*ui     ## Heterogeneity variance tau2
       ## etai: a phantom latent variable to represent the covariate x
       etai =~ b1*yi     ## yi = b1*etai
       etai ~ data.xi*1  ## etai = xi
       etai ~~ 0*etai    ## Var(etai) = 0
"

plot(m7, color = "yellow", layout = "spring")

## Only yi is an observed variable
ram7 <- lavaan2RAM(m7, obs.variables = "yi", std.lv = FALSE)

impliedS(ram7)

fit7 <- sem("Covariate_design", RAM=ram7, data=dat1, intervals.type = "LB")
summary(fit7)
plot(fit7, color = "green", sizeInt=6,  nDigits = 4)
```

### A random-x meta-regression
```{r}
## Mixed-effects model with the covariate as a variable
m8 <- "yi ~ b0*1 + b1*xi ## Intercept (b0) and slope (b1)
       yi ~~ data.vi*yi  ## Known sampling variance vi
       ui =~ 1*yi
       ui ~~ tau2*ui     ## Heterogeneity variance tau2
       xi ~ Meanx*1      ## Mean(xi) = Meanx
       xi ~~ Varx*xi     ## Var(xi) = Varx
       ## Define R2: explained variance/total variance
       R2 := b1^2*Varx/(b1^2*Varx + tau2)
"

plot(m8, color = "yellow", layout = "spring", sizeInt=6)

## Both yi and xi are observed variables
ram8 <- lavaan2RAM(m8, obs.variables = c("yi", "xi"), std.lv = FALSE)

impliedS(ram8)

## Request LBCIs on the R2
fit8 <- sem("Covariate_variable", RAM=ram8, data=dat1, intervals.type = "LB")
summary(fit8)

plot(fit8, color="green", sizeInt=6, nDigits = 4, layout = "spring")
```

### Regressing a covariate on the true effect size
```{r}
## Regressing a covariate on the true effect size
m9 <- "fyi =~ 1*yi         ## fi: true effect size
       fyi ~~ tau2*fyi     ## Var(fi) = tau2
       fyi ~ Meany*1       ## Mean(fi) = Meany
       yi ~ 0*1         
       yi ~~ data.vi*yi    ## Var(ei) = vi
       zi ~ b0*1 + b1*fyi  ## Intercept (b0) and slope (b1)
       zi ~~ ErrVarz*zi    ## Error variance of zi
       ## Define an R2
       R2 := b1^2*tau2/(b1^2*tau2 + ErrVarz)
"

plot(m9, color = "yellow", sizeInt=6, layout = "spring")

## Both yi and xi are observed variables
ram9 <- lavaan2RAM(m9, obs.variables = c("yi", "zi"), std.lv = FALSE)

impliedS(ram9)

fit9 <- sem("Regress_covariate", RAM=ram9, data=dat1, intervals.type = "LB")
summary(fit9)

plot(fit9, color = "green", sizeInt=6, nDigits = 4)
```

### Correlating a covariate with a true effect size
```{r}
## Correlation between fyi and fxi
m10 <- "## fyi: Standardized true effect size of yi
        fyi =~ tauy*yi     ## tauy: sd of fyi
        fyi ~~ 1*fyi       ## Var(fyi) = 1
        fyi ~ y_mean_div_tauy*1   ## rescaled mean
        yi ~ 0*1
        yi ~~ data.vi*yi
        ## fxi: Standardized score of xi
        fxi =~ sigma*xi    ## sigma: sd of xi
        fxi ~~ 1*fxi       ## Var(fxi) = 1
        fxi ~~ Cor*fyi
        fxi ~ x_mean_div_sigmax*1 ## rescaled mean
        xi ~ 0*1
        xi ~~ 0*xi       
"

plot(m10, color = "yellow", sizeInt=6)

## Both yi and xi are observed variables
ram10 <- lavaan2RAM(m10, obs.variables = c("yi", "xi"), std.lv = TRUE)

impliedS(ram10)

fit10 <- sem("Correlation", RAM=ram10, data=dat1)
summary(fit10)
plot(fit10, color = "green", sizeInt=6, nDigits = 4)
```

### Meta-regression with completely and partially standardized variables
```{r}
## Completely standardized fyi and xi
m11 <- "## fyi: Standardized true effect size of yi
        fyi =~ tauy*yi       ## tauy: sd of fyi
        fyi ~~ tau2_res*fyi  ## tau2_res: residual heterogeneity
        fyi ~ b0*1 + b1*fxi
        yi ~ 0*1
        yi ~~ data.vi*yi
        ## fxi: Standardized score of xi
        fxi =~ sigma*xi      ## sigma: sd of xi
        fxi ~~ 1*fxi         ## Var(fxi) =1
        fxi ~ mu_x_sd*1
        xi ~ 0*1
        xi ~~ 0*xi
        ## Nonlinear constraint to ensure Var(fyi) = 1
        tau2_res == 1 - b1^2
        ## Define R2
        R2 := b1^2
"

plot(m11, color="yellow", layout="spring", sizeInt=6)

## Both yi and xi are observed variables
ram11 <- lavaan2RAM(m11, obs.variables = c("yi", "xi"), std.lv = FALSE)

impliedS(ram11, replace.constraints = TRUE)

fit11 <- sem("Complete_standardize", RAM=ram11, data=dat1, intervals.type = "LB",
             lbound = list(tauy=0))
summary(fit11)
plot(fit11, color="green", sizeInt=6, nDigits=4)

## Partial standardized xi but not yi
m12 <- "## fyi: Unstandardized true effect size of yi
        fyi =~ 1*yi
        fyi ~~ tau2_res*fyi   ## tau2_res: residual heterogeneity
        fyi ~ b0*1 + b1*fxi
        yi ~ 0*1
        yi ~~ data.vi*yi
        ## fxi: Standardized score of xi
        fxi =~ sigma*xi       ## sigma: sd of xi
        fxi ~~ 1*fxi          ## Var(fxi) =1
        fxi ~ mu_x_sd*1
        xi ~ 0*1
        xi ~~ 0*xi
        ## Define R2
        R2 := b1^2/(b1^2 + tau2_res)
"

plot(m12, color="yellow", layout="spring", sizeInt=6)

## Both yi and xi are observed variables
ram12 <- lavaan2RAM(m12, obs.variables = c("yi", "xi"), std.lv = FALSE)
ram12
impliedS(ram12)

fit12 <- sem("Partial_standardize", RAM=ram12, data=dat1, intervals.type = "LB")
summary(fit12)
plot(fit12, color="green", sizeInt=6, nDigits=4)
```

### Location-scale and nonlinear models
```{r}
## Location-scale model with an additive heterogeneity variance
m13 <- "yi ~ mu*1          ## Mean(yi) = mu
        yi ~~ data.vi*yi   ## Var(ei) = vi 
        ## ui is the latent variable of the random effect
        ui =~ 1*yi         ## yi = 1*ui
        ui ~~ tau2*ui      ## Var(ui) = tau2
        ## Specify nonlinear on mean and tau2
        mu == b0 + b1*data.xi
        tau2 == exp(a0 + a1*data.xi)
"

ram13 <- lavaan2RAM(m13, obs.variables="yi", std.lv = FALSE)

## Get the model implied structures
## We need to replace the constraints with the new parameters
impliedS(ram13, replace.constraints=TRUE)

fit13 <- sem("Location_additive", RAM=ram13, data=dat1,
             replace.constraints=TRUE)
summary(fit13)


## Location-scale model with a multiplicative heterogeneity variance
m14 <- "yi ~ mu*1          ## Mean(yi) = mu
        yi ~~ sigma2*yi    ## Var(yi) = sigma2 
        ## Specify nonlinear on mean and sigma2
        mu == b0 + b1*data.xi
        sigma2 == exp(a0 + a1*data.xi)*data.vi
"

ram14 <- lavaan2RAM(m14, obs.variables="yi", std.lv = FALSE)

impliedS(ram14, replace.constraints = TRUE)

fit14 <- sem("Location_multiplicative", RAM=ram14, data=dat1,
             replace.constraints=TRUE)
summary(fit14)
```

# Multivariate Meta-Analyses
## Data preparation
```{r}
## Replace missing values with 0 as NA is not allowed in definition variables
Chan17$v_misinfo[is.na(Chan17$v_misinfo)] <-0 
Chan17$c_mis_deb[is.na(Chan17$c_mis_deb)] <-0
Chan17$v_debunk[is.na(Chan17$v_debunk)] <-0

dat2 <- data.frame(y1i=Chan17$g_misinfo,
                   y2i=Chan17$g_debunk,
                   V11i=Chan17$v_misinfo,
                   V21i=Chan17$c_mis_deb,
                   V22i=Chan17$v_debunk,
                   se11i=sqrt(Chan17$v_misinfo),
                   se22i=sqrt(Chan17$v_debunk),
                   xi=scale(Chan17$PublicationYear))

sessionInfo()
```


