---
title: 'Illustration of MASEM: Theory of planned behavior'
author: "Mike Cheung and Ryan Hong"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
---

This file includes examples used in Cheung and Hong (in press):
Cheung, M. W.-L., & Hong, R. Y. (in press). Applications of meta-analytic structural equation modeling in health psychology: Examples, issues, and recommendations. *Health Psychology Review*.

# Displaying some descriptive statistics of the data
```{r, message=FALSE}
## Load the libraries required for the analysis
library("metaSEM")
library("semPlot")

## Read the data
# source("Cooke16.R")

## Check whether the correlation matrices are valid (positive definite)
## Since the correlation matrix in Study 3 is not valid, we excluded it from the analyses.
is.pd(Cooke16$data)

## Since the correlation matrix in Study 3 is not positive definite,
## we may exclude it the following analyses
my.data <- Cooke16$data[-3]
my.n <- Cooke16$n[-3]

## Show the no. of studies per correlation
pattern.na(my.data, show.na = FALSE)

## Show the total sample sizes per correlation
pattern.n(my.data, my.n)
```

# Stage 1 analysis
```{r, message=FALSE}
## Fixed-effects model: method="FEM"
fixed1 <- tssem1(my.data, my.n, method="FEM")
summary(fixed1)

## Common correlation matrix under a fixed-effects model
coef(fixed1)

## Random-effects model
random1 <- tssem1(my.data, my.n, method="REM", RE.type="Diag")
summary(random1)

## Average correlation matrix under a random-effects model
vec2symMat(coef(random1, select="fixed"), diag = FALSE)
```

# Stage 2 analysis: original model (Model A)
* The models are specified in the [RAM formulation](http://openmx.psyc.virginia.edu/docs/OpenMx/latest/Examples_Matrix.html). Matrices `A`, `S` and `F` represent the regression coefficients, variance-covariance matrix of variables, and selection matrix of the latent variables, respectively. We may skip the `F` matrix when there is no latent variable.
* Consider the element `SN2BI` in `A1` as an example. It is the regression coefficient from `SN` to `BI` with a starting value of 0.2.
```{r, message=FALSE}
A1 <- create.mxMatrix(c(0,0,0,0,0,
                        0,0,0,0,0,
                        0,0,0,0,0,
                        "0.2*SN2BI","0.2*ATT2BI","0.2*PBC2BI",0,0,
                        0,0,"0.2*PBC2BEH","0.2*BI2BEH",0),
                        type="Full", ncol=5, nrow=5,
                        byrow=TRUE, as.mxMatrix=FALSE)

## This step is not necessary but it is useful for inspecting the model.
dimnames(A1)[[1]] <- dimnames(A1)[[2]] <- colnames(Cooke16$data[[1]])

## Display A1
A1

S1 <- create.mxMatrix(c(1,
                        "0.1*ATT_SN", 1,
                        "0.1*PBC_SN", "0.1*PBC_ATT", 1,
                        0, 0, 0, "0.5*VarBI",
                        0, 0, 0, 0, "0.5*VarBEH"),
                      type = "Symm", ncol=5, nrow=5,
                      byrow=TRUE, as.mxMatrix=FALSE)

dimnames(S1)[[1]] <- dimnames(S1)[[2]] <- colnames(Cooke16$data[[1]])
S1
```

* The columns `lbound` and `ubound` are the lower and upper bounds of the 95% confidence intervals. If they include 0, the parameter estimates are not statistically significant at $\alpha=.05$.
```{r}
ModelA <- tssem2(random1, Amatrix=A1, Smatrix=S1, diag.constraints=TRUE, intervals.type="LB")
summary(ModelA)
```

* To facilitate the interpretations, we may print the model with the parameter estimates using the `semPlot` package.
```{r}
## Convert the model to semPlotModel object
my.plota <- meta2semPlot(ModelA)

## Plot the parameter estimates
semPaths(my.plota, whatLabels="est", edge.label.cex=1, color="yellow")
```

# Stage 2 analysis: no direct effect from PBC (Model C)
```{r}
## Delete the direct effect from PBC to BEH
A2 <- create.mxMatrix(c(0,0,0,0,0,
                        0,0,0,0,0,
                        0,0,0,0,0,
                        "0.2*SN2BI","0.2*ATT2BI","0.2*PBC2BI",0,0,
                        0,0,0,"0.2*BI2BEH",0),
                        type="Full", ncol=5, nrow=5,
                        byrow=TRUE, as.mxMatrix=FALSE)

## This step is not necessary but it is useful for inspecting the model.
dimnames(A2)[[1]] <- dimnames(A2)[[2]] <- colnames(Cooke16$data[[1]])

## Display A1
A2

ModelC <- tssem2(random1, Amatrix=A2, Smatrix=S1, diag.constraints=TRUE, intervals.type="LB")
summary(ModelC)

## Convert the model to semPlotModel object
my.plotc <- meta2semPlot(ModelC)

## Plot the parameter estimates
semPaths(my.plotc, whatLabels="est", edge.label.cex=1, color="yellow")
```

# Stage 2 analysis: all direct effects (Model B)
```{r}
## Delete the direct effect from PBC to BEH
A3 <- create.mxMatrix(c(0,0,0,0,0,
                        0,0,0,0,0,
                        0,0,0,0,0,
                        "0.2*SN2BI","0.2*ATT2BI","0.2*PBC2BI",0,0,
                        "0.2*SN2BEH","0.2*ATT2BEH","0.2*PBC2BEH","0.2*BI2BEH",0),
                        type="Full", ncol=5, nrow=5,
                        byrow=TRUE, as.mxMatrix=FALSE)

## This step is not necessary but it is useful for inspecting the model.
dimnames(A3)[[1]] <- dimnames(A3)[[2]] <- colnames(Cooke16$data[[1]])

## Display A1
A3
```

* We may calculate functions of parameters with the `mx.algebras` argument. In this example, we calculate the indirect and direct effects. After the analysis, 95% CI on the indirect and direct effects are estimated.
```{r}
ModelB <- tssem2(random1, Amatrix=A3, Smatrix=S1, diag.constraints=TRUE, intervals.type="LB",
                 mx.algebras=list(indirect=mxAlgebra((SN2BI+ATT2BI+PBC2BI)*BI2BEH, name="indirect"),
                                  direct=mxAlgebra(SN2BEH+ATT2BEH+PBC2BEH, name="direct")))
summary(ModelB)

## Convert the model to semPlotModel object
my.plotb <- meta2semPlot(ModelB)

## Plot the parameter estimates
semPaths(my.plotb, whatLabels="est", edge.label.cex=1, color="yellow")

## Settings of R for the analysis
sessionInfo()
```
