---
title: "Emotional Intelligence Mediates the Connection Between Mindfulness and Gratitude"
author: "Nicola Schutte, Keng Shian Ling, and Mike Cheung"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
    toc_depth: 2
  pdf_document:
    toc: no
  word_document: null
editor_options: 
  chunk_output_type: console
---

# TSSEM

## Data preparation
```{r, message=FALSE}
## Whether to output figures as SVG format
plot.figures <- FALSE

library(metaSEM)
mxOption(key='Number of Threads', value=parallel::detectCores()-2)

## Read data
my.df <- foreign::read.spss("data.sav", use.value.labels = TRUE, to.data.frame=TRUE)

## A function to convert rows into a 3x3 correlation matrix
create.matrix <- function(x, type=c(1, 2, 3)) {
  mat <- matrix(NA, ncol=3, nrow=3)
  diag(mat) <- 1
  type <- as.character(type)
  ## Mindfulness, EI, Gratitude
  ## 1: Mindfulness and EI
  ## 2: Mindfulness and Gratitude
  ## 3: EI and Gratitude
  switch(type,
         "1" = mat[1, 2] <- mat[2, 1] <- unlist(x),
         "2" = mat[1, 3] <- mat[3, 1] <- unlist(x),
         "3" = mat[2, 3] <- mat[3, 2] <- unlist(x))
  mat
}

varlist <- c("Mindfulness", "EI", "Gratitude")

my.cor <- lapply(split(my.df, seq(nrow(my.df))),
                 function(x, y) create.matrix(x["Effect_size"], x["Type_of_Association"]))
my.cor <- lapply(my.cor, function(x) {dimnames(x) <- list(varlist, varlist); x}  )
names(my.cor) <- my.df$Study_name

## Correlation matrices in the analysis
my.cor

## Sample sizes
my.n <- my.df$N
my.n

## Number of studies in each cell
pattern.na(my.cor, show.na = FALSE)

## Total sample sizes in each cell
pattern.n(my.cor, my.n)
```

## First-stage of analysis
```{r}
## Stage 1 analysis: find an average correlation matrix
stage1 <- tssem1(my.cor, my.n)
summary(stage1)

## Average correlation matrix
meanR <- vec2symMat(coef(stage1, select = "fixed"), diag = FALSE)
dimnames(meanR) <- list(varlist, varlist)
meanR

## Absolute heterogeneity variance: tau^2
tau2 <- vec2symMat(coef(stage1, select = "random"), diag = FALSE)
dimnames(tau2) <- list(varlist, varlist)
tau2

## Relative heterogeneity index: I^2
I2 <- vec2symMat(summary(stage1)$I2.values[, "Estimate"], diag = FALSE)
dimnames(I2) <- list(varlist, varlist)
I2
```

## Second-stage of analysis
```{r, message=FALSE}
## Proposed model
model <- "Gratitude ~ c*Mindfulness + b*EI
          EI ~ a*Mindfulness
          Mindfulness ~~ 1*Mindfulness"
plot(model, color="yellow")

RAM1 <- lavaan2RAM(model, obs.variables = varlist)
RAM1

## Stage 2 analysis: fit the path model
stage2 <- tssem2(stage1, RAM=RAM1, intervals.type = "LB",
                 mx.algebras = list(Indirect=mxAlgebra(a*b, name="Indirect"),
                                    Direct=mxAlgebra(c, name="Direct")))
summary(stage2)

plot(stage2, color="yellow")

# svg("Fig4.svg", width=5, height=5)
# plot(stage2, col="yellow")
# dev.off()

## Testing the hypothesis c = a*b
stage2b <- tssem2(stage1, RAM=RAM1, intervals.type = "LB",
                 mx.algebras = list(Indirect=mxAlgebra(a*b, name="Indirect"),
                                    Direct=mxAlgebra(c, name="Direct")),
                 run=FALSE)
## Add a constraint on c=a*b
## Rerun to remove errors
stage2b <- mxModel(stage2b, mxConstraint(c==a*b, name="constraint"))
stage2b <- mxRun(stage2b)
stage2b <- mxTryHard(stage2b, extraTries = 100)
stage2b <- mxRun(stage2b, intervals = TRUE)
summary(stage2b)

anova(stage2$mx.fit, stage2b)
```

# OSMASEM

## Data preparation
```{r}
## Mindfulness scale for the correlation between mindfulness and EI
table(my.df[my.df$Type_of_Association=="mindfulness and emotional intelligence", "Mindfulness_Scale"], useNA="always")

## Mindfulness scale for the correlation between mindfulness and gratitude
table(my.df[my.df$Type_of_Association=="mindfulness and gratitude", "Mindfulness_Scale"], useNA="always")

## Indicators for mindfulness scale
FFMQ<- ifelse(my.df$Mindfulness_Scale=="FFMQ", yes=1, no=0)
Freiburg<- ifelse(my.df$Mindfulness_Scale=="Freiburg", yes=1, no=0)
MAAS<- ifelse(my.df$Mindfulness_Scale=="MAAS", yes=1, no=0)
kentucky<- ifelse(my.df$Mindfulness_Scale=="kentucky", yes=1, no=0)
CAMM <- ifelse(my.df$Mindfulness_Scale=="CAMM", yes=1, no=0)
other<- ifelse(my.df$Mindfulness_Scale=="other", yes=1, no=0)

FFMQ[is.na(FFMQ)] <- 0
Freiburg[is.na(Freiburg)] <- 0
MAAS[is.na(MAAS)] <- 0
kentucky[is.na(kentucky)] <- 0
CAMM[is.na(CAMM)] <- 0
other[is.na(other)] <- 0

tssem.df <- list(data=my.cor, 
                 n=my.n,
                 ## Center the Age and Female for ease of interpretations
                 Age=c(scale(my.df$Age, scale=FALSE)), 
                 Female=c(scale(my.df$Percent_female, scale=FALSE)),
                 Publish=ifelse(my.df$Publication_status=="published", yes=1, no=0),
                 FFMQ=FFMQ,
                 Freiburg=Freiburg,
                 MAAS=MAAS,
                 kentucky=kentucky,
                 CAMM=CAMM,
                 other=other)

## Summary of mean age
summary(tssem.df$Age)

## Summary of percentage of female participants
summary(tssem.df$Female)

## No. of studies published vs. unpublished
table(tssem.df$Publish)

os.df <- Cor2DataFrame(tssem.df)

## Show the first few studies
# head(os.df)

## Checking only
# os.fit0 <- osmasem("No moderator", RAM=RAM1, data=os.df)
# summary(os.fit0)
# VarCorr(os.fit0)
# plot(os.fit0)
```

## Mindfulness scale
```{r, message=FALSE, message=FALSE}
## Select data without missing data in Female
indexMindScale <- !is.na(os.df$data$FFMQ)

## Note that some studies are dropped because of missing values in mindfulness scale
os.fit0 <- osmasem("No moderator", RAM=RAM1, data=os.df)
summary(os.fit0)

## Create several A matrices to test the moderating effect from Mindfulness to EI [2,1]
A_FFMQ <- matrix(0, ncol=3, nrow=3,
                 dimnames = list(c("Mindfulness", "EI", "Gratitude"),
                                 c("Mindfulness", "EI", "Gratitude")))
A_Freiburg <- A_MAAS <- A_kentucky <- A_CAMM <- A_other <- A_FFMQ
A_FFMQ[2,1] <- "0*data.FFMQ"
A_Freiburg[2,1] <- "0*data.Freiburg"
A_MAAS[2,1] <- "0*data.MAAS"
A_kentucky[2,1] <- "0*data.kentucky"
A_CAMM[2,1] <- "0*data.CAMM"
## A_other[2,1] <- "0*data.other"

## A_other is not included as it is used as the reference. 
os.fitMind <- osmasem("Mindfulness scale", RAM=RAM1, data=os.df,
                      Ax=list(A_FFMQ, A_Freiburg, A_MAAS, A_kentucky, A_CAMM))
summary(os.fitMind)

anova(os.fitMind, os.fit0)

## Compute the path coefficients on each mindfulness scale
est <- c(mxEval(c(a, a+a_1, a+a_2, a+a_3, a+a_4, a+a_5), os.fitMind$mx.fit))
names(est) <- c("Other", "FFMQ", "Freiburg", "MAAS", "Kentucky", "CAMM")

Cov <- mxSE(rbind(a, a+a_1, a+a_2, a+a_3, a+a_4, a+a_5), os.fitMind$mx.fit, details = TRUE,
           silent=TRUE)$Cov
dimnames(Cov) <- list(c("Other", "FFMQ", "Freiburg", "MAAS", "Kentucky", "CAMM"),
                      c("Other", "FFMQ", "Freiburg", "MAAS", "Kentucky", "CAMM"))
se <- sqrt(diag(Cov))

p <- 2*pnorm(-abs(est/se))
z <- qnorm(.975)
lbound <- est - z*se
ubound <- est + z*se
Mindscale <- cbind(est, se, p, lbound, ubound)
dimnames(Mindscale) <- list(c("Other", "FFMQ", "Freiburg", "MAAS", "Kentucky", "CAMM"),
                            c("Estimate", "se", "p value", "lbound (95%)", "ubound (95%)"))  
Mindscale

## Compute the post-hoc comparisons with adjusted with multiple comparisons 
library(cg)
comp <- comparisons(est, Cov, endptscale = "original", type = "pairwise", mcadjust=TRUE, display="none")

knitr::kable(comp, caption="Comparisons Table: Differences (A-B)\n95% Confidence (alpha of 0.05), Multiplicity Adjusted")
```

## Mean age
```{r, message=FALSE}
## Select data without missing data in Age
indexAge <- !is.na(os.df$data$Age)

## For checking only
os.fit0 <- osmasem("No moderator", RAM=RAM1, data=os.df, subset.rows=indexAge)
summary(os.fit0)

A.A <- create.modMatrix(RAM1, mod="Age")
os.fitA <- osmasem("Age as a moderator", RAM=RAM1, Ax=A.A, data=os.df, 
                   subset.rows=indexAge)
os.fitA <- rerun(os.fitA)
summary(os.fitA)

anova(os.fitA, os.fit0)

osmasemR2(os.fitA, os.fit0)

## Test A[2,1] (a): significant
A.21 <- A.A
A.21[3,1] <- A.21[3,2] <- 0
A.21
os.fitA21 <- osmasem("Age A21", RAM=RAM1, Ax=A.21, data=os.df, subset.rows=indexAge)
summary(os.fitA21)
anova(os.fitA21, os.fit0)

## Test A[3,2] (b): not significant
A.32 <- A.A
A.32[2,1] <- A.32[3,1] <- 0
A.32
os.fitA32 <- osmasem("Age A32", RAM=RAM1, Ax=A.32, data=os.df, subset.rows=indexAge)
summary(os.fitA32)
anova(os.fitA32, os.fit0)

## Test A[3,1] (c): not significant
A.31 <- A.A
A.31[2,1] <- A.31[3,2] <- 0
A.31
os.fitA31 <- osmasem("Age A31", RAM=RAM1, Ax=A.31, data=os.df, subset.rows=indexAge)
summary(os.fitA31)
anova(os.fitA31, os.fit0)

## Mean of Age: 0
## sd of Age
(Age.sd <- sd(os.df$data$Age, na.rm=TRUE))

## Path coefficients when Age is 0
mxEval(A0, os.fitA$mx.fit)

## Change of path coefficients when Age increases by 1 unit
mxEval(A1, os.fitA$mx.fit)

## Path coefficients when Age is -1SD
mxEval(A0, os.fitA$mx.fit) - Age.sd*mxEval(A1, os.fitA$mx.fit)

## Path coefficients when Age is +1SD
mxEval(A0, os.fitA$mx.fit) + Age.sd*mxEval(A1, os.fitA$mx.fit)
```

## Percentage of female participants
```{r, message=FALSE}
## Select data without missing data in Female
indexFemale <- !is.na(os.df$data$Female)

## For checking only
os.fit0 <- osmasem("No moderator", RAM=RAM1, data=os.df, subset.rows=indexFemale)
summary(os.fit0)

A.F <- create.modMatrix(RAM1, mod="Female")
os.fitF <- osmasem("Female as a moderator", RAM=RAM1, Ax=A.F, data=os.df,
                   subset.rows=indexFemale)
os.fitF <- rerun(os.fitF, extraTries = 100)
summary(os.fitF)

anova(os.fitF, os.fit0)
```

## Published vs. unpublished
```{r, message=FALSE}
## No missing covariate
## For checking only
os.fit0 <- osmasem("No moderator", RAM=RAM1, data=os.df)
summary(os.fit0)

A.P <- create.modMatrix(RAM1, mod="Publish")
os.fitP <- osmasem("Publish as a moderator", RAM=RAM1, Ax=A.P, data=os.df)
os.fitP <- rerun(os.fitP, extraTries =200)
summary(os.fitP)

anova(os.fitP, os.fit0)
```

# Sensitivity analyses on the correlations
```{r, message=FALSE}
library(metafor)

## Create study labels        
study <- row.names(os.df$data)
```

## Correlation between Mindfulness and Gratitude
```{r, message=FALSE}
fit1 <- rma(yi=os.df$data$Gratitude_Mindfulness, 
            vi=os.df$data$`C(Gratitude_Mindfulness Gratitude_Mindfulness)`,
            method="ML", slab=study)
fit1

forest(fit1, main="Correlation between Mindfulness and Gratitude", xlab="Correlation")

funnel(fit1, xlab="Correlation", main="Correlation between Mindfulness and Gratitude",
       level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), legend=TRUE)

trimfill(fit1)
hc(fit1)
regtest(fit1)
ranktest(fit1)

if (plot.figures) {
  svg("Fig3.svg", width = 10, height = 5)
  forest(fit1, main="Correlation between Mindfulness and Gratitude", xlab="Correlation")
  dev.off()
  
  svg("Fig4.svg", width = 10, height = 5)
  funnel(fit1, xlab="Correlation", main="Correlation between Mindfulness and Gratitude",
         level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), legend=TRUE)
  dev.off()  
}
```

## Correlation between Mindfulness and Emotional Intelligence
```{r, message=FALSE}
fit2 <- rma(yi=os.df$data$EI_Mindfulness, 
            vi=os.df$data$`C(EI_Mindfulness EI_Mindfulness)`,
            method="ML", slab=study)
fit2

forest(fit2, main="Correlation between Mindfulness and Emotional Intelligence", xlab="Correlation")

funnel(fit2, xlab="Correlation", main="Correlation between Mindfulness and Emotional Intelligence",
       level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), legend=TRUE)

trimfill(fit2)
hc(fit2)
regtest(fit2)
ranktest(fit2)

if (plot.figures) {
  svg("Fig5.svg", width = 10, height = 5)
  forest(fit2, main="Correlation between Mindfulness and Emotional Intelligence", xlab="Correlation")
  dev.off()
  
  svg("Fig6.svg", width = 10, height = 5)
  funnel(fit2, xlab="Correlation", main="Correlation between Mindfulness and Emotional Intelligence",
         level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), legend=TRUE)
  dev.off()  
}
```

## Correlation between Emotional Intelligence and Gratitude
```{r, message=FALSE}
fit3 <- rma(yi=os.df$data$Gratitude_EI, 
            vi=os.df$data$`C(Gratitude_EI Gratitude_EI)`,
            method="ML", slab=study)
fit3

forest(fit3, main="Correlation between Emotional Intelligence and Gratitude", xlab="Correlation")

funnel(fit3, xlab="Correlation", main="Correlation between Emotional Intelligence and Gratitude",
       level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), legend=TRUE)

trimfill(fit3)
hc(fit3)
regtest(fit3)
ranktest(fit3)

if (plot.figures) {
  svg("Fig7.svg", width = 10, height = 5)
  forest(fit3, main="Correlation between Emotional Intelligence and Gratitude", xlab="Correlation")
  dev.off()
  
  svg("Fig8.svg", width = 10, height = 5)
  funnel(fit3, xlab="Correlation", main="Correlation between Emotional Intelligence and Gratitude",
         level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), legend=TRUE)
  dev.off()  
}

sessionInfo()
```


```{r, echo=FALSE, eval=FALSE, warning=FALSE}
library(ggplot2)
p <- ggplot(my.df, aes(x=Percent_female)) + geom_histogram(binwidth=5) + 
  ggtitle("Distribution of percentage of female participants")
ggsave("gender.svg", p)
```
