---
title: 'Supplemental Materials #3: Demonstration of the Incorrect Population Correlation
  Matrices and Model Specifications in Study 1 in Yu et al. (2016)'
author: 'Mike W.-L. Cheung'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document:
    toc: yes
    keep_tex: true
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes 
  word_document:
    toc: no
---

# Figure 1 in Yu et al. (2016)

## Incorrect (actually used) population correlation used to generate the data
* Pearson correlations were incorrectly used to represent the path coefficients to generate the random correlation matrices in Yu et al. The correct approach is to calculate the model implied correlation matrix based on the path diagram in Figure 1. 
* The incorrect population correlation matrix `IncorrectP1` was used to generate data in Study 1. 
* If we use `IncorrectP1` (the population values) to fit the proposed path model in Figure 1, the value of the minimum of the fit function is non-zero ($\chi^2(df=2)=59.215$ with $N=1,000$). Moreover, the residuals of the "covariance" matrix are also non-zero. This shows that the population correlation matrix `IncorrectP1` does not match the model specified in Figure 1.

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE)
```

```{r, message=FALSE}
## Required packages
lib2install <- c("lavaan", "semPlot", "knitr")

## Install them automatically if they have not been installed in your computer yet.
for (i in lib2install) {
  if (!(i %in% rownames(installed.packages()))) install.packages(i)
}

library(lavaan)
library(semPlot)
library(knitr)

labels <- c("x", "m1", "m2", "y")

IncorrectP1 <- matrix(c(1, .3, .3, 0,
                       .3, 1, 0, .3,
                       .3, 0, 1, .3,
                       0, .3, .3, 1), ncol=4, nrow=4, byrow=TRUE,
               dimnames = list(labels, labels))

## Population correlation matrix used in Yu's et al. simulation studies
kable(IncorrectP1)

## Population model: no direct effect used in the analysis
model1 <- 'm1 + m2 ~ x
           y ~ m1 + m2'

## Incorrect model. The fit is not perfect even the population correlation matrix is used.
fit.incorrect1 <- sem(model1, sample.cov=IncorrectP1, sample.nobs=1000)
summary(fit.incorrect1)

## Residuals of the "covariance" matrix
resid(fit.incorrect1)
```

* To see what the actual generating model is, we add the direct effect from $x$ to $y$ and allow the residues between $m1$ and $m2$ correlated. This model is now saturated. The results show that there is a direct effect of -0.22 from $x$ to $y$ and the correlation between the residues of $m1$ and $m2$ is -0.09. This model is different from the one specified in Figure 1 in Yu et al.

```{r}
## Population model: with direct effect              
model2 <- 'm1 + m2 ~ x
           y ~ m1 + m2 + x
           m1 ~~ m2'

fit.incorrect2 <- sem(model2, sample.cov=IncorrectP1, sample.nobs=1000)
summary(fit.incorrect2)  

## Residuals of the "covariance" matrix
resid(fit.incorrect2)

semPaths(fit.incorrect2, what="est", edge.label.cex=1.5, 
         sizeMan=8, color="yellow", edge.color = "black", 
         weighted=FALSE, layout="tree2")
```

```{r, echo=FALSE, eval=FALSE}
my.tab <- parTable(fit.incorrect2)

lavaan2dot <- function(x) {
  x$dir <- ifelse(x$op=="~", yes="forward", no="both")
  with(x, paste0(rhs, " -> ", lhs, " [dir=", dir, ", label=", round(est, 2), "];\n"))
}

cat(lavaan2dot(my.tab), file="temp.txt")
```


## Correct (intended to use) population correlation used to generate the data

* The following `R` code shows how to derive the correct population correlation matrix for the model in Figure 1 in Yu et al.
* We use the `impliedR()` function in the `metaSEM` package. Users may specify the population standardized regression coefficients. The function then generates the population correlation matrix for the model.
* When we fit the model in Figure 1 to the population correlation matrix `CorrectP1`, the discrepancy is exact zero ($\chi^2(df=2)=0$ with $N=1,000$). The residuals of the "covariance" matrix are also zero. The parameters are identical to the values in Figure 1 in Yu et al. Therefore, this is the correct population correlation matrix.
```{r, message=FALSE}
library(metaSEM)

## A matrix for the regression paths as defined in Figure 1
## All of them are fixed values.
A2 <- matrix(c(0,0,0,0,
               0.3,0,0,0,
               0.3,0,0,0,
               0,0.3,0.3,0), ncol=4, nrow=4, byrow=TRUE,
             dimnames=list(labels, labels))
A2

## The variance of x is fixed at 1 whereas the others are starting values.
S2 <- Diag(c(1, "0.1*Err_m1", "0.1*Err_m2", "0.1*Err_y"))
dimnames(S2) <- list(labels, labels)
S2

CorrectP1 <- impliedR(A2, S2, labels=labels)$SigmaObs
kable(CorrectP1)

fit.correct1 <- sem(model1, sample.cov=CorrectP1, sample.nobs=1000)
summary(fit.correct1)

## Residuals of the "covariance" matrix
resid(fit.correct1)

semPaths(fit.correct1, what="est", edge.label.cex=1.5, 
         sizeMan=8, color="lightseagreen", edge.color = "black", 
         weighted=FALSE, layout="tree2")
```

```{r, echo=FALSE, eval=FALSE}
my.tab <- parTable(fit.correct1)

cat(lavaan2dot(my.tab), file="temp.txt")
```

## Incorrect (actually used) model specification in analyzing the data
* The roles of independent and dependent variables in the `A` matrix in `OpenMx` and `metaSEM` were incorrectly reversed in Yu et al. Moreover, the error variances were incorrectly fixed at 1.0 in the `S` matrix. Therefore, the results in Figure 1 were incorrect in Yu et al.
* The following was the model used in the simulation study in Yu et al.
```{r}
varnames <- c('X','M1','M2','Y')

A <- mxMatrix('Full', ncol=4, nrow=4, byrow=T,
              values = c(0,0.3,0.3,0,
                         0,0,0,0.3,
                         0,0,0,0.3,
                         0,0,0,0),
              free=c(F,T,T,F,
                     F,F,F,T,
                     F,F,F,T,
                     F,F,F,F
              ),
              labels=c(NA,"betaxm1","betaxm2",NA,
                       NA,NA,NA,"betam1y",
                       NA,NA,NA,"betam2y",
                       NA,NA,NA,NA
              ),
              name="A")

S <- mxMatrix('Full', ncol=4, nrow=4, byrow=T,
              values = c(1,0,0,0,
                         0,1,.2,0,
                         0,.2,1,0,
                         0,0,0,1),
              free=c(F,F,F,F,
                     F,F,T,F,
                     F,T,F,F,
                     F,F,F,F),
              labels=c("varx",NA,NA,NA,
                       NA,"varm1","covm1m2",NA,
                       NA,"covm1m2","varm2",NA,
                       NA,NA,NA,"vary"
              ),
              name="S")

## Extract the values and draw the model
Amatrix <- A$values
dimnames(Amatrix) <- list(labels, labels)
Amatrix

Smatrix <- S$values
dimnames(Smatrix) <- list(labels, labels)
Smatrix

Fmatrix <- diag(4)

incorrect.plot <- ramModel(Amatrix, Smatrix, Fmatrix, manNames=varnames)

## All the directions are incorrect.
semPaths(incorrect.plot, layout="tree2", sizeMan=8, edge.color = "black", 
         whatLabels = "hide", color="yellow")
```

## Correct (intended to use) model specification in analyzing the data
* The following is the correct model specified in Figure 1 in Yu et al.
* There is no correlation between the residuals in the generating model. We include it to follow the model specified in Yu et al.
```{r}
A <- create.mxMatrix(c(0,0,0,0,
                       "0.3*betaxm1",0,0,0,
                       "0.3*betaxm2",0,0,0,
                       0,"0.3*betam1y","0.3*betam2y",0), 
                     ncol=4, nrow=4, byrow=TRUE, name="A")

S <- create.mxMatrix(c(1,
                       0,"0.2*varm1",
                       0,"0.1*covm1m2","0.2*varm2",
                       0,0,0,"0.2*vary"), 
                     type="Symm", ncol=4, nrow=4, byrow=TRUE, name="S")

## Extract the values and draw the model
Amatrix <- A$values
dimnames(Amatrix) <- list(labels, labels)
Amatrix

Smatrix <- S$values
dimnames(Smatrix) <- list(labels, labels)
Smatrix

correct.plot <- ramModel(Amatrix, Smatrix, Fmatrix, manNames=varnames)

## All the directions are correct.
semPaths(correct.plot, layout="tree2", sizeMan=8, edge.color = "black", 
         whatLabels = "hide", color="lightseagreen")
```

# Figure 2 in Yu et al. (2016)

## Incorrect (actually used) population correlation used to generate the data
* The same errors were also observed in Figure 2 in Yu et al. Pearson correlations were incorrectly used to represent the path coefficients to generate the random correlation matrices.
* The incorrect population correlation matrix `IncorrectP2` was used to generate data in Run 6 Study 1 and their online Supplemental-Material-2.docx. 
* If we use `IncorrectP2` (the population values) to fit the proposed path model in Figure 2, the value of the discrepancy function is non-zero ($\chi^2(df=8)=1,276.262$ with $N=1,000$). Moreover, the residuals of the "covariance" matrix are non-zero. This shows that the population correlation matrix `IncorrectP2` does not match the model in Figure 2.

```{r}
labels <- c("x1","x2","x3","m1","m2","y1","y2")

## These are the values used as the population correlation matrix 
## to generate the data in Figure 2
rho <- 0.3
r <- c(0,0,rho,rho,0,0,0,rho,rho,0,0,rho,rho,0,0,0,rho,rho,rho,rho,0)
IncorrectP2 <- lav_matrix_vech_reverse(r, diagonal=FALSE)
diag(IncorrectP2) <- 1
dimnames(IncorrectP2) <- list(labels, labels)

kable(IncorrectP2)

## Population model: no direct effect used in the analysis
model3 <- 'm1 + m2 ~ x1 + x2 + x3
           y1 + y2 ~ m1 + m2
           y1 ~~ 0*y2'

## Incorrect model. The fit is not perfect even the population correlation matrix is used.
fit.incorrect3 <- sem(model3, sample.cov=IncorrectP2, sample.nobs=1000)
summary(fit.incorrect3)  

## Residuals of the "covariance" matrix
resid(fit.incorrect3)
```

* To see what the actual generating model is, we add the direct effects and the correlated residuals. The model is now saturated. The results show that there are negative direct effects (-0.391) from $x1$, $x2$ and $x3$ to $y1$ and $y2$. Moreover, there are correlated residuals between $m1$ and $m2$, and between $y1$ and $y2$. This model is different from the one in Figure 2 in Yu et al.

```{r}
## Population model: no direct effect used in the analysis
model4 <- 'y1 + y2 + m1 + m2 ~ x1 + x2 + x3
           y1 + y2 ~ m1 + m2 
           m1 ~~ m2
           y1 ~~ y2'

## Incorrect model. The fit is not perfect even population correlation matrix is used.
fit.incorrect4 <- sem(model4, sample.cov=IncorrectP2, sample.nobs=1000)
summary(fit.incorrect4)

## Residuals of the "covariance" matrix
resid(fit.incorrect4)

semPaths(fit.incorrect4, what="est", edge.label.cex=1.5, 
         sizeMan=8, color="yellow", edge.color = "black", 
         weighted=FALSE, layout="tree2")
```
```{r, echo=FALSE, eval=FALSE}
my.tab <- parTable(fit.incorrect4)

cat(lavaan2dot(my.tab), file="temp.txt")
```

## Correct (intended to use) population correlation used to generate the data
* The following R code shows how to generate the correct correlation matrix for Figure 2 in Yu et al.
* We specify the regression paths in the `A4` matrix. Then we generate the population correlation matrix with the `impliedR()` function.
* When we fit the model in Figure 2 to the population correlation matrix `CorrectP2`, the discrepancy is exact zero ($\chi^2(df=8)=0$ with $N=1,000$). The residuals of the "covariance" matrix are zero. The parameters are identical to the values in Figure 2. Therefore, this is the correct population correlation matrix.
```{r}
## A matrix for the regression paths as defined in Figure 2
## All of them are fixed values.
A4 <- matrix(c(0,0,0,0,0,0,0,
               0,0,0,0,0,0,0,
               0,0,0,0,0,0,0,
               0.3,0.3,0.3,0,0,0,0,
               0.3,0.3,0.3,0,0,0,0,
               0,0,0,0.3,0.3,0,0,
               0,0,0,0.3,0.3,0,0), ncol=7, nrow=7, byrow=TRUE,
             dimnames=list(labels, labels))
A4

## The variances of x1 to x3 are fixed at 1 whereas the others are starting values.
S4 <- Diag(c(1,1,1,"0.1*Err_m1","0.1*Err_m2","0.1*Err_y1", "0.1*Err_y2"))
dimnames(S4) <- list(labels, labels)
S4

CorrectP2 <- impliedR(A4, S4, labels=labels)$SigmaObs
kable(CorrectP2)

fit.correct2 <- sem(model3, sample.cov=CorrectP2, sample.nobs=1000)
summary(fit.correct2)  

## Residuals of the "covariance" matrix
resid(fit.correct2)

semPaths(fit.correct2, what="est", edge.label.cex=1.5, 
         sizeMan=8, color="lightseagreen", edge.color = "black", 
         weighted=FALSE, layout="tree2")
```
```{r, echo=FALSE, eval=FALSE}
my.tab <- parTable(fit.correct2)

cat(lavaan2dot(my.tab), file="temp.txt")
```

## Incorrect (actually used) model specification in analyzing the data
* The roles of independent and dependent variables in the `A` matrix in `OpenMx` and `metaSEM` were incorrectly reversed in Yu et al. Moreover, the error variances were incorrectly fixed at 1.0 in the `S` matrix. Therefore, the results were incorrect in Figure 2 in Yu et al.
* The following was the model Yu et al. used in their simulation study.
```{r}
varnames <- c('X1','X2','X3','M1','M2','Y1','Y2')

rho <- 0.3

A <- mxMatrix('Full', ncol=7, nrow=7, byrow=T,
              values = c(0,0,0,rho,rho,0,0,
                         0,0,0,rho,rho,0,0,
                         0,0,0,rho,rho,0,0,
                         0,0,0,0,0,rho,rho,
                         0,0,0,0,0,rho,rho,
                         0,0,0,0,0,0,0,
                         0,0,0,0,0,0,0),
              free=c(F,F,F,T,T,F,F,
                     F,F,F,T,T,F,F,
                     F,F,F,T,T,F,F,
                     F,F,F,F,F,T,T,
                     F,F,F,F,F,T,T,
                     F,F,F,F,F,F,F,
                     F,F,F,F,F,F,F
              ),
              labels=c(NA,NA,NA,"betax1m1","betax1m2",NA,NA,
                       NA,NA,NA,"betax2m1","betax2m2",NA,NA,
                       NA,NA,NA,"betax3m1","betax3m2",NA,NA,
                       NA,NA,NA,NA,NA,"betam1y1","betam1y2",
                       NA,NA,NA,NA,NA,"betam2y1","betam2y2",
                       NA,NA,NA,NA,NA,NA,NA,
                       NA,NA,NA,NA,NA,NA,NA
              ),
              name="A")

S <- mxMatrix('Full',ncol=7,nrow=7,byrow=T,
              values = c(1,0,0,0,0,0,0,
                         0,1,0,0,0,0,0,
                         0,0,1,0,0,0,0,
                         0,0,0,1,0.1,0,0,
                         0,0,0,0.1,1,0,0,
                         0,0,0,0,0,1,0.1,
                         0,0,0,0,0,0.1,1
              ),
              free=c(F,F,F,F,F,F,F,
                     F,F,F,F,F,F,F,
                     F,F,F,F,F,F,F,
                     F,F,F,F,T,F,F,
                     F,F,F,T,F,F,F,
                     F,F,F,F,F,F,T,
                     F,F,F,F,F,T,F
              ),
              labels=c("varx1",NA,NA,NA,NA,NA,NA,
                       NA,"varx2",NA,NA,NA,NA,NA,
                       NA,NA,"varx3",NA,NA,NA,NA,
                       NA,NA,NA,"varm1","covm1m2",NA,NA,
                       NA,NA,NA,"covm1m2","varm2",NA,NA,
                       NA,NA,NA,NA,NA,"vary1","covy1y2",
                       NA,NA,NA,NA,NA,"covy1y2","vary2"
              ),
              name="S")

## Extract the values and draw the model
Amatrix <- A$values
dimnames(Amatrix) <- list(labels, labels)
Amatrix

Smatrix <- S$values
dimnames(Smatrix) <- list(labels, labels)
Smatrix

Fmatrix <- diag(7)

incorrect.plot <- ramModel(Amatrix, Smatrix, Fmatrix, manNames=varnames)

## All the directions are incorrect.
semPaths(incorrect.plot, layout="tree2", sizeMan=8, edge.color = "black", 
         whatLabels = "hide", color="yellow")
```

## Correct (intended to use) model specification in analyzing the data
* The following is the correct model for Figure 2 in Yu et al.
* There is no correlation between the residuals in the generating model. We include them to follow the model specified in Yu et al.
```{r}
varnames <- c('X1','X2','X3','M1','M2','Y1','Y2')

A <- create.mxMatrix(c(0,0,0,0,0,0,0,
                       0,0,0,0,0,0,0,
                       0,0,0,0,0,0,0,
                       "0.3*betax1m1","0.3*betax2m1","0.3*betax3m1",0,0,0,0,
                       "0.3*betax1m2","0.3*betax2m2","0.3*betax3m2",0,0,0,0,
                       0,0,0,"0.3*betam1y1","0.3*betam2y1",0,0,
                       0,0,0,"0.3*betam1y2","0.3*betam2y2",0,0), 
                     ncol=7, nrow=7, byrow=TRUE, name="A")

S <- create.mxMatrix(c(1,
                       0,1,
                       0,0,1,
                       0,0,0,"0.2*varm1",
                       0,0,0,"0.1*covm1m2","0.2*varm2",
                       0,0,0,0,0,"0.2*vary1",
                       0,0,0,0,0,"0.1*covy1y2","0.2*vary2"), 
                     type="Symm", ncol=7, nrow=7, byrow=TRUE, name="S")

## Extract the values and draw the model
Amatrix <- A$values
dimnames(Amatrix) <- list(labels, labels)
Amatrix

Smatrix <- S$values
dimnames(Smatrix) <- list(labels, labels)
Smatrix

correct.plot <- ramModel(Amatrix, Smatrix, Fmatrix, manNames=varnames)

## All the directions are correct.
semPaths(correct.plot, layout="tree2", sizeMan=8, edge.color = "black", 
         whatLabels = "hide", color="lightseagreen")

sessionInfo()
```
