---
title: "Correlation-based versus parameter-based MASEM"
author: "Mike W.-L. Cheung and Shu Fai Cheung"
date: "19 Feb 2016"
output:
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: no
  pdf_document:
    toc: no
  word_document: default
---

This file includes examples used in Cheung and Cheung (2016):

Cheung, M.W.-L., and Cheung, S.-F. (2016). Random effects models for meta-analytic structural equation modeling: Review, issues, and illustrations. *Research Synthesis Methods*. Manuscript accepted for publication.

# Installing the R packages
* R can be downloaded at http://www.r-project.org/.
* Before running the analyses, we need to install some R packages and download the data. 
* We only need to install them once.
```{r, eval=FALSE}
## Installing R packages from the CRAN
install.packages(c("semPlot", "lavaan", "metaSEM"))
```

# Illustration 1: World Values Survey

## Show some sample data
```{r, message=FALSE}
## Load the libraries for TSSEM
library("metaSEM")
library("semPlot")
library("lavaan")

## Display some data
head(wvs94b$data)

## Display some sample sizes
head(wvs94b$n)
```

## Correlation-based MASEM
```{r}
## First stage of TSSEM: Random-effects model
rand1 <- tssem1(wvs94b$data, wvs94b$n, method="REM")

## Show the summary
summary(rand1)

## Plot the effect sizes
plot.meta(rand1, axis.labels=c("Cor(LS,JS)", "Cor(LS,JA)", "Cor(JS,JA)"), 
          study.min.cex=0.5, main="Effect sizes and their CIs")

## Build a model with indirect effect
## A1: Regression coefficients; 
## parameters are labelled with "a", "b", and "c"
## S1: Variance-covariance of parameters
A1 <- create.mxMatrix(c(0,"0.2*b","0.2*c",
                        0,0,"0.2*a",
                        0,0,0), 
    	                type="Full", ncol=3, nrow=3, as.mxMatrix=FALSE, byrow=TRUE)
dimnames(A1) <- list(c("LS","JS","JA"), c("LS","JS","JA"))
A1

S1 <- create.mxMatrix(c("0.2*Error_y","0.2*Error_m",1), type="Diag", as.mxMatrix=FALSE)
dimnames(S1) <- list(c("LS","JS","JA"), c("LS","JS","JA"))
S1

## Second stage of TSSEM: Random-effects model
## Estimate direct and indirect effects
rand2 <- tssem2(rand1, Amatrix=A1, Smatrix=S1, 
                diag.constraints=TRUE, intervals.type="LB",
                mx.algebras=list( indirect=mxAlgebra(a*b, name="indirect"), 
                                  direct=mxAlgebra(c, name="direct") ))

## Show the summary
summary(rand2)

## Plot the model
my.plot1 <- meta2semPlot(rand2)
semPaths(my.plot1, whatLabels="est", edge.label.cex=1.5, sizeMan=8, color="yellow")
```

## Parameter-based MASEM
```{r}
## First stage of analysis: Extract the direct 
## and indirect effects per study
(indirect1 <- indirectEffect(wvs94b$data, wvs94b$n))

## Second stage of analysis: Multivariate meta-analysis on 
## the direct and indirect effect sizes
indirect2 <- meta(indirect1[, c("ind_eff", "dir_eff")], 
                  indirect1[, c("ind_var", "ind_dir_cov", "dir_var")], intervals.type="z")

## Show the summary
summary(indirect2)  

## Plot the indirect and direct effects as effect sizes
plot.meta(indirect2, axis.labels=c("Indirect effect", "Direct effect"),
          study.min.cex=0.5, main="Effect sizes and their CIs")
```

# Illustration 2: Theory of Planned Behavior

## Show some sample data and check the data
```{r}
## Display some data
head(Cheung00$data)

## Display some sample sizes
head(Cheung00$n)

# Check pd in the data
pd_check <- is.pd(Cheung00$data)
sum(!is.na(pd_check))

## Studies with positive definite matrices happen to be studies without missing data.
my.pd <- list()
my.pd$data <- Cheung00$data[!is.na(pd_check)]
my.pd$n <- Cheung00$n[!is.na(pd_check)]

# No. of studies: All data
pattern.na(Cheung00$data, show.na=FALSE)

# No. of studies: Positive definite
pattern.na(my.pd$data, show.na=FALSE)

# Sample sizes: All data
pattern.n(Cheung00$data, Cheung00$n)

# Sample sizes: Positive definite
pattern.n(my.pd$data, my.pd$n)
```

## Correlation-based MASEM
```{r}
## Stage 1 analysis: Random-effects model
## All studies used.
tra_random_1 <- tssem1(Cheung00$data, Cheung00$n, method="REM")
summary(tra_random_1)

## Variable labels
labels <- c("att","sn","bi","beh")

## Complete mediation model
tra_S <- Diag(c("1", "1", ".2*e_bi", ".2*e_beh"))
tra_S[2,1] <- tra_S[1,2] <- ".2*c_att_sn"
dimnames(tra_S) <- list(labels, labels)
tra_S

tra_A <- matrix(c("0","0","0","0",
                  "0","0","0","0",
                  ".2*att2bi", ".2*sn2bi", "0", "0",
                  "0", "0", ".2*bi2beh", "0"),
                byrow=TRUE, 4, 4)
dimnames(tra_A) <- list(labels, labels)
tra_A

## Stage 2
tra_random_2 <- tssem2(tra_random_1, Amatrix=tra_A, Smatrix=tra_S, 
                       intervals.type="LB", diag.constraints=FALSE)
summary(tra_random_2)

my.plot2 <- meta2semPlot(tra_random_2)
semPaths(my.plot2, whatLabels="est", rotation=2, edge.label.cex=1.5, sizeMan=8, 
         color="yellow")

## Stage 1
# Only use studies with positive definite matrices
## Not enough data to estimate the variance component for all random effects
## A diagonal matrix is used
tra_random_pd_1 <- tssem1(my.pd$data, my.pd$n, method="REM", RE.type="Diag")
summary(tra_random_pd_1)

## Stage 2
tra_random_pd_2 <- tssem2(tra_random_pd_1, Amatrix=tra_A, Smatrix=tra_S, 
                          intervals.type="LB", diag.constraints=FALSE)
summary(tra_random_pd_2)

my.plot3 <- meta2semPlot(tra_random_pd_2)
semPaths(my.plot3, whatLabels="est", rotation=2, edge.label.cex=1.5, sizeMan=8,
         color="yellow")
```

## Parameter-based MASEM
```{r, message=FALSE}
## Stage 1 analysis

# Use lavaan to the model
tra_model <- 'bi  ~ att + sn
              beh ~ bi
              att ~~ sn'
  
tra_study_lavaan <- function(id, data, para_names=NULL, model, ...) {
    cor_i   <- data$data[[id]]
    n_i     <- data$n[[id]]
    fit_i <- sem(model=model, sample.cov=cor_i, sample.nobs=n_i,  ...)
    results <- list()
    coefs <- standardizedSolution(fit_i)$est.std[1:4]
    vcoefs <- vcov(fit_i)[1:4,1:4]
    if (is.null(para_names)) para_names <- names(coefs)
    names(coefs) <- para_names
    colnames(vcoefs) <- rownames(vcoefs) <- para_names
    results$coefs <- coefs
    results$vcoefs <- vcoefs
    results$fit <- fit_i
    results
}  
  
# Analyze complete data with positive definite matrices
k <- length(my.pd$data)
para_names <- c("att2bi","sn2bi","bi2beh","c_att_sn")
var_names <- c("att", "sn", "bi", "beh")

tra_fit_all <- lapply(1:k, FUN=tra_study_lavaan, data=my.pd, para_names=para_names,
                      model=tra_model, estimator="ML", fixed.x=FALSE)                
                            
## Parameter-based: Stage 2: Random and Fixed         
tra_coefs_all <- t(sapply(1:k, function(x) {tra_fit_all[[x]]$coefs}))
tra_vcoefs_all <- t(sapply(1:k, function(x) {lav_matrix_vech(tra_fit_all[[x]]$vcoefs)}))

## Not enough data to estimate the variance component for all random effects
## A diagonal matrix is used
tra_pbased_random <- meta(tra_coefs_all, tra_vcoefs_all,
                          RE.constraints=Diag(paste("0.2*Tau2_",1:4,"_", 1:4, sep="")))
summary(tra_pbased_random)

tra_pbased_fixed  <- meta(tra_coefs_all, tra_vcoefs_all, 
                          RE.constraints=matrix("0", ncol=4, nrow=4))
summary(tra_pbased_fixed)

# Compare All Methods

## Parameter estimates
coef_cbased_random  <- coef(tra_random_2)[para_names]
coef_cbased_random_pd  <- coef(tra_random_pd_2)[para_names]
coef_pbased_random  <- coef(tra_pbased_random)[1:4]

coef_all <- as.data.frame(rbind(coef_cbased_random,
                                coef_cbased_random_pd,
                                coef_pbased_random))
```

```{r, message=FALSE}
library(lavaan)

## Estimated sampling variances of parameter estimates
### Make the results reproducible.
set.seed(98717341)                                
secoef_cbased_random <- sqrt(diag(vcov(tra_random_2))[1:4])
secoef_cbased_random_pd <- sqrt(diag(vcov(tra_random_pd_2))[1:4])
secoef_pbased_random <- summary(tra_pbased_random)$coefficients$Std.Error[1:4]

secoef_all <- as.data.frame(rbind(secoef_cbased_random,
                                  secoef_cbased_random_pd,
                                  secoef_pbased_random))
                                  
## Compare them
coef_all
secoef_all

## Variance components
## Cbased_random
c_labels <- lav_matrix_vech(outer(labels, labels, 
                                  function(x,y) paste("c_",x, "_", y,sep="")), 
                            diagonal = FALSE)

cbased.vc <- vec2symMat( coef(tra_random_1, select="random"), diag=TRUE, byrow=TRUE )
dimnames(cbased.vc) <- list(c_labels, c_labels)
cbased.vc

cbased.vc_pd <- diag(coef(tra_random_pd_1, select="random"))
dimnames(cbased.vc_pd) <- list(c_labels, c_labels)
cbased.vc_pd

## Pbased_random
pbased.vc <- diag( coef(tra_pbased_random, select="random") )
dimnames(pbased.vc) <- list(para_names, para_names)
pbased.vc

## Chi-squares and p values of the parameter-based models
tra_model_fit <- t(sapply(tra_fit_all, function(x) { 
                                       fitMeasures(x$fit, c("chisq", "pvalue", "ntotal"))}))
row.names(tra_model_fit) <- names(my.pd$data)
options(scipen=50)
round(tra_model_fit,4)
```

```{r}
sessionInfo()
```
