---
title: 'Illustration of MASEM: Theory of planned behavior'
author: "Mike Cheung and Ryan Hong"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
---

This file includes examples used in Cheung and Hong (2017):
Cheung, M. W.-L., & Hong, R. Y. (2017). Applications of meta-analytic structural equation modeling in health psychology: Examples, issues, and recommendations. *Health Psychology Review*, *11*(3), 265-279. http://doi.org/10.1080/17437199.2017.1343678

# Displaying some descriptive statistics of the data
```{r, message=FALSE}
## Load the library required for the analysis
library("metaSEM")

## Check whether the correlation matrices are valid (positive definite)
## Since the correlation matrix in Study 3 is not valid, we excluded it from the analyses.
is.pd(Cooke16$data)

## Since the correlation matrix in Study 3 is not positive definite,
## we may exclude it the following analyses
my.data <- Cooke16$data[-3]
my.n <- Cooke16$n[-3]

## Show the no. of studies per correlation
pattern.na(my.data, show.na = FALSE)

## Show the total sample sizes per correlation
pattern.n(my.data, my.n)
```

# Stage 1 analysis
```{r, message=FALSE}
## Fixed-effects model: method="FEM"
fixed1 <- tssem1(my.data, my.n, method="FEM")
summary(fixed1)

## Common correlation matrix under a fixed-effects model
coef(fixed1)

## Random-effects model
random1 <- tssem1(my.data, my.n, method="REM", RE.type="Diag")
summary(random1)

## Average correlation matrix under a random-effects model
vec2symMat(coef(random1, select="fixed"), diag = FALSE)
```

# Stage 2 analysis: original model (Model A)
* The models can specified in the [RAM formulation](http://openmx.psyc.virginia.edu/docs/OpenMx/latest/Examples_Matrix.html). Matrices `A`, `S` and `F` represent the regression coefficients, variance-covariance matrix of variables, and selection matrix of the latent variables, respectively. We may skip the `F` matrix when there is no latent variable.
* Alternatively, we may also use a subset of the [lavaan syntax](http://lavaan.ugent.be/tutorial/syntax1.html).
* Consider the element `SN2BI` in `A1` as an example. It is the regression coefficient from `SN` to `BI` with a starting value of 0.2.
```{r, message=FALSE}
## lavaan synatx
modelA <- "# Regression paths
           BEH ~ BI + PBC
           BI ~ SN + ATT + PBC
           # Fix the variances of independent variables to 1
           SN ~~ 1*SN
           ATT ~~ 1*ATT
           PBC ~~ 1*PBC
           # Free the correlations among the independent variables
           SN ~~ ATT
           SN ~~ PBC
           ATT ~~ PBC"

## Plot the model
plot(modelA)

## Convert the lavaan model to RAM specification
RAMA <- lavaan2RAM(modelA, obs.variables = c("SN", "ATT", "PBC", "BI", "BEH"),
                   A.notation="on", S.notation="with")
RAMA
```

* The columns `lbound` and `ubound` are the lower and upper bounds of the 95% confidence intervals. If they include 0, the parameter estimates are not statistically significant at $\alpha=.05$.
```{r}
ModelA <- tssem2(random1, RAM=RAMA, intervals.type="LB")
summary(ModelA)
```

* To facilitate the interpretations, we may print the model with the parameter estimates using the `semPlot` package.
```{r, warning=FALSE}
## Plot the parameter estimates
plot(ModelA, color="yellow")
```

# Stage 2 analysis: no direct effect from PBC (Model C)
```{r, warning=FALSE}
modelC <- "# Regression paths
           BEH ~ BI
           BI ~ SN + ATT + PBC
           # Fix the variances of independent variables to 1
           SN ~~ 1*SN
           ATT ~~ 1*ATT
           PBC ~~ 1*PBC
           # Free the correlations among the independent variables
           SN ~~ ATT
           SN ~~ PBC
           ATT ~~ PBC"

## Plot the model
plot(modelC)

## Convert the lavaan model to RAM specification
RAMC <- lavaan2RAM(modelC, obs.variables = c("SN", "ATT", "PBC", "BI", "BEH"),
                   A.notation="on", S.notation="with")
RAMC

ModelC <- tssem2(random1, RAM=RAMC, intervals.type="LB")
summary(ModelC)

## Plot the parameter estimates
plot(ModelC, color="yellow")
```

# Stage 2 analysis: all direct effects (Model B)
```{r}
## Delete the direct effect from PBC to BEH
modelB <- "# Regression paths
           BEH ~ BI + SN + ATT + PBC
           BI ~ SN + ATT + PBC
           # Fix the variances of independent variables to 1
           SN ~~ 1*SN
           ATT ~~ 1*ATT
           PBC ~~ 1*PBC
           # Free the correlations among the independent variables
           SN ~~ ATT
           SN ~~ PBC
           ATT ~~ PBC"

## Plot the model
plot(modelB)

## Convert the lavaan model to RAM specification
RAMB <- lavaan2RAM(modelB, obs.variables = c("SN", "ATT", "PBC", "BI", "BEH"),
                   A.notation="on", S.notation="with")
RAMB
```

* We may calculate functions of parameters with the `mx.algebras` argument. In this example, we calculate the indirect and direct effects. After the analysis, 95% CI on the indirect and direct effects are estimated.
```{r, warning=FALSE}
ModelB <- tssem2(random1, RAM=RAMB, intervals.type="LB",
                 mx.algebras=list(indirect=mxAlgebra((BIonSN+BIonATT+BIonPBC)*BEHonBI, name="indirect"),
                                  direct=mxAlgebra(BEHonSN+BEHonATT+BEHonPBC, name="direct")))
summary(ModelB)

## Plot the parameter estimates
plot(ModelB, color="yellow")

## Settings of R for the analysis
sessionInfo()
```
