---
title: 'Supplemental materials of "A Tutorial on Fitting Flexible Meta-Analytic Models with Structural Equation Modeling in R"'
author: "Mike Cheung"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
  word_document: default
editor_options: 
  chunk_output_type: console
---

# Univariate Meta-Analyses

## Data preparation
```{r, message=FALSE}
## Load the required libraries
library(metaSEM)
library(symSEM)

## Prepare a sample data
dat1 <- data.frame(yi=Jaramillo05$r, 
                   vi=Jaramillo05$r_v,
                   xi=scale(Jaramillo05$IDV, scale=FALSE),
                   zi=scale(Jaramillo05$Citations, scale=TRUE))
```

## Models without any external variables
### Fixed-effect model
```{r}
## Fixed-effect model with a latent variable ei
m1a <- "ei =~ 1*yi           ## e as a latent variable and yi is its indicator
        yi ~ mu*1            ## Mean(yi) = mu
        yi ~~ 0*yi           ## Set the error variance of yi at 0
        ei ~~ data.vi*ei     ## Var(yi) = vi, which is assigned from data
"

## Fixed-effects model without specifying ei
m1b <- "yi ~ mu*1            ## Mean(yi) = mu
        yi ~~ data.vi*yi     ## Var(yi) = vi, which is assigned from data
"

## Display the conceptual model using the semPlot package
## sizeInt=6: Set 6 as the size of the intercept, see help(semPaths) in R
plot(m1a, color = "yellow", sizeInt = 7)

## Convert the lavaan model to RAM specification
## obs.variables = "yi": "yi" is the observed variable
ram1 <- lavaan2RAM(m1a, obs.variables = "yi", std.lv = FALSE)

## Display the RAM model
ram1

## Get the model-implied mean and variance structures
impliedS(ram1)

## Fit the model to the data
fit1 <- sem("Fixed", RAM = ram1, data = dat1)
summary(fit1)

## Display the model with results using the semPlot package
## nDigits=4: Display 4 digits
plot(fit1, color = "green", sizeInt = 7, nDigits = 4)
```

### Random-effects model
```{r}
## Random-effects model
m2 <- "ui =~ 1*yi          ## ui is the random effect
       ui ~~ tau2*ui       ## Var(ui) = tau2
       yi ~ mu*1           ## Mean(yi) = mu
       yi ~~ data.vi*yi    ## Var(yi) = vi, which is assigned from data
"

## Display the conceptual model
plot(m2, color="yellow", sizeInt = 7)

## Convert the lavaan model to RAM specification
## std.lv = FALSE: Ensure the latent variable is not standardized
ram2 <- lavaan2RAM(m2, obs.variables = "yi", std.lv = FALSE)

## Get the model-implied mean and variance structures
impliedS(ram2)

## intervals.type = "LB": Request likelihood-based CI
fit2 <- sem("Random", RAM = ram2, data = dat1, intervals.type = "LB")
summary(fit2)

## Display the model with results
plot(fit2, color="green", sizeInt = 7, nDigits = 4)
```

### Multiplicative-error model
```{r}
## Multiplicative error, same as UWLS model
m3 <- "ei =~ phi_sqrt*yi   ## phi (phi_sqrt^2) is the heterogeneity scaling factor
       ei ~~ data.vi*ei    ## Var(ei) = vi
       yi ~ mu*1
       yi ~~ 0*yi          ## Set the default error variance of yi at 0
       phi := phi_sqrt^2   ## Define phi as a function of phi_sqrt
"

plot(m3, color = "yellow", sizeInt = 7)

ram3 <- lavaan2RAM(m3, obs.variables = "yi", std.lv = FALSE)

impliedS(ram3)

fit3 <- sem("Multiplicative", RAM = ram3, data = dat1, intervals.type = "LB")
summary(fit3)
plot(fit3, color="green", sizeInt=6, nDigits=4)
```

### Hybrid models of additive and multiplicative errors
```{r}
## Hybrid model 1: multiplicative and addictive error
m4 <- "yi ~ mu*1
       yi ~~ 0*yi
       ## additive error
       ui =~ 1*yi
       ui ~~ tau2*ui      ## Var(ui) = tau2
       ## mulitplicative error
       ei =~ phi_sqrt*yi  ## phi is a scaling factor
       ei ~~ data.vi*ei   ## Var(ei) = vi
       phi := phi_sqrt^2  ## Define phi as a function of phi_sqrt
"
plot(m4, color="yellow", sizeInt=6)

ram4 <- lavaan2RAM(m4, obs.variables = "yi", std.lv = FALSE)

impliedS(ram4)

fit4 <- sem("Hybrid1", RAM = ram4, data = dat1, intervals.type = "LB")
summary(fit4)

plot(fit4, color="green", sizeInt=6, nDigits=4)


## Hybrid model 2: multiplicative and addictive error
m5 <- "yi ~ mu*1         ## Mean(yi) = mu
       yi ~~ Vary*yi     ## Var(yi) = vary
       ## Define Vary as a function of vi, tau2, and omega
       Vary == data.vi + tau2*data.vi^omega
"

ram5 <- lavaan2RAM(m5, obs.variables = "yi", std.lv = FALSE)

impliedS(ram5, replace.constraints = TRUE)

fit5 <- sem("Hybrid2", RAM = ram5, data = dat1, intervals.type = "LB", 
            replace.constraints = TRUE)
summary(fit5)

## Hybrid model 3: Hartung-Knapp-Sidik-Jonkman method
m6 <- "yi ~ mu*1
       yi ~~ 0*yi
       ## additive error
       ui =~ phi_sqrt*yi    ## phi is a scaling factor
       ui ~~ 0.01703407*ui  ## tau2 obtained from the random-effects model
       ## mulitplicative error
       ei =~ phi_sqrt*yi    ## phi is a scaling factor
       ei ~~ data.vi*ei     ## Var(ei) = vi
       phi := phi_sqrt^2    ## Define phi as a function of phi_sqrt
"
plot(m6, color="yellow", sizeInt=6)

ram6 <- lavaan2RAM(m6, obs.variables = "yi", std.lv = FALSE)

impliedS(ram6)

fit6 <- sem("Hybrid3", RAM = ram6, data = dat1, intervals.type = "LB")
summary(fit6)

plot(fit6, color="green", sizeInt=6, nDigits=4)
```

### Comparison of the six models
```{r}
## Refit some models to get SEs rather than LBCIs for comparisons
fit2 <- sem("Random", RAM = ram2, data = dat1)
fit3 <- sem("Multiplicative", RAM = ram3, data = dat1)
fit4 <- sem("Hybrid1", RAM = ram4, data = dat1)
fit5 <- sem("Hybrid2", RAM = ram5, data = dat1, replace.constraints = TRUE)
fit6 <- sem("Hybrid3", RAM = ram6, data = dat1)

extractResults <- function(x) {
  out <- summary(x)
  para <- out$coefficients[c("mu", "tau2", "phi_sqrt", "omega"), 
                           c("Estimate", "Std.Error")]
  out <- c(unlist(para), out$estPara, out$Minus2LL, 
           out$informationCriteria["AIC:", "Parameters Penalty"],
           out$informationCriteria["BIC:", "Parameters Penalty"])
  names(out) <- c("mu", "tau2", "phi_sqrt", "omega",
                  "mu_SE", "tau2_SE", "phi_sqrt_SE", "omega_SE",
                  "No_parameters",  "-2LL", "AIC", "BIC")
  out
}

table1 <- sapply(list(fit1, fit2, fit3, fit4, fit5, fit6), extractResults)
table1 <- data.frame(Name=rownames(table1), table1)
rownames(table1) <- NULL
colnames(table1) <- c("Name", "Fixed", "Random", "Multiplicative", "Hybrid1", 
                      "Hybrid2", "Hybrid3")
## Add the estimated tau2 and tau2_SE from the random-effects model to
## the table in an ad-hoc manner
table1[table1$Name=="tau2", "Hybrid3"] <- table1[table1$Name=="tau2", "Random"]
table1[table1$Name=="tau2_SE", "Hybrid3"] <- table1[table1$Name=="tau2_SE", "Random"]
knitr::kable(table1, digits=4, caption="Results of the Six Meta-Analytic Models")
```

## Models with external variables
### A fixed-x meta-regression
```{r, message=FALSE}
## Mixed-effects model with the covariate as a design matrix
m7 <- "yi ~ b0*1         ## Intercept of yi: b0
       yi ~~ data.vi*yi  ## Known sampling variance vi
       ui =~ 1*yi
       ui ~~ tau2*ui     ## Heterogeneity variance tau2
       ## etai: a phantom latent variable to represent the covariate x
       etai =~ b1*yi     ## yi = b1*etai
       etai ~ data.xi*1  ## etai = xi
       etai ~~ 0*etai    ## Var(etai) = 0
"

plot(m7, color = "yellow", layout = "spring")

## Only yi is an observed variable
ram7 <- lavaan2RAM(m7, obs.variables = "yi", std.lv = FALSE)

impliedS(ram7)

fit7 <- sem("Covariate_design", RAM=ram7, data=dat1, intervals.type = "LB")
summary(fit7)
plot(fit7, color = "green", sizeInt=6,  nDigits = 4)
```

### A random-x meta-regression
```{r}
## Mixed-effects model with the covariate as a variable
m8 <- "yi ~ b0*1 + b1*xi ## Intercept (b0) and slope (b1)
       yi ~~ data.vi*yi  ## Known sampling variance vi
       ui =~ 1*yi
       ui ~~ tau2*ui     ## Heterogeneity variance tau2
       xi ~ Meanx*1      ## Mean(xi) = Meanx
       xi ~~ Varx*xi     ## Var(xi) = Varx
       ## Define R2: explained variance/total variance
       R2 := b1^2*Varx/(b1^2*Varx + tau2)
"

plot(m8, color = "yellow", layout = "spring", sizeInt=6)

## Both yi and xi are observed variables
ram8 <- lavaan2RAM(m8, obs.variables = c("yi", "xi"), std.lv = FALSE)

impliedS(ram8)

## Request LBCIs on the R2
fit8 <- sem("Covariate_variable", RAM=ram8, data=dat1, intervals.type = "LB")
summary(fit8)

plot(fit8, color="green", sizeInt=6, nDigits = 4, layout = "spring")
```

### Regressing a covariate on the true effect size
```{r}
## Regressing a covariate on the true effect size
m9 <- "fyi =~ 1*yi         ## fi: true effect size
       fyi ~~ tau2*fyi     ## Var(fi) = tau2
       fyi ~ Meany*1       ## Mean(fi) = Meany
       yi ~ 0*1         
       yi ~~ data.vi*yi    ## Var(ei) = vi
       zi ~ b0*1 + b1*fyi  ## Intercept (b0) and slope (b1)
       zi ~~ ErrVarz*zi    ## Error variance of zi
       ## Define an R2
       R2 := b1^2*tau2/(b1^2*tau2 + ErrVarz)
"

plot(m9, color = "yellow", sizeInt=6, layout = "spring")

## Both yi and xi are observed variables
ram9 <- lavaan2RAM(m9, obs.variables = c("yi", "zi"), std.lv = FALSE)

impliedS(ram9)

fit9 <- sem("Regress_covariate", RAM=ram9, data=dat1, intervals.type = "LB")
summary(fit9)

plot(fit9, color = "green", sizeInt=6, nDigits = 4)
```

### Correlating a covariate with a true effect size
```{r}
## Correlation between fyi and fxi
m10 <- "yi ~ 0*1
        yi ~~ data.vi*yi
        ## fyi: Standardized true effect size of yi
        fyi =~ tauy*yi     ## tauy: sd of fyi
        fyi ~~ 1*fyi       ## Var(fyi) = 1
        fyi ~ y_mean_div_tauy*1
        ## fxi: Standardized score of xi
        fxi =~ sigma*xi    ## sigma: sd of xi
        fxi ~~ 1*fxi       ## Var(fxi) = 1
        fxi ~~ Cor*fyi
        fxi ~ x_mean_div_sigmax*1
        xi ~ 0*1
        xi ~~ 0*xi       
"

plot(m10, color = "yellow", sizeInt=6)

## Both yi and xi are observed variables
ram10 <- lavaan2RAM(m10, obs.variables = c("yi", "xi"), std.lv = TRUE)

impliedS(ram10)

fit10 <- sem("Correlation", RAM=ram10, data=dat1)
summary(fit10)
plot(fit10, color = "green", sizeInt=6, nDigits = 4)
```

### Meta-regression with completely and partially standardized variables
```{r}
## Completely standardized fyi and xi
m11 <- "yi ~ 0*1
        yi ~~ data.vi*yi
        ## fyi: Standardized true effect size of yi
        fyi =~ tauy*yi       ## tauy: sd of fyi
        fyi ~~ tau2_res*fyi  ## tau2_res: residual heterogeneity
        fyi ~ b0*1 + b1*fxi
        ## fxi: Standardized score of xi
        fxi =~ sigma*xi      ## sigma: sd of xi
        fxi ~~ 1*fxi         ## Var(fxi) =1
        fxi ~ mu_x_sd*1
        xi ~ 0*1
        xi ~~ 0*xi
        ## Nonlinear constraint to ensure Var(fyi) = 1
        tau2_res == 1 - b1^2
        ## Define R2
        R2 := b1^2
"

plot(m11, color="yellow", layout="spring", sizeInt=6)

## Both yi and xi are observed variables
ram11 <- lavaan2RAM(m11, obs.variables = c("yi", "xi"), std.lv = FALSE)

impliedS(ram11, replace.constraints = TRUE)

fit11 <- sem("Complete_standardize", RAM=ram11, data=dat1, intervals.type = "LB",
             lbound = list(tauy=0))
summary(fit11)
plot(fit11, color="green", sizeInt=6, nDigits=4)

## Partial standardized xi but not yi
m12 <- "yi ~ 0*1
        yi ~~ data.vi*yi
        ## fyi: Unstandardized true effect size of yi
        fyi =~ 1*yi
        fyi ~~ tau2_res*fyi   ## tau2_res: residual heterogeneity
        fyi ~ b0*1 + b1*fxi
        ## fxi: Standardized score of xi
        fxi =~ sigma*xi       ## sigma: sd of xi
        fxi ~~ 1*fxi          ## Var(fxi) =1
        fxi ~ mu_x_sd*1
        xi ~ 0*1
        xi ~~ 0*xi
        ## Define R2
        R2 := b1^2/(b1^2 + tau2_res)
"

plot(m12, color="yellow", layout="spring", sizeInt=6)

## Both yi and xi are observed variables
ram12 <- lavaan2RAM(m12, obs.variables = c("yi", "xi"), std.lv = FALSE)
ram12
impliedS(ram12)

fit12 <- sem("Partial_standardize", RAM=ram12, data=dat1, intervals.type = "LB")
summary(fit12)
plot(fit12, color="green", sizeInt=6, nDigits=4)
```

### Location-scale and nonlinear models
```{r}
## Location-scale model with an additive heterogeneity variance
m13 <- "yi ~ mu*1          ## Mean(yi) = mu
        yi ~~ data.vi*yi   ## Var(ei) = vi 
        ## ui is the latent variable of the random effect
        ui =~ 1*yi         ## yi = 1*ui
        ui ~~ tau2*ui      ## Var(ui) = tau2
        ## Specify nonlinear on mean and tau2
        mu == b0 + b1*data.xi
        tau2 == exp(a0 + a1*data.xi)
"

ram13 <- lavaan2RAM(m13, obs.variables="yi", std.lv = FALSE)

## Get the model implied structures
## We need to replace the constraints with the new parameters
impliedS(ram13, replace.constraints=TRUE)

fit13 <- sem("Location_additive", RAM=ram13, data=dat1,
             replace.constraints=TRUE)
summary(fit13)


## Location-scale model with a multiplicative heterogeneity variance
m14 <- "yi ~ mu*1          ## Mean(yi) = mu
        yi ~~ sigma2*yi    ## Var(yi) = sigma2 
        ## Specify nonlinear on mean and sigma2
        mu == b0 + b1*data.xi
        sigma2 == exp(a0 + a1*data.xi)*data.vi
"

ram14 <- lavaan2RAM(m14, obs.variables="yi", std.lv = FALSE)

impliedS(ram14, replace.constraints = TRUE)

fit14 <- sem("Location_multiplicative", RAM=ram14, data=dat1,
             replace.constraints=TRUE)
summary(fit14)
```

# Multivariate Meta-Analyses
## Data preparation
```{r}
## Replace missing values with 0 as NA is not allowed in definition variables
Chan17$v_misinfo[is.na(Chan17$v_misinfo)] <-0 
Chan17$c_mis_deb[is.na(Chan17$c_mis_deb)] <-0
Chan17$v_debunk[is.na(Chan17$v_debunk)] <-0

dat2 <- data.frame(y1i=Chan17$g_misinfo,
                   y2i=Chan17$g_debunk,
                   V11i=Chan17$v_misinfo,
                   V21i=Chan17$c_mis_deb,
                   V22i=Chan17$v_debunk,
                   se11i=sqrt(Chan17$v_misinfo),
                   se22i=sqrt(Chan17$v_debunk),
                   xi=scale(Chan17$PublicationYear))
```

## Models without any external variables
### Bivariate random-effects model with known sampling covariance
```{r}
## Multivariate random-effects model
m15 <- "## Means of y1 and y2
        y1i ~ mu_y1*1
        y2i ~ mu_y2*1
        ## Known sampling variances
        y1i ~~ data.V11i*y1i
        y2i ~~ data.V22i*y2i
        ## Known sampling covariance
        y1i ~~ data.V21i*y2i
        ## Random effects
        f1i =~ 1*y1i
        f2i =~ 1*y2i
        ## Heterogeneity variance-covariance matrix
        f1i ~~ T_11*f1i
        f1i ~~ T_21*f2i        
        f2i ~~ T_22*f2i
        ## Define the correlation of the random effects
        cor := T_21/sqrt(T_11*T_22)
"
plot(m15, color="yellow")

ram15 <- lavaan2RAM(m15, obs.variables = c("y1i", "y2i"), std.lv = FALSE)

impliedS(ram15)

fit15 <- sem("Mul_MA", RAM=ram15, data=dat2, intervals.type = "LB")
summary(fit15)

plot(fit15, color="green", nDigits=4)


## Multivariate random-effects model with equal means
m15b <- "## Means of y1 and y2
         y1i ~ mu_y*1
         y2i ~ mu_y*1
         ## Known sampling variances
         y1i ~~ data.V11i*y1i
         y2i ~~ data.V22i*y2i
         ## Known sampling covariance
         y1i ~~ data.V21i*y2i
         ## Random effects
         f1i =~ 1*y1i
         f2i =~ 1*y2i
         ## Heterogeneity variance-covariance matrix
         f1i ~~ T_11*f1i
         f1i ~~ T_21*f2i        
         f2i ~~ T_22*f2i
         ## Define the correlation of the random effects
         cor := T_21/sqrt(T_11*T_22)
"
plot(m15b, color="yellow")

ram15b <- lavaan2RAM(m15b, obs.variables = c("y1i", "y2i"), std.lv = FALSE)

impliedS(ram15b)

fit15b <- sem("Mul_MA_eq", RAM=ram15b, data=dat2, intervals.type = "LB")
summary(fit15b)

## Compare the two models
anova(fit15, fit15b)
```

### Bivariate random-effects model with a common between- and within-correlation
```{r}
## Multivariate random-effects model with a common between- and within-correlation
m16 <- "## Means of y1 and y2
        y1i ~ mu_y1*1
        y2i ~ mu_y2*1
        ## Error variances of y1i and y2i are set at 0 because e1i and e2i 
        ## are used to represent the sampling errors.
        y1i ~~ 0*y1i
        y2i ~~ 0*y2i
        ## e1i and e2i: sampling errors
        ## Known sampling sds
        e1i =~ data.se11i*y1i
        e2i =~ data.se22i*y2i
        ## Common correlation (within)
        e1i ~~ cor*e2i
        ## e1i and e2i variances are set at 1
        e1i ~~ 1*e1i
        e2i ~~ 1*e2i
        ## Random effect taus (SDs)
        f1i =~ sd_11*y1i
        f2i =~ sd_22*y2i
        ## Common correlation (between)
        f1i ~~ cor*f2i
        ## f1i and f2i variances are set at 1
        f1i ~~ 1*f1i
        f2i ~~ 1*f2i
        ## Compute the original T
        T_11 := sd_11^2
        T_21 := cor*sd_11*sd_22
        T_22 := sd_22^2
"
plot(m16, color="yellow")

ram16 <- lavaan2RAM(m16, obs.variables = c("y1i", "y2i"), std.lv = TRUE)

impliedS(ram16)

fit16 <- sem("Mul_no_cor", RAM=ram16, data=dat2, intervals.type = "LB")
summary(fit16)

plot(fit16, color="green", nDigits=4)
```

### Bivariate random-effects model by regressing a true effect on another true effect
```{r}
## Multivariate meta-regression: regressing f2i on f1i
m17 <- "## Means of y1 and y2
        y1i + y2i ~ 0*1
        ## Known sampling variances and covariance
        y1i ~~ data.V11i*y1i
        y2i ~~ data.V22i*y2i
        y1i ~~ data.V21i*y2i
        ## f1i and f2i: True effect sizes
        f1i =~ 1*y1i
        f2i =~ 1*y2i
        ## Heterogeneity variances
        f1i ~~ T_11*f1i
        f2i ~~ T_22*f2i
        ## Regressing f2i on f1i
        f2i ~ b0*1+ b1*f1i
        f1i ~ mu_1*1
        ## Define R2 on f2i
        R2 := b1^2*T_11/(b1^2*T_11 + T_22)
"
plot(m17, color="yellow", layout="spring")

ram17 <- lavaan2RAM(m17, obs.variables = c("y1i", "y2i"), std.lv = FALSE)

impliedS(ram17)

fit17 <- sem("Mul_meta_reg", RAM=ram17, data=dat2, intervals.type = "LB")
summary(fit17)

plot(fit17, color="green", nDigits=4, layout="spring")
```

## Models with external variables

### Bivariate meta-regression
```{r}
## Multivariate meta-regression: regressing f2i and f1i on xi
## Treating xi as a variable
# m18 <- "## Means of y1 and y2
#         y1i + y2i ~ 0*1
#         ## Known sampling variances
#         y1i ~~ data.V11i*y1i
#         y2i ~~ data.V22i*y2i
#         y1i ~~ data.V21i*y2i
#         ## f1i and f2i: True effect sizes
#         f1i =~ 1*y1i
#         f2i =~ 1*y2i
#         ## Heterogeneity variances and covariance
#         f1i ~~ T_11*f1i
#         f1i ~~ T_21*f2i
#         f2i ~~ T_22*f2i
#         ## Regressing f2i on f1i
#         f1i ~ b0_1*1 + b1_1*xi
#         f2i ~ b0_2*1 + b1_2*xi
#         xi ~ meanx*1
#         xi ~~ varx*xi
#         # Define R2s
#         R2_1 := b1_1^2*varx/(b1_1^2*varx + T_11)
#         R2_2 := b1_2^2*varx/(b1_2^2*varx + T_22)
# "

## Multivariate meta-regression: regressing f2i and f1i on xi
## Treating xi as a design matrix
m18 <- "## Means of y1 and y2
        y1i + y2i ~ 0*1
        ## Known sampling variances
        y1i ~~ data.V11i*y1i
        y2i ~~ data.V22i*y2i
        y1i ~~ data.V21i*y2i
        ## f1i and f2i: True effect sizes
        f1i =~ 1*y1i
        f2i =~ 1*y2i
        ## Heterogeneity variances and covariance
        f1i ~~ T_11*f1i 
        f1i ~~ T_21*f2i        
        f2i ~~ T_22*f2i
        ## etai: a phantom latent variable to represent the covariate x
        etai =~ 0*y1i        ## Arbitarily loaded on either y1i or y2i
        etai ~ data.xi*1     ## etai = xi
        etai ~~ 0*etai       ## Var(etai) = 0
        ## Regressing f2i on f1i
        f1i ~ b0_1*1 + b1_1*etai
        f2i ~ b0_2*1 + b1_2*etai
"

plot(m18, color="yellow", layout="tree")

ram18 <- lavaan2RAM(m18, obs.variables = c("y1i", "y2i"), std.lv = FALSE)

impliedS(ram18)

fit18 <- sem("Mul_covariate", RAM=ram18, data=dat2, intervals.type = "z")
summary(fit18)

plot(fit18, color="green", nDigits=4, layout="tree")


## Multivariate meta-regression: regressing f2i and f1i on xi
## The regression coefficients are assumed the same
m18b <- "## Means of y1 and y2
         y1i + y2i ~ 0*1
         ## Known sampling variances
         y1i ~~ data.V11i*y1i
         y2i ~~ data.V22i*y2i
         y1i ~~ data.V21i*y2i
         ## f1i and f2i: True effect sizes
         f1i =~ 1*y1i
         f2i =~ 1*y2i
         ## Heterogeneity variances and covariance
         f1i ~~ T_11*f1i 
         f1i ~~ T_21*f2i        
         f2i ~~ T_22*f2i
         ## etai: a phantom latent variable to represent the covariate x
         etai =~ 0*y1i        ## Arbitarily loading on either y1i or y2i
         etai ~ data.xi*1     ## etai = xi
         etai ~~ 0*etai       ## Var(etai) = 0
         ## Regressing f2i on f1i with the same regression coefficients
         f1i ~ b0_1*1 + b1*etai
         f2i ~ b0_2*1 + b1*etai
"

plot(m18b, color="yellow", layout="tree")

ram18b <- lavaan2RAM(m18b, obs.variables = c("y1i", "y2i"), std.lv = FALSE)

impliedS(ram18b)

fit18b <- sem("Mul_covariate_eq_beta", RAM=ram18b, data=dat2, intervals.type = "z")
summary(fit18b)

plot(fit18b, color="green", nDigits=4, layout="tree")

## Compare the two models
anova(fit18, fit18b)
```

### Mediation model with a true effect size as a mediator
```{r}
## Multivariate random-effects model: a mediation model
## Treating xi as a variable
# m19 <- "## Means of y1 and y2
#         y1i + y2i ~ 0*1
#         ## Known sampling variances
#         y1i ~~ data.V11i*y1i
#         y2i ~~ data.V22i*y2i
#         y1i ~~ data.V21i*y2i
#         ## f1i and f2i: True effect sizes
#         f1i =~ 1*y1i
#         f2i =~ 1*y2i
#         ## Heterogeneity variances
#         f1i ~~ T_11*f1i
#         f2i ~~ T_22*f2i
#         ## Regressing f2i on f1i
#         f2i ~ b*f1i + cp*xi
#         f1i ~ a*xi
#         xi ~ meanx*1
#         xi ~~ varx*xi
#         f1i ~ interceptf1*1
#         f2i ~ interceptf2*1
#         ## Define the direct, indirect, and total effects
#         direct := cp
#         indirect := a*b
#         total := a*b + cp
# "

## Multivariate random-effects model: a mediation model
## Treating xi as a design matrix
m19 <- "## Means of y1 and y2 are 0
        y1i + y2i ~ 0*1
        ## Known sampling variances
        y1i ~~ data.V11i*y1i
        y2i ~~ data.V22i*y2i
        y1i ~~ data.V21i*y2i
        ## f1i and f2i: True effect sizes
        f1i =~ 1*y1i
        f2i =~ 1*y2i
        ## Heterogeneity variances
        f1i ~~ T_11*f1i
        f2i ~~ T_22*f2i
        ## etai: a phantom latent variable to represent the covariate x
        etai =~ 0*y1i        ## Arbitarily loading on either y1i or y2i
        etai ~ data.xi*1     ## etai = xi
        etai ~~ 0*etai       ## Var(etai) = 0     
        ## Mediation model
        f1i ~ interceptf1*1 + a*etai
        f2i ~ interceptf2*1 + b*f1i + cp*etai
        ## Define the direct, indirect, and total effects
        direct := cp
        indirect := a*b
        total := a*b + cp
"
plot(m19, color="yellow", layout="tree")

ram19 <- lavaan2RAM(m19, obs.variables = c("y1i", "y2i"), std.lv = FALSE)

impliedS(ram19, replace.constraints = TRUE)

fit19 <- sem("Mul_mediation", RAM=ram19, data=dat2, intervals.type = "LB")
summary(fit19)

plot(fit19, color="green", nDigits=4, layout="tree")
```

## One-stage meta-analytic structural equation modeling (OSMASEM)

### OSMASEM using the osmasem() function
```{r}
#### Data preparation
dat_osmasem <- Cor2DataFrame(Becker92)

#### Specify the model to fit
m20 <- "Math ~ s2m*Spatial + v2m*Verbal
        Spatial ~~ r_sv*Verbal
        Spatial ~~ 1*Spatial
        Verbal ~~ 1*Verbal
"
plot(m20, col = "yellow")

## std.lv = TRUE to standardize the latent variables
ram20 <- lavaan2RAM(m20, obs.variables = c("Math", "Spatial", "Verbal"), std.lv = TRUE)

## Fit the osmasem using m20
fit.osmasem <- osmasem(RAM=ram20, data=dat_osmasem)
summary(fit.osmasem)

plot(fit.osmasem, col = "green")
```

### OSMASEM using the sem() function
```{r}
#### Data preparation
## Use the correlations and sampling covariance matrices prepared by Cor2DataFrame(Becker92)
dat3 <- dat_osmasem$data

## Use better names for the correlations and sampling covariance matrices
colnames(dat3) <- c("rS_M", "rV_M", "rV_S", 
                    "cSM_SM", "cVM_SM", "cVS_SM", "cVM_VM", "cVS_VM", "cVS_VS")

## Data set including the correlation coefficients and sampling covariance matrices
dat3

## corr = TRUE to get the model-implied correlation vector
impliedS(ram20, corr = TRUE)

#### OSMASEM using the SEM-based meta-analysis
m21 <- "## Create true effect sizes
        rhoS_M =~ 1*rS_M
        rhoV_M =~ 1*rV_M                  
        rhoV_S =~ 1*rV_S
        ## Average effect sizes
        rhoS_M ~ muS_M*1
        rhoV_M ~ muV_M*1
        rhoV_S ~ muV_S*1
        ## Heterogeneity variances
        ## As there are only a few studies, random effects are assumed to be independent.
        rhoS_M ~~ tau2S_M*rhoS_M
        rhoV_M ~~ tau2V_M*rhoV_M
        rhoV_S ~~ tau2V_S*rhoV_S
        ## Known sampling variances and covariances
        rS_M ~~ data.cSM_SM*rS_M
        rV_M ~~ data.cVM_SM*rS_M
        rV_S ~~ data.cVS_SM*rS_M
        rV_M ~~ data.cVM_VM*rV_M
        rV_S ~~ data.cVS_VM*rV_M
        rV_S ~~ data.cVS_VS*rV_S
        ## Fit the model-implied correlation vector based on impliedS(ram20, corr=TRUE)
        muS_M == r_sv*v2m + s2m
        muV_M == r_sv*s2m + v2m
        muV_S == r_sv
"

## std.lv = FALSE to keep the true effect sizes unstandardized
ram21 <- lavaan2RAM(m21, obs.variables = c("rS_M", "rV_M", "rV_S"), std.lv = FALSE)

fit21 <- sem("OSMASEM", RAM=ram21, data=dat3, replace.constraints = TRUE)
summary(fit21)

sessionInfo()
```


