---
title: 'Special Issue on Meta-analytic Structural Equation Modeling: Introduction
  from the Guest Editors'
author: "Mike W.-L. Cheung and Adam R. Hafdahl"
date: "10 Mar 2016"
output:
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
  word_document: default
---

# Introduction
*  Becker (2009) tested a mediation model on "Sport Performance" (`SP`). The independent variables were "Cognitive Anxiety" (`CA`) and "Somatic Anxiety" (`SA`) while the mediator was "Self confidence" (`SC`). The following figure displays the model with the labels of the parameters.

```{r, engine='dot', echo=FALSE, eval=TRUE, fig.align='center'}
digraph "reduced" {
   node [style=filled, shape=box, fontname="Arial", fontsize=16];
  {rank=same; CA; SA};
  SC [fillcolor="green"];
  SP [fillcolor="yellow"];
  CA -> SP [label="CA2SP"];
  CA -> SC [label="CA2SC"];  
  SA -> SP [label="SA2SP"];
  SA -> SC [label="SA2SC"];
  SC -> SP [label="SC2SP"];
  CA -> SA [dir=both, label="Cor_CA_SA"];
  CA:n -> CA:n [dir=both, label="1.0"];
  SA:n -> SA:n [dir=both, label="1.0"];
  SP:s -> SP:s [dir=both, label="Err_Var_SP"];
  SC:e -> SC:e [dir=both, label="Err_Var_SC"];
}
```


* The analysis was conducted in [R](https://www.r-project.org/) with the [metaSEM package](https://courses.nus.edu.sg/course/psycwlm/internet/metasem/). If you have not installed it yet, you may install it by using the following code in R.
```{r, eval=FALSE}
## Install the stable version from CRAN
install.packages("metaSEM")
```

* The dataset is stored as an R object `Becker09` in the metaSEM package. The correlation matrices and the sample sizes are stored as `Becker09$data` and `Becker09$n`, respectively. `NA` represents missing value.
* In this illustration, we calculated the sampling variances based on the unweighted means of the correlation matrices.

```{r, message=FALSE}
options(width=100)

## Load the library for data analysis 
library(metaSEM)

## Display the correlation matrices
Becker09$data

## Display the sample sizes
Becker09$n

## Display the numbers of studies present in each cell of the correlation matrix
pattern.na(Becker09$data, show.na = FALSE)

## Display the cumulative sample size in each cell of the correlation matrix
pattern.n(Becker09$data, Becker09$n)
```


# Fixed-effects GLS approach
## Stage 1 analysis

```{r}
## Construct a column vector y of correlation coefficients
## Extract the correlation coefficients without missing values for each study
y <- lapply(Becker09$data, function(x) {x <- vechs(x)
                                        x[complete.cases(x)]})
y

## Convert it into a column vector
y <- matrix(unlist(y), ncol=1)
y

## Create indicators on whether the correlation matrices are missing
X.design <- lapply(Becker09$data, function(x) {ifelse(is.na(vechs(x)), yes = 0, no = 1)})
X.design

## Create the design matrices
X.design <- lapply(X.design, function(x) create.Fmatrix(x, as.mxMatrix = FALSE))
X.design

## Combine into a matrix
X <- do.call(rbind, X.design)
X

## Sampling variance covariance matrix based on the unweighted mean of correlations
V <- asyCov(Becker09$data, Becker09$n, acov="unweighted", as.matrix = FALSE)

## Function to remove missing values in V: x*v*x'
filter <- function(x,v) { x %*% v %*% t(x) }

## Extract known sampling covariance matrices without the missing values
V <- mapply(filter, X.design, V)

## Convert it into a diagonal matrix
V <- bdiagMat(V)

## Only show part of V as it is too big
V[1:8, 1:8]

## Textbook formula; it may not be numerically stable
## Common correlation matrix under the fixed-effects model
r.fixed <- solve( t(X)%*%solve(V)%*%X ) %*% t(X) %*% solve(V) %*% y
r.fixed

## Sampling covariance matrix of r.fixed
V.fixed <- solve( t(X)%*%solve(V)%*%X )
V.fixed

## Q statistic in testing the homogeneity of correlation matrices
Q.fixed <- t(y) %*% (solve(V)-solve(V)%*% X %*%solve(t(X)%*%solve(V)%*%X) %*% t(X) %*%solve(V)) %*% y
Q.fixed

## df for the Q statistic
Q.df <- sum(vechs(pattern.na(Becker09$data, show.na = FALSE))) - 6
Q.df

## p value of the Q statistic
pchisq(Q.fixed, df=Q.df, lower.tail=FALSE)
```

## Stage 2 analysis
* The models were specified in the [RAM formulation](http://openmx.psyc.virginia.edu/docs/OpenMx/latest/Examples_Matrix.html) and analyzed with the `wls()` function.
```{r}
## A matrix for regression coefficients
A1 <- create.mxMatrix(c(0, "0.1*CA2SP", "0.1*SA2SP", "0.1*SC2SP",
                        0, 0, 0, 0,
                        0, 0, 0, 0,
                        0, "0.1*CA2SC", "0.1*SA2SC",0),
                      type="Full", byrow=TRUE, ncol=4, nrow=4,
                      as.mxMatrix=FALSE)

## Variable names
var.names <- c("Performance","Cognitive","Somatic","Self confidence")

## This step is not necessary but it is useful for inspecting the model.
dimnames(A1) <- list(var.names, var.names)
A1

## S matrix for variance and covariance matrix
## Since the correlation matrix is used, the variances of the predictors are fixed at 1.0.
S1 <- create.mxMatrix(c("0.1*Err_Var_SP",
                        0, 1,
                        0, "0.1*Cor_CA_SA", 1,
                        0, 0, 0, "0.1*Err_Var_SC"), byrow=TRUE, type="Symm",
                      as.mxMatrix=FALSE)

## This step is not necessary but it is useful for inspecting the model.
dimnames(S1) <- list(var.names, var.names)
S1

## Total sample size
n.total <- sum(Becker09$n)

## Convert r vector to a matrix
R.fixed <- vec2symMat(r.fixed, diag = FALSE)
R.fixed

gls.fixed2 <- wls(Cov=R.fixed, asyCov=V.fixed, n=n.total, Amatrix=A1, Smatrix=S1, 
                  model.name="Fixed effects GLS 2nd stage with direct effects")
summary(gls.fixed2)
```

* We also illustrated how to fit a model without the direct effects from `CA` and `SA` to `SP`. We created a new matrix `A2` to represent the regression coefficients. The matrix `S1` was the same as the one we used before.
```{r}
## A matrix for regression coefficients with the direct effects
A2 <- create.mxMatrix(c(0, 0, 0, "0.1*SC2SP",
                        0, 0, 0, 0,
                        0, 0, 0, 0,
                        0, "0.1*CA2SC", "0.1*SA2SC",0),
                      type="Full", byrow=TRUE, ncol=4, nrow=4,
                      as.mxMatrix=FALSE)

## This step is not necessary but it is useful for inspecting the model.
dimnames(A2) <- list(var.names, var.names)
A2

gls.fixed3 <- wls(Cov=R.fixed, asyCov=V.fixed, n=n.total, Amatrix=A2, Smatrix=S1, 
                  model.name="Fixed effects GLS 2nd stage without direct effect)")
summary(gls.fixed3)
```


# Random-effects GLS approach
## Stage 1 analysis
* Following Becker (2009), we only estimated the diagonals of the random effects. For completeness, we repeated the process by reading the data.

```{r}
## Extract the correlation coefficients without missing values
y <- lapply(Becker09$data, function(x) {x <- vechs(x); x[complete.cases(x)]})

## Convert it into a column vector
y <- matrix(unlist(y), ncol=1)
y

## Extract whether the correlation matrices are missing
X.design <- lapply(Becker09$data, function(x) {ifelse(is.na(vechs(x)), yes = 0, no = 1)})

## Create the design matrices
X.design <- lapply(X.design, function(x) create.Fmatrix(x, as.mxMatrix = FALSE))

## Combine into a matrix
X <- do.call(rbind, X.design)
X
```

* We estimated the heterogeneity variance $\tau^2$ `Tau2` on each correlation coefficient with the maximum likelihood (ML) estimation method.
```{r}
## Stack the correlation coefficients
y.vec <- t(sapply(Becker09$data, function(x) vechs(x)))
y.vec

## Extract the data by effect sizes (by column)
y.vec <- lapply(1:6, function(x) y.vec[, x] )
y.vec

## Function to estimate the tau^2 with ML estimation method
est.tau2 <- function(x, n) {
  ## remove missing values in x
  y <- x[!is.na(x)]
  ## unweighted mean of the correlation coefficients
  r <- mean(x, na.rm = TRUE)
  ## sampling variances based on the unweighted mean
  v <- (1-r^2)^2/n[!is.na(x)]
  ## meta-analysis with ML estimation
  out <- meta(y, v)
  ## extract the tau2
  coef(out, select = "random")
}

## Digaonal matrix of random effects (tau2)
Tau2 <- diag(sapply(y.vec, est.tau2, n=Becker09$n))
Tau2

## Sampling variance covariance matrix based on the unweighted mean of correlations
V <- asyCov(Becker09$data, Becker09$n, acov="unweighted", as.matrix = FALSE)

## Function to filter the elements with missing data and add Tau2
filter <- function(x,v) { x %*% (v+Tau2) %*% t(x) }

## New V = old V + Tau2
V <- mapply(filter, X.design, V)

## Convert it into a diagonal matrix
V <- bdiagMat(V)

## Show part of V
V[1:8, 1:8]

## Average correlation matrix under the random-effects model
r.random <- solve(t(X)%*%solve(V)%*%X) %*% t(X) %*% solve(V) %*% y
r.random

## Sampling covariance matrix of r.random
V.random <- solve(t(X)%*%solve(V)%*%X)
V.random
```

## Stage 2 analysis
* We fitted the model with the direct effects from `CA` and `SA` to `SP`.
```{r}
## Total sample size
n.total <- sum(Becker09$n)

## Convert r vector to a matrix
R.random <- vec2symMat(r.random, diag = FALSE)
R.random

gls.random2 <- wls(Cov=R.random, asyCov=V.random, n=n.total, Amatrix=A1, Smatrix=S1, 
                   model.name="Random effects GLS 2nd stage with direct effects")

summary(gls.random2)
```

* We fitted the model without the direct effects from `CA` and `SA` to `SP`.
```{r}
gls.random3 <- wls(Cov=R.random, asyCov=V.fixed, n=n.total, Amatrix=A2, Smatrix=S1, 
                   model.name="Random effects GLS 2nd stage without direct effect")
summary(gls.random3)
```

# Fixed-effects TSSEM approach
## Stage 1 analysis
* Since the fixed-effects TSSEM uses a multiple-group structural equation modeling approach, it does not need to estimate the sampling covariance matrix of the correlation coefficients.

```{r}
tssem.fixed1 <- tssem1(Becker09$data, Becker09$n, method="FEM", 
                       model.name="Fixed effects TSSEM 1st stage")
summary(tssem.fixed1)

## Estimated common correlation matrix
coef(tssem.fixed1)
```

## Stage 2 analysis
* We fitted the model with the direct effects from `CA` and `SA` to `SP`.
```{r}
tssem.fixed2 <- tssem2(tssem.fixed1, Amatrix=A1, Smatrix=S1, 
                       model.name="Fixed effects TSSEM 2nd stage with direct effects")
summary(tssem.fixed2)
```

* We fitted the model without the direct effects from `CA` and `SA` to `SP`.
```{r}
tssem.fixed3 <- tssem2(tssem.fixed1, Amatrix=A2, Smatrix=S1, 
                       model.name="Fixed effects TSSEM 2nd stage without direct effect)")
summary(tssem.fixed3)
```


# Random-effects TSSEM approach
## Stage 1 analysis
```{r}
## Run the analysis with the unweighted correlation matrx to calculate 
## the sampling covariance matrices of the correlation coefficients.
tssem.random1 <- tssem1(Becker09$data, Becker09$n, method="REM", RE.type="Diag",
                        acov="unweighted", model.name="Random effects TSSEM 1st stage")
summary(tssem.random1)

## Estimated average correlation matrix
vec2symMat( coef(tssem.random1, select="fixed"), diag = FALSE)
```

## Stage 2 analysis
* We fitted the model with the direct effects from `CA` and `SA` to `SP`.
```{r}
tssem.random2 <- tssem2(tssem.random1, Amatrix=A1, Smatrix=S1, 
                        model.name="Random effects TSSEM 2nd stage with direct effects")
summary(tssem.random2)
```

* We fitted the model without the direct effects from `CA` and `SA` to `SP`.
```{r}
tssem.random3 <- tssem2(tssem.random1, Amatrix=A2, Smatrix=S1, 
                       model.name="Random effects TSSEM 2nd stage without direct effect)")
summary(tssem.random3)
```


# Summary of parameter estimates and and their standard errors
## Tables
* We combined the results for the ease of comparisions for the model with direct effects.
```{r}
gls.fixed.est <- coef(gls.fixed2)
gls.fixed.se <- sqrt(diag(vcov(gls.fixed2)))

gls.random.est <- coef(gls.random2)
gls.random.se <- sqrt(diag(vcov(gls.random2)))

tssem.fixed.est <- coef(tssem.fixed2)
tssem.fixed.se <- sqrt(diag(vcov(tssem.fixed2)))

tssem.random.est <- coef(tssem.random2)
tssem.random.se <- sqrt(diag(vcov(tssem.random2)))

para <- data.frame(gls.fixed=gls.fixed.est, tssem.fixed=tssem.fixed.est, 
                   gls.random=gls.random.est, tssem.random=tssem.random.est)

se <- data.frame(gls.fixed=gls.fixed.se, tssem.fixed=tssem.fixed.se, 
                 gls.random=gls.random.se, tssem.random=tssem.random.se)
## Arrange rows according to para
se <- se[row.names(para), ]

knitr::kable(para, digits=4, caption="Parameter estimates of various methods (model with direct effects)")

knitr::kable(se, digits=4, caption="Standard errors of various methods (model with direct effects)")
```

* We combined the results for the ease of comparisions for the model without direct effect.
```{r}
gls.fixed.est <- coef(gls.fixed3)
gls.fixed.se <- sqrt(diag(vcov(gls.fixed3)))

gls.random.est <- coef(gls.random3)
gls.random.se <- sqrt(diag(vcov(gls.random3)))

tssem.fixed.est <- coef(tssem.fixed3)
tssem.fixed.se <- sqrt(diag(vcov(tssem.fixed3)))

tssem.random.est <- coef(tssem.random3)
tssem.random.se <- sqrt(diag(vcov(tssem.random3)))

para <- data.frame(gls.fixed=gls.fixed.est, tssem.fixed=tssem.fixed.est, 
                   gls.random=gls.random.est, tssem.random=tssem.random.est)

se <- data.frame(gls.fixed=gls.fixed.se, tssem.fixed=tssem.fixed.se, 
                 gls.random=gls.random.se, tssem.random=tssem.random.se)
## Arrange rows according to para
se <- se[row.names(para), ]

knitr::kable(para, digits=4, caption="Parameter estimates of various methods (model without direct effects)")

knitr::kable(se, digits=4, caption="Standard errors of various methods (model without direct effects)")
```

## Figures
* We plotted the output of the model with the direct effects based on the TSSEM approach.
```{r, message=FALSE}
## Install the library if it is not available in R
# install.packages("semPlot")

## Load the library to plot the figure
library(semPlot)

## Convert the model to semPlotModel object
my.plot1 <- meta2semPlot(tssem.random2, manNames=c("SP", "CA", "SA", "SC"))

## Plot the parameter estimates
semPaths(my.plot1, whatLabels="est", nCharNodes=10, edge.label.cex=1.2,
         layout="spring", color="green")
```

* We may plot the output of the model without the direct effects based on the TSSEM approach.
```{r}
## Convert the model to semPlotModel object
my.plot2 <- meta2semPlot(tssem.random3, manNames=c("SP", "CA", "SA", "SC"))

## Plot the parameter estimates
semPaths(my.plot2, whatLabels="est", nCharNodes=10, edge.label.cex=1.2, color="yellow")

## R settings for the analyses
sessionInfo()
```