---
title: "Comparisons of random split with with the World Values Survey data"
author: "Mike Cheung and Suzanne Jak"
date: "1 Apr 2016"
output:
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
  word_document: default
---


# Multiple regression: Fixed-effects model
* We randomly split the data into *k*=1 (raw data) to *k*=1000 *studies*.
* We regress *satisfaction with your life* (A170) on *subjective state of health* (A009), *freedom of choice and control* (A173), *satisfaction with financial situation of household* (C006), *sex* (X001), and *age* (X003) in each *study*.
* The following figure displays the regression model.

```{r, echo=FALSE, message=FALSE}
library("semPlot")
library("lavaan")
set.seed(1000)
my.df <- data.frame(A170=rnorm(100), A009=rnorm(100), A173=rnorm(100),
                    C006=rnorm(100), X001=rnorm(100), X003=rnorm(100))
my.model <- "A170~y1*A009+y2*A173+y3*C006+y4*X001+y5*X003"
my.fit <- sem(my.model, data=my.df)
semPaths(my.fit, what="mod", nCharNodes=5, edge.label.cex=1.3, col="yellow")  
```

* The estimated regression coefficients with their estimated sampling covariance matrices are treated as multiple effect sizes for a multivariate fixed-effects meta-analysis.
* The variables used in this demonstration are:
    + *State of health (subjective)* (A009): 1 (Very good); 4 (Very poor) (it is reversed before the analyses)
    + *Satisfaction with your life* (A170): 1 (Dissatisfied); 10 (Satisfied)
    + *How much freedom of choice and control* (A173): 1 (None at all); 10 (A great deal)
    + *Satisfaction with financial situation of household* (C006): 1 (None at all); 10 (A great deal)
    + *Sex* (X001): 1 (Male); 2 (Female)
    + *Age* (X003)
    + Negative values in the original dataset represent missing values. They are recoded into missing values (NA) before the analysis.
```{r, message=FALSE, cache=FALSE}
## Load the libraries
library("data.table")
library("lavaan")
library("metaSEM")

## Load the data
load("WVS.Rdata")
  
## Select the relevant variables to minimize the memory usage
WVS <- WVS[, list(A009, A170, A173, C006, X001, X003, S002, S003)]

## Reverse coding for A009
## Recode all negative values as NA
## Age (X003) is divided by 10 to improve numerical stability.
WVS[, `:=`(A009 = 5-ifelse(A009 < 0, yes=NA, no=A009),
           A170 =   ifelse(A170 < 0, yes=NA, no=A170),
           A173 =   ifelse(A173 < 0, yes=NA, no=A173),
           C006 =   ifelse(C006 < 0, yes=NA, no=C006),
           X001 =   ifelse(X001 < 0, yes=NA, no=X001),
           X003 =   ifelse(X003 < 0, yes=NA, no=X003/10))]
  
## x: no. of studies
random_split <- function(x) {
  ## Make a copy of WVS
  my.df <- copy(WVS)

  ## Randomly split the data into 100 studies
  Study <- sample(1:nrow(my.df)) %% x + 1

  ## Append "Study" into the dataset
  my.df[, Study:=Study]

  ## Set "Study" as the key for grouping
  setkeyv(my.df, "Study")

## Function to fit regression analysis
## y1 to y5: Regression coefficients from A170, A009, A173, C006, X001, and X003.
## v11 to v55: Sampling covariance matrix of the parameter estimates
  fun.reg <- function(dt) { fit <- try(lm(A170~A009+A173+C006+X001+X003, data=dt), silent=TRUE)

                            ## If there are errors during the analysis, it returns missing values.
                            if (is.element("try-error", class(fit))) {
                              list(y1=NaN,y2=NaN,y3=NaN,y4=NaN,y5=NaN,
                                   v11=NaN,v21=NaN,v31=NaN,v41=NaN,v51=NaN,
                                   v22=NaN,v32=NaN,v42=NaN,v52=NaN,v33=NaN,
                                   v43=NaN,v53=NaN,v44=NaN,v54=NaN,v55=NaN)
                            } else {
                              ## Extract the regression coefficients excluding the intercept
                              y <- coef(fit)
                              ## Extract the sampling covariance matrix excluding the intercept
                              v <- lav_matrix_vech(vcov(fit)[-1,-1])
                              list(y1=y[2],y2=y[3],y3=y[4],y4=y[5],y5=y[6],
                                   v11=v[1],v21=v[2],v31=v[3],v41=v[4],v51=v[5],
                                   v22=v[6],v32=v[7],v42=v[8],v52=v[9],v33=v[10],
                                   v43=v[11],v53=v[12],v44=v[13],v54=v[14],v55=v[15])
                              }
                            }

  ########## Split data by "Study" and analyze data with the fun.reg() function on each "Study"
  FEM1.reg <- my.df[, fun.reg(.SD), by=list(Study)]

  ########## Meta-analyze results with a multivariate fixed-effects meta-analysis:
  ########## Variance component is fixed at 0: RE.constraints=matrix(0, ncol=5, nrow=5)
  FEM2.reg <- meta(y=cbind(y1,y2,y3,y4,y5),
                   v=cbind(v11,v21,v31,v41,v51,v22,v32,v42,v52,v33,v43,v53,v44,v54,v55),
                   data=FEM1.reg, RE.constraints=matrix(0, ncol=5, nrow=5),
                   model.name="Regression analysis FEM")
  FEM2.reg
}

## Set seed for replicability
set.seed (871139100)
my.reg <- lapply(list(5,10,50,100,500,1000), random_split)

my.coeff <- sapply(my.reg, coef)
my.se <- sapply(my.reg, function(x) sqrt(diag(vcov(x))))

## Analysis of raw data (k=1)
my.lm <- lm(A170~A009+A173+C006+X001+X003, data=WVS)

my.coef <- cbind( coef(my.lm)[-1], my.coeff )
colnames(my.coef) <- paste0("k=",c(1,5,10,50,100,500,1000))

my.se <- cbind( sqrt(diag(vcov(my.lm)))[-1], my.se )
colnames(my.se) <- paste0("k=",c(1,5,10,50,100,500,1000))

write.csv(round(my.coef,4), file="coefficients.csv")
write.csv(round(my.se,4), file="SEs.csv")

knitr::kable(my.coef, digits=4, caption="Regression coefficients")
knitr::kable(my.se, digits=4, caption="Standard errors")
```

# Settings of the R system
```{r}
sessionInfo()
```
