---
title: '(Appendix B) Example 2: Testing equality of direct effects in a path model (random-effects analysis)'
author: 'Suzanne Jak and Mike W.-L. Cheung'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
  word_document: default
---

## Install the metaSEM package
* R can be downloaded at http://www.r-project.org/.
* We only need to install the metaSEM-package once.
```{r, eval=FALSE}
install.packages("metaSEM")
```

## Read in the data and load metaSEM package
```{r, message=FALSE}
library(metaSEM)

## Try to use multiple cores in the analyses
mxOption(key='Number of Threads', value=parallel::detectCores()-2)

## Load the functions to facilitate analysis (link is anonymised for review)
source("subgroup.functions.R")

head(Roorda11$data)
head(Roorda11$n)
head(Roorda11$SES)
```

## Stage 1 random-effects model on all data
```{r}
## Stage 1 analysis overall (random)
stage1random <- tssem1(Cov = Roorda11$data, n = Roorda11$n, method = "REM", RE.type = "Diag")
summary(stage1random)
```

## Stage 2 random-effects model on all data
* Create a matrix with direct effects (A) and a matrix with variances and covariances (S)

```{r}
varnames <- c("pos","neg","enga","achiev")

A <- create.mxMatrix(c( 0,0,0,0,
                        0,0,0,0,
                        "0.1*b31","0.1*b32",0,0,
                        0,0,"0.1*b43",0),
                      type = "Full", nrow = 4, ncol = 4, byrow = TRUE,
                      name = "A", as.mxMatrix = FALSE)
dimnames(A) <- list(varnames, varnames)
A

S <- create.mxMatrix(c(1,
                       ".5*p21",1,
                       0,0,"0.6*p33",
                       0,0,0,"0.6*p44"), 
                     type="Symm", byrow = TRUE,
                     name="S", as.mxMatrix = FALSE)
dimnames(S) <- list(varnames, varnames)
S
```

* Fitting the Stage 2 model on the pooled correlation matrix from the random effects Stage 1 analysis

```{r}
# Run the Stage 2 model
Stage2 <- tssem2(stage1random, Amatrix=A, Smatrix=S, diag.constraints=TRUE, intervals="LB", 
                 mx.algebras = list(Indirect_pos=mxAlgebra(b31*b43, name = "Indirect_pos"),
                                    Indirect_neg=mxAlgebra(b32*b43, name = "Indirect_neg")))

# Rerun it to remove the error code
Stage2 <- rerun(Stage2)

summary(Stage2)
```

# Subgroup analysis
* We split the studies based on the % SES

```{r}
# Data for studies with majority low SES 
data_low <- Roorda11$data[Roorda11$SES>50]
n_low <- Roorda11$n[Roorda11$SES>50]

# Data for studies with majority high SES
data_high <- Roorda11$data[Roorda11$SES<=50]
n_high <- Roorda11$n[Roorda11$SES<=50]
```

## Fitting a random-effects Stage 1 model in both subgroups

```{r, message=FALSE}
# Stage 1 analysis per subgroup (random-effects analysis)
stage1_low.fit <- tssem1(Cov = data_low, n = n_low, method = "REM", RE.type = "Diag")
stage1_high.fit <- tssem1(Cov = data_high, n = n_high, method = "REM", RE.type = "Diag")

## Rerun it to remove the error code
stage1_high.fit <- rerun(stage1_high.fit)

summary(stage1_low.fit)
summary(stage1_high.fit)
```

## Fitting the Stage 2 model in both subgroups

```{r, message=FALSE}
# Stage 2 analysis per subgroup (random-effect analysis)

stage2_low.fit <- tssem2(stage1_low.fit, Amatrix=A, Smatrix=S)

stage2_high.fit <- tssem2(stage1_high.fit, Amatrix=A, Smatrix=S)
summary(stage2_low.fit)
summary(stage2_high.fit)
```


## Testing the equality of regression coefficients
* We create and fit a model with equal direct effects (we use the same matrix A for both groups),but different variances and covariances, so we create an S matrix with different labels for the 'high' group

```{r}
S_high <- create.mxMatrix(c(1,
                            ".1*p21_high",1,
                             0,0,"0.6*p33_high",
                             0,0,0,"0.6*p44_high"), 
                      type="Symm", byrow = TRUE,
                      name="S", as.mxMatrix = FALSE)
S_high

# Create the models for the two groups, make sure to set the argument run=FALSE
stage2_low <- tssem2(stage1_low.fit, Amatrix=A, Smatrix=S, run=FALSE, model.name="low")

stage2_high <- tssem2(stage1_high.fit, Amatrix=A, Smatrix=S_high, run=FALSE, model.name="high")

# Create the multigroup model
stage2_constrained <- mxModel(model="same_regression_coef", stage2_low, stage2_high,
                              mxFitFunctionMultigroup(c("low", "high")))

# Fit multigroup model with equality constraints
Stage2_constrained.fit <- mxRun(stage2_constrained)

# first make a list of the fitted models in the separate groups
submodels.fit <- list(stage2_low.fit,stage2_high.fit)

subgroup.summary(submodels.fit,Stage2_constrained.fit)

sessionInfo()
```

