---
title: '(Appendix A) Example 1: Testing equality of factor loadings in CFA (fixed-effects analysis)'
author: 'Suzanne Jak and Mike W.-L. Cheung'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
  word_document: default
---

# Installing the metaSEM package
* R can be downloaded at http://www.r-project.org/.
* We only need to install the metaSEM-package once.
```{r, eval=FALSE}
install.packages("metaSEM")
```

# Read in the data and load metaSEM package
```{r, message=FALSE}
library(metaSEM)

## Try to use multiple cores in the analyses
mxOption(key='Number of Threads', value=parallel::detectCores()-2)

## Load the functions to facilitate analysis
source("subgroup.functions.R")

# The data "Norton13" are available after loading the metaSEM package
# We display correlation matrices for first (nr 1) and the last (nr 28) study
# With the round()-function we limit the number of decimals to 2
round(Norton13$data[[1]], 2)
round(Norton13$data[[28]], 2)

# Display sample sizes
Norton13$n

# Display population type
Norton13$population
```


# Stage 1 analysis overall 
```{r, cache=FALSE}

# Running Stage 1 model and getting the summary takes +- 4 minutes
# depending on the speed of the computer.
# The argument method = "FEM" indicates that we apply the fixed effects model
# For a random effects model one would use method = "REM"

Stage1.fit <- tssem1(Norton13$data, Norton13$n, method = "FEM")

# if the model did not converge (if OpenMx status is not 0 or 1), you may use the rerun() function to continue trying until convergence is reached

Stage1.fit <- rerun(Stage1.fit)
summary(Stage1.fit)

```

# Stage 2 analysis overall
* Specify the A-matrix with factor loadings, the S-matrix with variances and covariances, and the F-matrix in selecting observed variables. See Cheung (2015) or McArdle & McDonald (1984) for an explanation of the A, S and F matrices in the RAM formulation.

```{r}
varnames <- paste("x", 1:14, sep="")
factornames <- c("Dist","Anx","Dep")

# Vector to indicate whether variable is observed (1) or latent (0)
F <- create.Fmatrix(c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,0,0,0), name="F", as.mxMatrix=FALSE)

dimnames(F) <- list(varnames,c(varnames,factornames))

## Display F
F

# matrix with factor loadings, later included in matrix A
lambda <-matrix(
		c("0.1*L1_1","0.1*L1_2","0",
		  "0.1*L2_1","0","0.1*L2_3",
		  "0.1*L3_1","0.1*L3_2","0",
		  "0.1*L4_1","0","0.1*L4_3",
		  "0.1*L5_1","0.1*L5_2","0",
		  "0.1*L6_1","0","0.1*L6_3",
		  "0.1*L7_1","0.1*L7_2","0",
		  "0.1*L8_1","0","0.1*L8_3",
		  "0.1*L9_1","0.1*L9_2","0",
		  "0.1*L10_1","0","0.1*L10_3",
		  "0.1*L11_1","0.1*L11_2","0",
		  "0.1*L12_1","0","0.1*L12_3",
		  "0.1*L13_1","0.1*L13_2","0",
		  "0.1*L14_1","0","0.1*L14_3"), 
	    	nrow=14, ncol=3, byrow = TRUE)

A <- rbind(cbind(matrix(0,ncol=14,nrow=14),lambda),
           matrix(0, nrow=3, ncol=17))

dimnames(A) <- list(c(varnames,factornames),c(varnames,factornames))

## Display A
A

## Convert to OpenMx-matrices
A <- as.mxMatrix(A)

# Matrix with residual variances
psi <- matrix(0, nrow = 14, ncol = 14)
diag(psi) <- paste("0.1*","e",1:14,sep="")

# Matrix with factor variances and covariances
phi <- matrix(c(1,0,0,
                0,1,0,
                0,0,1),
              nrow = 3, ncol = 3)

# Psi and phi are combined in matrix S
S <- bdiagMat(list(psi, phi))

dimnames(S) <- list(c(varnames,factornames),c(varnames,factornames))

## Display S
S

## Convert to OpenMx-matrices
S <- as.mxMatrix(S)
```

* Fitting the Stage 2 model on the pooled correlation matrix from the fixed effects Stage 1 analysis

```{r, cache=FALSE}
# Run the Stage 2 model
Stage2.fit <- tssem2(Stage1.fit, Amatrix=A, Smatrix=S, Fmatrix=F)
summary(Stage2.fit)
```

# Subgroup analysis
* We group the sample characteristics into patients versus non-patients. The variable is stored in `Norton13$group`.

```{r, cache=FALSE}
# Display the populations grouped into patients vs. non-patients
Norton13$group

# Stage 1 per subgroup
Stage1_subgroup.fit <- tssem1(Norton13$data, Norton13$n, method = "FEM", 
                              cluster = Norton13$group)

summary(Stage1_subgroup.fit)
```

## Run the Stage 2 model without equality constraints across groups
```{r, cache=FALSE}
# By providing the R-object with the Stage 1 analysis (Stage1_subgroup.fit) to the tssem2()-function, 
# the factor model is fit separately to the data of each subgroup
# Running  this model takes some minutes
Stage2_free.fit <- tssem2(Stage1_subgroup.fit, Amatrix=A, Smatrix=S, Fmatrix=F)

## summary of individual models
summary(Stage2_free.fit)
```

```{r, cache=FALSE}

## summary of combined model
subgroup.summary(Stage2_free.fit)
```


## Run the Stage 2 model with equality constraints across groups (per factor)
*  Set factor loadings Distress equal across groups
*  Give the factor loadings of Anxiety (odd items) and Depression (even items) different labels
*  Give all residual variances different labels

```{r, echo = TRUE, eval = TRUE}
# Create a vector with the item numbers for each factor
Items_Anx <- c(1,3,5,7,9,11,13)
Items_Dep <- c(2,4,6,8,10,12,14)

# we create a new matrix, A_pat1, in which the factor loadings from the Anxiety facor (in column 16 of matrix A), and of the Depression factor (in column 17 of matrix A) are different
A_pat1 <- A
A_pat1$labels[Items_Anx,16] <- paste(A$labels[Items_Anx,16],"_pat",sep="")
A_pat1$labels[Items_Dep,17] <- paste(A$labels[Items_Dep,17],"_pat",sep="")

# Give the residual variances different labels for the patient group
S_pat <- S
diag(S_pat$labels)[1:14] <- paste(diag(S$labels)[1:14],"_pat",sep="") 
```

* Create the models (set run=FALSE) and multigroup model with equal factor loadings of Distress factor
```{r, cache=FALSE, message=FALSE}
# we use the new A and S matrix in the patient group
stage2_pat <- tssem2(Stage1_subgroup.fit[[1]], Amatrix=A_pat1, Smatrix=S_pat, Fmatrix=F, 
                    run=FALSE, model.name="patients")

stage2_nonpat <- tssem2(Stage1_subgroup.fit[[2]], Amatrix=A, Smatrix=S, Fmatrix=F, 
                        run=FALSE, model.name="nonpatients")

# Combine the two separate models into a multigroup model
stage2_constrained_distr <- mxModel(model="same_factor_loadings_Distr", 
                                    stage2_nonpat, stage2_pat,
                                    mxFitFunctionMultigroup(c("patients", "nonpatients")))

Stage2_constrained_distr.fit <- mxRun(stage2_constrained_distr)
```

```{r, cache=FALSE}
# get summary using subgroup.summary() function
subgroup.summary(Stage2_free.fit, Stage2_constrained_distr.fit)
```


* Create the models (set run=FALSE) and multigroup model with equal factor loadings of Anxiety factor

```{r, cache=FALSE, message=FALSE}
A_pat2 <- A_pat1
A_pat2$labels[1:14,15] <- paste(A$labels[1:14,15],"_pat",sep="")
A_pat2$labels[Items_Anx,16] <- A$labels[Items_Anx,16]

# Create the models, make sure to set the argument run=FALSE

stage2_pat <- tssem2(Stage1_subgroup.fit[[1]], Amatrix=A_pat2, Smatrix=S_pat, Fmatrix=F, 
                     run=FALSE, model.name="patients")

stage2_nonpat <- tssem2(Stage1_subgroup.fit[[2]], Amatrix=A, Smatrix=S, Fmatrix=F, 
                        run=FALSE, model.name="nonpatients")

stage2_constrained_anx <- mxModel(model="same_factor_loadings_Anx",
                                  stage2_nonpat, stage2_pat,
                                  mxFitFunctionMultigroup(c("patients", "nonpatients")))

Stage2_constrained_anx.fit <- mxRun(stage2_constrained_anx)
```


```{r, cache = FALSE}
subgroup.summary(Stage2_free.fit, Stage2_constrained_anx.fit)
```

* Create the models (set run=FALSE) and multigroup model with equal factor loadings of Depression factor

```{r, cache=FALSE, message=FALSE}
#  factor loadings Depression equal across groups
A_pat3 <- A_pat2
A_pat3$labels[Items_Anx,16] <- A_pat1$labels[Items_Anx,16]
A_pat3$labels[Items_Dep,17] <- A$labels[Items_Dep,17]

# Create the models, make sure to set the argument run=FALSE

stage2_pat <- tssem2(Stage1_subgroup.fit[[1]], Amatrix=A_pat3, Smatrix=S_pat, Fmatrix=F, 
                     run=FALSE, model.name="patients")

stage2_nonpat <- tssem2(Stage1_subgroup.fit[[2]], Amatrix=A, Smatrix=S, Fmatrix=F, 
                        run=FALSE, model.name="nonpatients")

stage2_constrained_dep <- mxModel(model="same_factor_loadings_Dep", 
                                  stage2_nonpat, stage2_pat,
                                  mxFitFunctionMultigroup(c("patients", "nonpatients")))

Stage2_constrained_dep.fit <- mxRun(stage2_constrained_dep)
```


```{r, cache = FALSE}
subgroup.summary(Stage2_free.fit, Stage2_constrained_dep.fit)

sessionInfo()
```

