---
title: "MASEM on Gnambs et al. (2018)"
author: "Suzanne Jak and Mike Cheung"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    keep_md: yes
    self_contained: yes
    theme: united
    toc: yes
  pdf_document:
    toc: no
  word_document: default
editor_options: 
  chunk_output_type: console
---

# Data preparation
```{r}
library(metaSEM)

## Create a new dataset without missing value in Individualism
index_na <- is.na(Gnambs18$Individualism)
Gnambs18 <- lapply(Gnambs18, function(x) x[!index_na])

## Select data with the correlation matrices, i.e., exclude studies with factor loadings
index <- Gnambs18$CorMat==1
Gnambs18 <- lapply(Gnambs18, function(x) x[index])

my.df <- Cor2DataFrame(Gnambs18$data, Gnambs18$n, acov = "weighted")

## Add the standardized individualism as the moderator
## Standardization of the moderator improves the convergence.
my.df$data <- data.frame(my.df$data,
                         Individualism=scale(Gnambs18$Individualism),
                         check.names=FALSE)
summary(my.df)
```

# TSSEM
## One general factor model

```{r}
rand1 <- tssem1(Gnambs18$data, Gnambs18$n, method="REM", RE.type="Diag")
summary(rand1)

## One general factor
model1 <- "G =~ g1*I1 + g2*I2 + g3*I3 + g4*I4 + g5*I5 +
                g6*I6 + g7*I7 + g8*I8 + g9*I9 + g10*I10"

RAM1 <- lavaan2RAM(model1, obs.variables = paste0("I", 1:10))
RAM1

rand2a <- tssem2(rand1, Amatrix = RAM1$A, Smatrix = RAM1$S, Fmatrix = RAM1$F)
summary(rand2a)
plot(rand2a, col="green")
```

## Bifactor model with positive and negative Self-Esteem
```{r}
model2 <- "G =~ g1*I1 + g2*I2 + g3*I3 + g4*I4 + g5*I5 + 
                g6*I6 + g7*I7 + g8*I8 + g9*I9 + g10*I10
           POS =~ p1*I1 + p3*I3 + p4*I4 + p7*I7 + p10*I10
           NEG =~ n2*I2 + n5*I5 + n6*I6 + n8*I8 + n9*I9"

RAM2 <- lavaan2RAM(model2, obs.variables = paste0("I", 1:10))
RAM2

rand2b <- tssem2(rand1, Amatrix = RAM2$A, Smatrix = RAM2$S, Fmatrix = RAM2$F)
summary(rand2b)
plot(rand2b, col="green")

## Compare the one general factor to the bifactor model
anova(rand2b, rand2a)
```

## Models with two subgroup analysis
```{r}
# Data for studies with individualism below the mean
data_g1 <- Gnambs18$data[my.df$data$Individualism < 0 ]
n_g1 <- Gnambs18$n[my.df$data$Individualism < 0 ]

# Data for studies with individualism above or equal the mean
data_g2 <- Gnambs18$data[my.df$data$Individualism >= 0 ]
n_g2 <- Gnambs18$n[my.df$data$Individualism >= 0 ]
```

### Fitting a random-effects Stage 1 model in two subgroups
```{r, message=FALSE}
## Stage 1 analysis per subgroup (random-effects analysis)
stage1_g1.fit <- tssem1(Cov = data_g1, n = n_g1, method = "REM", RE.type = "Diag")
stage1_g2.fit <- tssem1(Cov = data_g2, n = n_g2, method = "REM", RE.type = "Diag")

summary(stage1_g1.fit)
summary(stage1_g2.fit)
```

### Fitting the Stage 2 bifactor model in both subgroups
```{r}
## Stage 2 analysis per subgroup (random-effect analysis)
stage2_g1.fit <- tssem2(stage1_g1.fit, Amatrix=RAM2$A, Smatrix=RAM2$S, Fmatrix=RAM2$F)
stage2_g2.fit <- tssem2(stage1_g2.fit, Amatrix=RAM2$A, Smatrix=RAM2$S, Fmatrix=RAM2$F)

summary(stage2_g1.fit)
summary(stage2_g2.fit)
```

# OSMASEM
## One general factor model without any moderator
```{r}
## Create matrices with implicit diagonal constraints
M0a <- create.vechsR(A0=RAM1$A, S0=RAM1$S, F0=RAM1$F)

## Create heterogeneity variances
T0a <- create.Tau2(RAM=RAM1, RE.type="Diag", Transform="expLog", RE.startvalues=0.05)

fit0a <- osmasem(model.name="No moderator", Mmatrix=M0a, Tmatrix=T0a, data=my.df)
summary(fit0a, Saturated=TRUE)

## SRMR
osmasemSRMR(fit0a)

## Show the heterogeneity variances
diag(VarCorr(fit0a))
```

## One general factor model with `Individualism` as a moderator on the A matrix
```{r, error=TRUE}
## Replace the A matrix with the moderator "Individualism"
Ax1a <- RAM1$A
Ax1a[grep("\\*", Ax1a)] <- "0*data.Individualism"
Ax1a

## Create matrices with implicit diagonal constraints
M1a <- create.vechsR(A0=RAM1$A, S0=RAM1$S, F0=RAM1$F, Ax=Ax1a)

fit1a <- osmasem(model.name="Moderator with individualism", Mmatrix=M1a, Tmatrix=T0a, data=my.df)
summary(fit1a)

## Get the R2
osmasemR2(fit1a, fit0a)

## Compare the models with and without Individualism
anova(fit1a, fit0a)

## Get the estimated A0 and A1
A0 <- mxEval(A0, fit1a$mx.fit)
A0
A1 <- mxEval(A1, fit1a$mx.fit)
A1

## Compute the estimated A matrix at -1SD (-1) of the standardized individualism
A0 - A1

## Compute the estimated A matrix at 0 (mean) of the standardized individualism
A0

## Compute the estimated A matrix at +1SD (+1) of the standardized individualism
A0 + A1
```

## Bifactor model without any moderator
```{r}
## Create matrices with implicit diagonal constraints
M0b <- create.vechsR(A0=RAM2$A, S0=RAM2$S, F0=RAM2$F)

## Create heterogeneity variances
T0b <- create.Tau2(RAM=RAM2, RE.type="Diag", Transform="expLog", RE.startvalues=0.05)

fit0b <- osmasem(model.name="No moderator", Mmatrix=M0b, Tmatrix=T0b, data=my.df)
summary(fit0b, Saturated=TRUE)

## SRMR
osmasemSRMR(fit0b)

diag(VarCorr(fit0b))
```

## Bifactor model  with `Individualism` as a moderator on the A matrix
```{r, error=TRUE}
## Replace the A matrix with the moderator "Individualism"
Ax1b <- RAM2$A
Ax1b[grep("\\*", Ax1b)] <- "0*data.Individualism"
Ax1b

## Create matrices with implicit diagonal constraints
M1b <- create.vechsR(A0=RAM2$A, S0=RAM2$S, F0=RAM2$F, Ax=Ax1b)

fit1b <- osmasem(model.name="Moderator with individualism", Mmatrix=M1b, Tmatrix=T0b, data=my.df)
summary(fit1b)

## Get the R2
osmasemR2(fit1b, fit0b)

## Compare the models with and without Individualism
anova(fit1b, fit0b)

## Get the estimated A0 and A1
A0 <- mxEval(A0, fit1b$mx.fit)
A0
A1 <- mxEval(A1, fit1b$mx.fit)
A1

## Compute the estimated A matrix at -1SD (-1) of the standardized individualism
A0 - A1

## Compute the estimated A matrix at 0 (mean) of the standardized individualism
A0

## Compute the estimated A matrix at +1SD (+1) of the standardized individualism
A0 + A1

save.image(file="Gnambs2018.RData")

sessionInfo()
```

